### 2021-03-02 manuscript/manuscript.Rmd
Initial commit

---
title: "The document title"
output: 
  pdf_document:
    includes:
      in_header: "preamble.tex"
---

\begin{center}
Lucas Veras$^1$, Florêncio Diniz-Sousa$^1$, Giorjines Boppre$^1$, Vítor Devezas$^2$, Hugo Santos-Sousa$^2$, John Preto$^2$, João Paulo Vilas-Boas$^{3,4}$, Leandro Machado$^{3,4}$, José Oliveira$^1$, Hélder Fonseca$^1$
\end{center}

\bigskip

1 - Research Center in Physical Activity, Health and Leisure (CIAFEL), Faculty of Sport, University of Porto, Porto, Portugal

2 - General Surgery Department, São João Medical Center, Porto, Portugal

3 - Center of Research, Education, Innovation and Intervention in Sport (CIFI2D), Faculty of Sport, University of Porto, Porto, Portugal

4 - Biomechanics Laboratory (LABIOMEP-UP), University of Porto, Porto, Portugal

### 2021-03-03 code/02_explore.R
Compute number of peaks descriptive statistics by locomotion type

@@ -20,6 +20,10 @@ knitr::kable(sample_size)
 # Number of peaks median and IQR ------------------------------------------
 
 n_peaks_desc <- mechanical_load_data %>%
+  mutate(
+    locomotion_type = as.factor(ifelse(speed %in% 1:6, "walking", "running"))
+  ) %>%
+  group_by(locomotion_type) %>%
   summarise(
     n_peaks_median = median(n_peaks), n_peaks_iqr = IQR(n_peaks)
   )

### 2021-03-03 figures/bmi_speed.R
Initial commit

@@ -1,0 +1,65 @@
+# Load packages -----------------------------------------------------------
+
+library(here)
+library(tidyverse)
+library(lvmisc)
+library(ggsci)
+library(patchwork)
+library(ragg)
+
+# Load data ---------------------------------------------------------------
+
+load(here("data", "mechanical_load_data.rda"))
+sample_size <- mechanical_load_data %>%
+  filter(acc_placement == "hip" & vector == "resultant") %>%
+  group_by(speed) %>%
+  select(pGRF_N) %>%
+  summarise_all(~ sum(!is.na(.)))
+
+# Plot BMI distribution per speed -----------------------------------------
+
+bmi_speed_plot <- mechanical_load_data %>%
+  filter(acc_placement == "hip" & vector == "resultant") %>%
+  ggplot(aes(x = speed, y = BMI)) +
+  geom_boxplot(outlier.shape = NA, fill = "gray", alpha = 0.6, width = 0.5) +
+  geom_jitter(alpha = 0.5) +
+  scale_y_continuous(
+    limits = c(10, 60),
+    expand = c(0, 0),
+    breaks = seq(10, 60, 10)
+  ) +
+  scale_x_discrete(
+    labels = c(
+      "2" = quote(atop(2 ~ km %.% h^-1, paste("n = 64"))),
+      "3" = quote(atop(3 ~ km %.% h^-1, paste("n = 64"))),
+      "4" = quote(atop(4 ~ km %.% h^-1, paste("n = 64"))),
+      "5" = quote(atop(5 ~ km %.% h^-1, paste("n = 61"))),
+      "6" = quote(atop(6 ~ km %.% h^-1, paste("n = 46"))),
+      "7" = quote(atop(7 ~ km %.% h^-1, paste("n = 20"))),
+      "8" = quote(atop(8 ~ km %.% h^-1, paste("n = 20"))),
+      "9" = quote(atop(9 ~ km %.% h^-1, paste("n = 20"))),
+      "10" = quote(atop(10 ~ km %.% h^-1, paste("n = 19"))),
+      "11" = quote(atop(11 ~ km %.% h^-1, paste("n = 19"))),
+      "12" = quote(atop(12 ~ km %.% h^-1, paste("n = 19"))),
+      "13" = quote(atop(13 ~ km %.% h^-1, paste("n = 19"))),
+      "14" = quote(atop(14 ~ km %.% h^-1, paste("n = 17")))
+    )
+  ) +
[…]

### 2021-03-03 code/02_explore.R
Assign ggplot objects

@@ -15,7 +15,6 @@ sample_size <- mechanical_load_data %>%
   group_by(acc_placement, vector, speed) %>%
   select(acc_placement, vector, speed, pGRF_N, pLR_Ns) %>%
   summarise_all(~ sum(!is.na(.)))
-knitr::kable(sample_size)
 
 # Number of peaks median and IQR ------------------------------------------
 
@@ -31,25 +30,25 @@ n_peaks_desc <- mechanical_load_data %>%
 # GRF and ACC magnitude and rate per running speeed -----------------------
 
 # Ground reaction force
-mechanical_load_data %>%
+GRF_speed_plots <- mechanical_load_data %>%
   filter(acc_placement == "lower_back") %>%
   ggplot(aes(x = speed, y = pGRF_BW, fill = vector)) +
   geom_boxplot()
 
 # Loading rate
-mechanical_load_data %>%
+LR_speed_plots <- mechanical_load_data %>%
   filter(acc_placement == "lower_back") %>%
   ggplot(aes(x = speed, y = pLR_BWs, fill = vector)) +
   geom_boxplot()
 
 # Acceleration
-mechanical_load_data %>%
+ACC_speed_plots <- mechanical_load_data %>%
   ggplot(aes(x = speed, y = pACC_g, fill = vector)) +
   geom_boxplot() +
   facet_wrap(~ acc_placement)
 
 # Acceleration transient rate
-mechanical_load_data %>%
+ATR_speed_plots <- mechanical_load_data %>%
   ggplot(aes(x = speed, y = pATR_gs, fill = vector)) +
   geom_boxplot() +
   facet_wrap(~ acc_placement)
@@ -72,9 +71,13 @@ correlations <- map2(
 # Scatterplots ------------------------------------------------------------
 
 # GRF x ACC
-plot_scatter(mechanical_load_data, pACC_g, pGRF_N, color = BMI_cat) +
+GRF_ACC_scatterplot <- plot_scatter(
+  mechanical_load_data, pACC_g, pGRF_N, color = BMI_cat
+) +
   facet_wrap(~ acc_placement)
 
 # LR x ATR
-plot_scatter(mechanical_load_data, pATR_gs, pLR_Ns, color = BMI_cat) +
+LR_ATR_scatterplot <- plot_scatter(
+  mechanical_load_data, pATR_gs, pLR_Ns, color = BMI_cat
+) +
   facet_wrap(~ acc_placement)

### 2021-03-03 code/01_tidy_data.R
Add sex, age and height variables to mechanical_loading_data

@@ -73,7 +73,9 @@ mechanical_load_data <- rbind(
     subj = ifelse(str_length(subj) == 4, paste0("0", subj), subj)
   ) %>%
   left_join(
-    select(anthropometric, id, trial, body_mass, BMI, BMI_cat),
+    select(
+      anthropometric, id, trial, sex, age, height, body_mass, BMI, BMI_cat
+    ),
     by = c("id", "trial")
   ) %>%
   select(
@@ -129,7 +131,9 @@ mechanical_load_rates_df <- rbind(
     subj = ifelse(str_length(subj) == 4, paste0("0", subj), subj)
   ) %>%
   left_join(
-    select(anthropometric, id, trial, body_mass, BMI, BMI_cat),
+    select(
+      anthropometric, id, trial, sex, age, height, body_mass, BMI, BMI_cat
+    ),
     by = c("id", "trial")
   ) %>%
   select(

### 2021-03-03 code/02_explore.R
Compute sample descriptive statistics

@@ -16,6 +16,25 @@ sample_size <- mechanical_load_data %>%
   select(acc_placement, vector, speed, pGRF_N, pLR_Ns) %>%
   summarise_all(~ sum(!is.na(.)))
 
+# Sample descriptives -----------------------------------------------------
+
+sample_descriptives <- mechanical_load_data %>%
+  # Filter only one data point per participant
+  filter(acc_placement == "hip" & vector == "resultant" & speed == 2) %>%
+  select(age, height, body_mass, BMI) %>%
+  summarise(
+    across(where(is.double), list(mean = mean, sd = sd))
+  ) %>%
+  round(1)
+
+sample_size_sex <- mechanical_load_data %>%
+  # Filter only one data point per participant
+  filter(acc_placement == "hip" & vector == "resultant" & speed == 2) %>%
+  select(sex) %>%
+  as_vector() %>%
+  unname() %>%
+  table()
+
 # Number of peaks median and IQR ------------------------------------------
 
 n_peaks_desc <- mechanical_load_data %>%

### 2021-03-03 manuscript/manuscript.Rmd
Add objective and begin Methods section

@@
 ---
-title: "The document title"
+# title: "The document title"
 output: 
@@
 
+```{r, setup, include=FALSE}
+knitr::opts_chunk$set(
+  echo = FALSE, results = "hide", message = FALSE, warning = FALSE
+)
+```
+
+```{r source}
+library(here)
+source(here("code", "02_explore.R"))
+```
+
+<!--
 \begin{center}
@@
 4 - Biomechanics Laboratory (LABIOMEP-UP), University of Porto, Porto, Portugal
+-->
+
+# Introduction
+
+The aim of this study was to develop pGRF and pLR prediction models based on accelerometer data during walking and running.
+
+# Methods
+
+A convenience sample of `r max(sample_size$pGRF_N)` adults (`r unname(sample_size_sex["M"])` males; $`r sample_descriptives$age_mean`\pm`r sample_descriptives$age_sd`$ years; $`r sample_descriptives$height_mean`\pm`r sample_descriptives$height_sd`$ cm; $`r sample_descriptives$body_mass_mean`\pm`r sample_descriptives$body_mass_sd`$ kg; $`r sample_descriptives$BMI_mean`\pm`r sample_descriptives$BMI_sd`$ $\mathrm{kg\cdot m^{-2}}$, $\overline{X}\pm SD$) with no self-reported neurological or musculoskeletal limitations was recruited.
+Height and body mass were assessed following standard procedures with a stadiometer and a digital scale, respectively.
+All participants were informed about the experiments purpose and protocol before giving written informed consent.
+The study was approved by the local Ethics Committee (CES 192-14).
+
+Data collection took place at Porto Biomechanics Laboratory (LABIOMEP-UP), in an instrumented treadmill with built-in FPs (AMTI Corporation, Watertown, USA).
+Before starting the data collection, participants were given time to warm up on the treadmill.
+The protocol consisted of walking and running in 12 incremental speeds, with 0º inclination, from 2 to 14 $\mathrm{km\cdot h^{-1}}$ with 1 $\mathrm{km\cdot h^{-1}}$ increment.
+Walking and running speeds ranged from 2 to 6 $\mathrm{km\cdot h^{-1}}$ and from 7 to 14 $\mathrm{km\cdot h^{-1}}$, respectively.
+Each walking speed lasted for 1 minute and each running speed for 30 seconds.
+To ensure single foot contact on each FP during the trial, feedback regarding location on the treadmill was provided by a researcher.
+Participants wore their own sports shoes throughout the trials.

### 2021-03-04 code/03_predict.R
Add speed as a random factor

@@ -28,7 +28,7 @@ ver_mechanical_load_data <- map(
 # Build GRF models --------------------------------------------------------
 
 GRF_formula <- as.formula(
-  "pGRF_N ~ pACC_g + body_mass + pACC_g:body_mass + (1 | subj)"
+  "pGRF_N ~ pACC_g + body_mass + pACC_g:body_mass + (1 | subj) + (1 | speed)"
 )
 
 # Resultant vector

### 2021-03-04 manuscript/manuscript.Rmd
Finish Methods section

@@
       in_header: "preamble.tex"
+    latex_engine: lualatex
+bibliography: bibliography.bib
+csl: nlm.csl
 ---
@@
 
-Data collection took place at Porto Biomechanics Laboratory (LABIOMEP-UP), in an instrumented treadmill with built-in FPs (AMTI Corporation, Watertown, USA).
+\medskip
+
+Data collection took place at Porto Biomechanics Laboratory (LABIOMEP-UP), in an instrumented treadmill with built-in FP (AMTI Corporation, Watertown, USA).
 Before starting the data collection, participants were given time to warm up on the treadmill.
-The protocol consisted of walking and running in 12 incremental speeds, with 0º inclination, from 2 to 14 $\mathrm{km\cdot h^{-1}}$ with 1 $\mathrm{km\cdot h^{-1}}$ increment.
+The protocol consisted of walking and running in 13 incremental speeds, with 0º inclination, from 2 to 14 $\mathrm{km\cdot h^{-1}}$ with 1 $\mathrm{km\cdot h^{-1}}$ increment.
 Walking and running speeds ranged from 2 to 6 $\mathrm{km\cdot h^{-1}}$ and from 7 to 14 $\mathrm{km\cdot h^{-1}}$, respectively.
@@
 Participants wore their own sports shoes throughout the trials.
+
+\medskip
+
+During the protocol, participants wore three activity monitors, one at their right hip (along the anterior axillary line, at the level of the iliac crest), another at their lower back (at the midpoint between the two posterior superior iliac spines) and the last at their right ankle (immediately superior to the lateral malleolus).
+Activity monitors at waist level were secured on the same elastic belt with a clip and the activity monitor placed on the ankle was secured by an elastic belt fixed with adhesive tape.
+In all placements the accelerometer y-axis (vertical) was aligned with the standing body longitudinal axis.
+All participants wore the same three activity monitors always positioned at the same places.
+
+\medskip
+
+Activity monitors used were the GT9X Link model ($\mathrm{\pm16}$\textit{g} range; ActiGraph, Pensacola, USA) that incorporates primary and secondary triaxial accelerometers.
+As a manufacturer proprietary filter is applied on the primary, but not on the secondary accelerometer raw data, only data from the latter was used in this study, reducing filtering bias and enhancing the data processing replicability.
+
+\medskip
+
+Data from FP and accelerometers were collected at 1000 Hz and 100 Hz sampling frequency, respectively, and operated through the manufacturer supplied software (for the FP: Netforce, Version 3.5.1; AMTI Corporation, Watertown, USA; for the accelerometers: ActiLife version 6.13.3; Actigraph, Pensacola, USA).
+Raw GRF (expressed in N) and ACC (expressed in gravitational ACC units; 1\textit{g} =
+9.807 $\mathrm{m\cdot s^{-2}}$) from the x, y and z vectors were exported to plain text files.
+
+\medskip
+
+FP and accelerometer data were processed using MATLAB (Version 8.3, Mathworks, Natick, USA), as described elsewhere [@Veras2020].
+Briefly, both GRF and ACC signals were filtered using a Butterworth fourth-order low-pass filter, with 20 Hz cut-off frequency, and their resultant vector calculated ($r_i = \sqrt{x_i^2 + y_i^2 + z_i^2}$).
+After that, peak ACC (pACC) values were defined as minimum height of twice the positive ACC average, during any given walking speed, and separated by 0.4 seconds minimum.
+Then, GRF and ACC signals were synchronized using the following sequence: i) signal adjustment through time set by the systems clock; ii) manual correction based on visual inspection for possible errors in time adjustment; iii) determination of pGRF as the highest values between 0.4 seconds before and after each pACC; iv) synchronization by the maximum cross-correlation coefficient of GRF and ACC peak values for both the resultant and its vertical component.
+The rates of change, LR for the GRF and ACC rate (AR) for the ACC, were computed through a second-order centered derivative from the beggining of the foot contact to the pGRF, and pLR and peak AR (pAR) were defined as the maximum value in the array.
+
+\[f_i^{'} = \frac{f_{i + 1} - f_{i - 1}}{t_{i + 1} - t_{i - 1}}\]
+
+\[f_{i_{max}}^{'} = max(f_i^{'})\]
+
+Regions of interest for data analysis were manually selected where locomotion patterns were constant, which corresponded of 20-45 seconds periods, on average.
+Then, pGRF, pACC, pLR and pAR means of the resultant and its vertical component for each participant at each speed were extracted and used in all remaining analyses.
+In walking trials, a median of `r n_peaks_desc[which(n_peaks_desc$locomotion_type == "walking"), "n_peaks_median"]` peaks (interquartile range `r n_peaks_desc[which(n_peaks_desc$locomotion_type == "walking"), "n_peaks_median"] - n_peaks_desc[which(n_peaks_desc$locomotion_type == "walking"), "n_peaks_iqr"]`\textendash`r n_peaks_desc[which(n_peaks_desc$locomotion_type == "walking"), "n_peaks_median"] + n_peaks_desc[which(n_peaks_desc$locomotion_type == "walking"), "n_peaks_iqr"]`) were used to calculate these means, while in running trials it was a median of `r n_peaks_desc[which(n_peaks_desc$locomotion_type == "running"), "n_peaks_median"]` peaks (interquartile range `r n_peaks_desc[which(n_peaks_desc$locomotion_type == "running"), "n_peaks_median"] - n_peaks_desc[which(n_peaks_desc$locomotion_type == "running"), "n_peaks_iqr"]`\textendash`r n_peaks_desc[which(n_peaks_desc$locomotion_type == "running"), "n_peaks_median"] + n_peaks_desc[which(n_peaks_desc$locomotion_type == "running"), "n_peaks_iqr"]`).
+
+\medskip
+
+Statistical analyses were conducted using R statistical software (version `r paste(R.Version()$major, R.Version()$minor, sep = ".")`, R Foundation for Statistical Computing, Vienna, Austria).
+All statistical analyses were registered in an open platform, where R code utilized and detailed information can be assessed (LINK).
+
+\medskip
+
+Regression equations to predict pRGFR, pVGRF, pRLR and pVLR were developed by linear mixed models.
+Body mass (kg) was tested as a predictor in all models.
+In addition, in the pRGRF and pVGRF models, pRACC and peak vertical ACC (pVACC; \textit{g}), respectively, were tested as predictors, while in the pRLR and pVLR models, peak resultant ATR (pRATR; $g\phantom{ }\mathrm{\phantom{ }\cdot s^{-1}}$) and peak vertical ATR (pVATR; $g\phantom{ }\mathrm{\phantom{ }\cdot s^{-1}}$), respectively, were tested as predictors.
+All predictors were tested as fixed effects and were shown to be significant.
+Subject and speed were tested as random effects, and the inclusion of both has showed to improve the models.
+Final models were chosen according to -2 log-likelihood statistics [@Field2012].
+Traditional coefficient of determination (R\textsuperscript{2}) was represented by conditional R\textsuperscript{2}, that estimates the variance explained by the whole model [@Nakagawa2013].
+
+\medskip
+
+Model was validated using the leave-one-out cross-validation approach [@Staudenmayer2012].
+For leave-one-out cross-validation each participant’s data was separated into a testing dataset (one participant at a time) with the remaining data in the training dataset.
+New linear mixed models with the same outcomes and predictors as determined for the entire sample were developed using the training dataset and then used to predict the outcome for the participant in the testing dataset.
+This process was repeated for all participants (`r max(sample_size$pGRF_N)` times).
+Data from the testing dataset was used in the remaining statistical analysis.
+
+\medskip
+
+To assess the models prediction accuracy, mean absolute error (MAE), mean absolute percent error (MAPE) and root mean square error (RMSE) were calculated.
+Also, methods agreement was evaluated with Bland-Altman plots by plotting the difference between the actual (FP measured) and predicted values against their mean.
+Bias was computed as the mean of these differences and the limits of agreement were obtained using \pm 1.96 standard deviation of the mean between actual and predicted values [@Bland1986].
+Linear regressions were applied to identify if the magnitude of the mean between actual and predicted values influenced the magnitude of their difference.
+This was done to ascertain if the prediction accuracy was constant throughout the assayed magnitude range [@Giavarina2015].
+These analyses were conducted separately for data from each accelerometer placement and outcome.
+
+\medskip
+
+Our pGRF prediction equation was compared with a previously published reference equation [@Neugebauer2014], that was developed using similar methods.
+This compaarison was performed in three ways using: i) the whole sample; ii) a subsample of walking trials; iii) a subsample of running trials.
+To assess the prediction accuracy, MAE, MAPE and RMSE were calculated using pVGRF values predicted from both equations.
+No pLR prediction equation suited to be compared with ours was found.
+
+\newpage
+
+# References

### 2021-03-04 figures/fig1.R
Change output name

@@ -54,7 +54,7 @@ bmi_speed_plot <- mechanical_load_data %>%
 # Save plot ---------------------------------------------------------------
 
 agg_tiff(
-  here("figures", "bmi_speed.tiff"),
+  here("figures", "fig1.tiff"),
   width = 60,
   height = 50,
   units = "cm",

### 2021-03-04 figures/fig1.R
Update figures

@@ -18,7 +18,7 @@ sample_size <- mechanical_load_data %>%
 
 # Plot BMI distribution per speed -----------------------------------------
 
-bmi_speed_plot <- mechanical_load_data %>%
+fig1 <- mechanical_load_data %>%
   filter(acc_placement == "hip" & vector == "resultant") %>%
   ggplot(aes(x = speed, y = BMI)) +
   geom_boxplot(outlier.shape = NA, fill = "gray", alpha = 0.6, width = 0.5) +
@@ -61,5 +61,5 @@ agg_tiff(
   res = 100,
   scaling = 2
 )
-plot(bmi_speed_plot)
+plot(fig1)
 dev.off()

### 2021-03-04 figures/fig2.R

@@ -1,0 +1,87 @@
+# Load packages -----------------------------------------------------------
+
+library(here)
+library(tidyverse)
+library(ggsci)
+library(patchwork)
+library(ragg)
+
+# Load data ---------------------------------------------------------------
+
+load(here("data", "mechanical_load_data.rda"))
+mechanical_load_data <- mechanical_load_data %>%
+  mutate(activity = as.factor(ifelse(speed %in% 1:6, "walking", "running")))
+
+# Hip pRGRF x pRACC plot --------------------------------------------------
+
+scatterplot_GRF_res_hip <- mechanical_load_data %>%
+  filter(vector == "resultant" & acc_placement == "hip") %>%
+  ggplot() +
+  geom_point(
+    aes(x = pACC_g, y = pGRF_N, color = BMI_cat, shape = activity),
+    show_guide = FALSE
+  ) +
+  geom_smooth(
+    aes(x = pACC_g, y = pGRF_N, color = BMI_cat),
+    method = "lm", se = FALSE
+  ) +
+  scale_color_nejm() +
+  scale_y_continuous(
+    limits = c(0, 3500),
+    expand = c(0, 0),
+    breaks = seq(0, 3500, 500)
+  ) +
+  theme_light() +
+  theme(
+    plot.title = element_text(hjust = 0.5),
+    legend.title = element_blank()
+  ) +
+  labs(title = "Hip", x = quote("pRACC" ~ (italic(g))), y = "pRGRF (N)")
+
+# Back pRLR x pRAR plot ---------------------------------------------------
+
+scatterplot_LR_res_back <- mechanical_load_data %>%
+  filter(vector == "resultant" & acc_placement == "lower_back") %>%
+  ggplot() +
+  geom_point(
+    aes(x = pATR_gs, y = pLR_Ns, color = BMI_cat, shape = activity),
+    show_guide = FALSE
+  ) +
+  geom_smooth(
+    aes(x = pATR_gs, y = pLR_Ns, color = BMI_cat),
+    method = "lm", se = FALSE
+  ) +
+  scale_color_nejm() +
+  scale_y_continuous(
+    limits = c(0, 60000),
+    expand = c(0, 0),
+    breaks = seq(0, 60000, 10000)
+  ) +
+  theme_light() +
+  theme(
+    plot.title = element_text(hjust = 0.5),
+    legend.title = element_blank()
+  ) +
+  labs(
[…]

### 2021-03-04 figures/fig3.R

@@ -1,0 +1,87 @@
+# Load packages -----------------------------------------------------------
+
+library(here)
+library(tidyverse)
+library(lvmisc)
+library(ggsci)
+library(patchwork)
+library(ragg)
+
+# Load data ---------------------------------------------------------------
+
+load(here("output", "loocv_data.rda"))
+cv_res_GRF_models <- map(
+  cv_res_GRF_models,
+  ~ mutate(
+    .x, activity = as.factor(ifelse(speed %in% 1:6, "walking", "running"))
+  )
+)
+cv_res_LR_models <- map(
+  cv_res_LR_models,
+  ~ mutate(
+    .x, activity = as.factor(ifelse(speed %in% 1:6, "walking", "running"))
+  )
+)
+
+# Hip pRGRF Bland-Altman plot ---------------------------------------------
+
+BA_GRF_res_hip <- cv_res_GRF_models$hip %>%
+  plot_bland_altman(color = BMI_cat, shape = activity) +
+  guides(shape = FALSE) +
+  scale_color_nejm() +
+  scale_y_continuous(
+    limits = c(-600, 600),
+    expand = c(0, 0),
+    breaks = seq(-600, 600, 200)
+  ) +
+  theme_light() +
+  theme(
+    plot.title = element_text(hjust = 0.5),
+    legend.title = element_blank()
+  ) +
+  labs(
+    title = "Hip",
+    x = "Mean of Actual and Predicted pRGRF (N)",
+    y = "Actual - Predicted pRGRF (N)"
+  )
+
+# Back pRLR Bland-Altman plot ---------------------------------------------
+
+BA_LR_res_back <- cv_res_LR_models$lower_back %>%
+  plot_bland_altman(color = BMI_cat, shape = activity) +
+  guides(shape = FALSE) +
+  scale_color_nejm() +
+  scale_y_continuous(
+    labels = scales::label_number(),
+    limits = c(-30000, 30000),
+    expand = c(0, 0),
+    breaks = seq(-30000, 30000, 10000)
+  ) +
+  theme_light() +
+  theme(
+    plot.title = element_text(hjust = 0.5),
+    legend.title = element_blank()
+  ) +
+  labs(
+    title = "Lower Back",
+    x = quote("Mean of Actual and Predicted pRLR" ~ (N %.% s^-1)),
+    y = quote("Actual - Predicted pRLR" ~ (N %.% s^-1))
+  )
+
[…]

### 2021-03-04 figures/figS2.R

@@ -12,176 +12,200 @@ load(here("data", "mechanical_load_data.rda"))
 mechanical_load_data <- mechanical_load_data %>%
   mutate(activity = as.factor(ifelse(speed %in% 1:6, "walking", "running")))
 
-# GRF x ACC plots ---------------------------------------------------------
+# LR x ATR plots ----------------------------------------------------------
 
 # Resultant: Ankle
-scatterplot_GRF_res_ankle <- mechanical_load_data %>%
+scatterplot_LR_res_ankle <- mechanical_load_data %>%
   filter(vector == "resultant" & acc_placement == "ankle") %>%
   ggplot() +
   geom_point(
-    aes(x = pACC_g, y = pGRF_N, color = BMI_cat, shape = activity),
+    aes(x = pATR_gs, y = pLR_Ns, color = BMI_cat, shape = activity),
     show_guide = FALSE
   ) +
   geom_smooth(
-    aes(x = pACC_g, y = pGRF_N, color = BMI_cat),
+    aes(x = pATR_gs, y = pLR_Ns, color = BMI_cat),
     method = "lm", se = FALSE
   ) +
   scale_color_nejm() +
   scale_y_continuous(
-    limits = c(0, 3500),
+    limits = c(0, 60000),
     expand = c(0, 0),
-    breaks = seq(0, 3500, 500)
+    breaks = seq(0, 60000, 10000)
   ) +
   theme_light() +
   theme(
     plot.title = element_text(hjust = 0.5),
     legend.title = element_blank()
   ) +
-  labs(title = "Ankle", x = quote("pRACC" ~ (italic(g))), y = "pRGRF (N)")
+  labs(
+    title = "Ankle",
+    x = quote("pRATR" ~ (italic(g) %.% s^-1)),
+    y = quote("pRLR" ~ (N %.% s^-1))
+  )
 
 # Resultant: Lower back
-scatterplot_GRF_res_back <- mechanical_load_data %>%
+scatterplot_LR_res_back <- mechanical_load_data %>%
   filter(vector == "resultant" & acc_placement == "lower_back") %>%
   ggplot() +
   geom_point(
-    aes(x = pACC_g, y = pGRF_N, color = BMI_cat, shape = activity),
+    aes(x = pATR_gs, y = pLR_Ns, color = BMI_cat, shape = activity),
[…]

### 2021-03-04 figures/figS3.R

@@ -22,18 +22,6 @@ cv_ver_GRF_models <- map(
     .x, activity = as.factor(ifelse(speed %in% 1:6, "walking", "running"))
   )
 )
-cv_res_LR_models <- map(
-  cv_res_LR_models,
-  ~ mutate(
-    .x, activity = as.factor(ifelse(speed %in% 1:6, "walking", "running"))
-  )
-)
-cv_ver_LR_models <- map(
-  cv_ver_LR_models,
-  ~ mutate(
-    .x, activity = as.factor(ifelse(speed %in% 1:6, "walking", "running"))
-  )
-)
 
 # GRF plots ---------------------------------------------------------------
 
@@ -163,146 +151,9 @@ BA_GRF_ver_hip <- cv_ver_GRF_models$hip %>%
     y = "Actual - Predicted pVGRF (N)"
   )
 
-# LR plots ----------------------------------------------------------------
-
-# Resultant: Ankle
-BA_LR_res_ankle <- ggplot()
-# BA_LR_res_ankle <- cv_res_LR_models$ankle %>%
-#   plot_bland_altman(color = BMI_cat, shape = activity) +
-#   guides(shape = FALSE) +
-#   scale_color_nejm() +
-#   scale_y_continuous(
-#     labels = scales::label_number(),
-#     limits = c(-100000, 125000),
-#     expand = c(0, 0),
-#     breaks = seq(-100000, 125000, 50000)
-#   ) +
-#   theme_light() +
-#   theme(
-#     plot.title = element_text(hjust = 0.5),
-#     legend.title = element_blank()
-#   ) +
-#   labs(
-#     title = "Ankle",
-#     x = quote("Mean of Actual and Predicted pRLR" ~ (N %.% s^-1)),
-#     y = quote("Actual - Predicted pRLR" ~ (N %.% s^-1))
-#   )
-
-# Resultant: Lower back
-BA_LR_res_back <- cv_res_LR_models$lower_back %>%
-  plot_bland_altman(color = BMI_cat, shape = activity) +
-  guides(shape = FALSE) +
-  scale_color_nejm() +
-  scale_y_continuous(
-    labels = scales::label_number(),
-    limits = c(-30000, 30000),
-    expand = c(0, 0),
-    breaks = seq(-30000, 30000, 10000)
-  ) +
-  theme_light() +
-  theme(
[…]

### 2021-03-04 figures/figS4.R

@@ -1,0 +1,182 @@
+# Load packages -----------------------------------------------------------
+
+library(here)
+library(tidyverse)
+library(lvmisc)
+library(ggsci)
+library(patchwork)
+library(ragg)
+
+# Load data ---------------------------------------------------------------
+
+load(here("output", "loocv_data.rda"))
+cv_res_LR_models <- map(
+  cv_res_LR_models,
+  ~ mutate(
+    .x, activity = as.factor(ifelse(speed %in% 1:6, "walking", "running"))
+  )
+)
+cv_ver_LR_models <- map(
+  cv_ver_LR_models,
+  ~ mutate(
+    .x, activity = as.factor(ifelse(speed %in% 1:6, "walking", "running"))
+  )
+)
+
+# LR plots ----------------------------------------------------------------
+
+# Resultant: Ankle
+BA_LR_res_ankle <- ggplot()
+# BA_LR_res_ankle <- cv_res_LR_models$ankle %>%
+#   plot_bland_altman(color = BMI_cat, shape = activity) +
+#   guides(shape = FALSE) +
+#   scale_color_nejm() +
+#   scale_y_continuous(
+#     labels = scales::label_number(),
+#     limits = c(-100000, 125000),
+#     expand = c(0, 0),
+#     breaks = seq(-100000, 125000, 50000)
+#   ) +
+#   theme_light() +
+#   theme(
+#     plot.title = element_text(hjust = 0.5),
+#     legend.title = element_blank()
+#   ) +
+#   labs(
+#     title = "Ankle",
+#     x = quote("Mean of Actual and Predicted pRLR" ~ (N %.% s^-1)),
+#     y = quote("Actual - Predicted pRLR" ~ (N %.% s^-1))
+#   )
+
+# Resultant: Lower back
+BA_LR_res_back <- cv_res_LR_models$lower_back %>%
+  plot_bland_altman(color = BMI_cat, shape = activity) +
+  guides(shape = FALSE) +
+  scale_color_nejm() +
+  scale_y_continuous(
+    labels = scales::label_number(),
+    limits = c(-30000, 30000),
+    expand = c(0, 0),
+    breaks = seq(-30000, 30000, 10000)
+  ) +
+  theme_light() +
+  theme(
+    plot.title = element_text(hjust = 0.5),
+    legend.title = element_blank()
[…]

### 2021-03-04 figures/fig1.R
Change figure settings

@@ -46,6 +46,11 @@ fig1 <- mechanical_load_data %>%
     )
   ) +
   theme_light() +
+  theme(
+    axis.title.y = element_text(size = 18),
+    axis.text.y = element_text(size = 16),
+    axis.text.x = element_text(size = 16)
+  ) +
   labs(
     x = "",
     y = quote("BMI" ~ (kg %.% m^-2))
@@ -55,7 +60,7 @@ fig1 <- mechanical_load_data %>%
 
 agg_tiff(
   here("figures", "fig1.tiff"),
-  width = 60,
+  width = 80,
   height = 50,
   units = "cm",
   res = 100,

### 2021-03-04 figures/fig2.R

@@ -19,7 +19,7 @@ scatterplot_GRF_res_hip <- mechanical_load_data %>%
   ggplot() +
   geom_point(
     aes(x = pACC_g, y = pGRF_N, color = BMI_cat, shape = activity),
-    show_guide = FALSE
+    show.legend = FALSE
   ) +
   geom_smooth(
     aes(x = pACC_g, y = pGRF_N, color = BMI_cat),
@@ -31,10 +31,20 @@ scatterplot_GRF_res_hip <- mechanical_load_data %>%
     expand = c(0, 0),
     breaks = seq(0, 3500, 500)
   ) +
+  scale_x_continuous(
+    limits = c(0.5, 6.5),
+    expand = c(0, 0),
+    breaks = seq(1, 6, 1)
+  ) +
   theme_light() +
   theme(
-    plot.title = element_text(hjust = 0.5),
-    legend.title = element_blank()
+    plot.title = element_text(size = 15, hjust = 0.5),
+    legend.title = element_blank(),
+    legend.text = element_text(size = 13),
+    axis.title.y = element_text(size = 13),
+    axis.title.x = element_text(size = 13),
+    axis.text.y = element_text(size = 13),
+    axis.text.x = element_text(size = 13)
   ) +
   labs(title = "Hip", x = quote("pRACC" ~ (italic(g))), y = "pRGRF (N)")
 
@@ -53,14 +63,24 @@ scatterplot_LR_res_back <- mechanical_load_data %>%
   ) +
   scale_color_nejm() +
   scale_y_continuous(
-    limits = c(0, 60000),
+    limits = c(0, 30000),
     expand = c(0, 0),
-    breaks = seq(0, 60000, 10000)
+    breaks = seq(0, 30000, 5000)
+  ) +
+  scale_x_continuous(
+    limits = c(0, 50),
+    expand = c(0, 0),
+    breaks = seq(5, 45, 5)
   ) +
   theme_light() +
   theme(
-    plot.title = element_text(hjust = 0.5),
-    legend.title = element_blank()
+    plot.title = element_text(size = 15, hjust = 0.5),
+    legend.title = element_blank(),
+    legend.text = element_text(size = 13),
+    axis.title.y = element_text(size = 13),
+    axis.title.x = element_text(size = 13),
+    axis.text.y = element_text(size = 13),
+    axis.text.x = element_text(size = 13)
   ) +
   labs(
     title = "Lower Back",
[…]

### 2021-03-04 figures/fig3.R

@@ -36,8 +36,13 @@ BA_GRF_res_hip <- cv_res_GRF_models$hip %>%
   ) +
   theme_light() +
   theme(
-    plot.title = element_text(hjust = 0.5),
-    legend.title = element_blank()
+    plot.title = element_text(size = 15, hjust = 0.5),
+    legend.title = element_blank(),
+    legend.text = element_text(size = 13),
+    axis.title.y = element_text(size = 13),
+    axis.title.x = element_text(size = 13),
+    axis.text.y = element_text(size = 13),
+    axis.text.x = element_text(size = 13)
   ) +
   labs(
     title = "Hip",
@@ -53,14 +58,24 @@ BA_LR_res_back <- cv_res_LR_models$lower_back %>%
   scale_color_nejm() +
   scale_y_continuous(
     labels = scales::label_number(),
-    limits = c(-30000, 30000),
+    limits = c(-20000, 20000),
     expand = c(0, 0),
-    breaks = seq(-30000, 30000, 10000)
+    breaks = seq(-20000, 20000, 10000)
+  ) +
+  scale_x_continuous(
+    limits = c(2500, 27500),
+    expand = c(0, 0),
+    breaks = seq(5000, 25000, 5000)
   ) +
   theme_light() +
   theme(
-    plot.title = element_text(hjust = 0.5),
-    legend.title = element_blank()
+    plot.title = element_text(size = 15, hjust = 0.5),
+    legend.title = element_blank(),
+    legend.text = element_text(size = 13),
+    axis.title.y = element_text(size = 13),
+    axis.title.x = element_text(size = 13),
+    axis.text.y = element_text(size = 13),
+    axis.text.x = element_text(size = 13)
   ) +
   labs(
     title = "Lower Back",
@@ -74,11 +89,14 @@ fig3 <- BA_GRF_res_hip +
   BA_LR_res_back +
   plot_annotation(tag_levels = "A") +
   plot_layout(guides = "collect") &
-  theme(legend.position = "bottom")
+  theme(
+    legend.position = "bottom",
+    plot.tag = element_text(size = 16)
+  )
 agg_tiff(
   here("figures", "fig3.tiff"),
-  width = 120,
-  height = 50,
+  width = 80,
+  height = 25,
   units = "cm",
   res = 100,
   scaling = 2

### 2021-03-04 figures/figS1.R

@@ -20,7 +20,7 @@ scatterplot_GRF_res_ankle <- mechanical_load_data %>%
   ggplot() +
   geom_point(
     aes(x = pACC_g, y = pGRF_N, color = BMI_cat, shape = activity),
-    show_guide = FALSE
+    show.legend = FALSE
   ) +
   geom_smooth(
     aes(x = pACC_g, y = pGRF_N, color = BMI_cat),
@@ -32,10 +32,20 @@ scatterplot_GRF_res_ankle <- mechanical_load_data %>%
     expand = c(0, 0),
     breaks = seq(0, 3500, 500)
   ) +
+  scale_x_continuous(
+    limits = c(0.5, 14.5),
+    expand = c(0, 0),
+    breaks = seq(1, 14, 2)
+  ) +
   theme_light() +
   theme(
-    plot.title = element_text(hjust = 0.5),
-    legend.title = element_blank()
+    plot.title = element_text(size = 15, hjust = 0.5),
+    legend.title = element_blank(),
+    legend.text = element_text(size = 13),
+    axis.title.y = element_text(size = 13),
+    axis.title.x = element_text(size = 13),
+    axis.text.y = element_text(size = 13),
+    axis.text.x = element_text(size = 13)
   ) +
   labs(title = "Ankle", x = quote("pRACC" ~ (italic(g))), y = "pRGRF (N)")
 
@@ -45,7 +55,7 @@ scatterplot_GRF_res_back <- mechanical_load_data %>%
   ggplot() +
   geom_point(
     aes(x = pACC_g, y = pGRF_N, color = BMI_cat, shape = activity),
-    show_guide = FALSE
+    show.legend = FALSE
   ) +
   geom_smooth(
     aes(x = pACC_g, y = pGRF_N, color = BMI_cat),
@@ -57,10 +67,20 @@ scatterplot_GRF_res_back <- mechanical_load_data %>%
     expand = c(0, 0),
     breaks = seq(0, 3500, 500)
   ) +
+  scale_x_continuous(
+    limits = c(0.5, 5.5),
+    expand = c(0, 0),
+    breaks = seq(1, 6, 1)
+  ) +
   theme_light() +
   theme(
-    plot.title = element_text(hjust = 0.5),
-    legend.title = element_blank()
+    plot.title = element_text(size = 15, hjust = 0.5),
+    legend.title = element_blank(),
+    legend.text = element_text(size = 13),
[…]

### 2021-03-04 figures/figS2.R

@@ -28,14 +28,24 @@ scatterplot_LR_res_ankle <- mechanical_load_data %>%
   ) +
   scale_color_nejm() +
   scale_y_continuous(
-    limits = c(0, 60000),
+    limits = c(0, 30000),
     expand = c(0, 0),
-    breaks = seq(0, 60000, 10000)
+    breaks = seq(0, 30000, 10000)
+  ) +
+  scale_x_continuous(
+    limits = c(0, 50),
+    expand = c(0, 0),
+    breaks = seq(5, 45, 5)
   ) +
   theme_light() +
   theme(
-    plot.title = element_text(hjust = 0.5),
-    legend.title = element_blank()
+    plot.title = element_text(size = 15, hjust = 0.5),
+    legend.title = element_blank(),
+    legend.text = element_text(size = 13),
+    axis.title.y = element_text(size = 13),
+    axis.title.x = element_text(size = 13),
+    axis.text.y = element_text(size = 13),
+    axis.text.x = element_text(size = 13)
   ) +
   labs(
     title = "Ankle",
@@ -57,14 +67,24 @@ scatterplot_LR_res_back <- mechanical_load_data %>%
   ) +
   scale_color_nejm() +
   scale_y_continuous(
-    limits = c(0, 60000),
+    limits = c(0, 30000),
     expand = c(0, 0),
-    breaks = seq(0, 60000, 10000)
+    breaks = seq(0, 30000, 5000)
+  ) +
+  scale_x_continuous(
+    limits = c(0, 50),
+    expand = c(0, 0),
+    breaks = seq(5, 45, 5)
   ) +
   theme_light() +
   theme(
-    plot.title = element_text(hjust = 0.5),
-    legend.title = element_blank()
+    plot.title = element_text(size = 15, hjust = 0.5),
+    legend.title = element_blank(),
+    legend.text = element_text(size = 13),
+    axis.title.y = element_text(size = 13),
+    axis.title.x = element_text(size = 13),
+    axis.text.y = element_text(size = 13),
+    axis.text.x = element_text(size = 13)
   ) +
   labs(
     title = "Lower Back",
@@ -86,14 +106,24 @@ scatterplot_LR_res_hip <- mechanical_load_data %>%
   ) +
   scale_color_nejm() +
   scale_y_continuous(
-    limits = c(0, 60000),
+    limits = c(0, 30000),
     expand = c(0, 0),
-    breaks = seq(0, 60000, 10000)
[…]

### 2021-03-04 figures/figS3.R

@@ -31,14 +31,19 @@ BA_GRF_res_ankle <- cv_res_GRF_models$ankle %>%
   guides(shape = FALSE) +
   scale_color_nejm() +
   scale_y_continuous(
-    limits = c(-1500, 600),
+    limits = c(-600, 600),
     expand = c(0, 0),
-    breaks = seq(-1500, 600, 300)
+    breaks = seq(-600, 600, 300)
   ) +
   theme_light() +
   theme(
-    plot.title = element_text(hjust = 0.5),
-    legend.title = element_blank()
+    plot.title = element_text(size = 15, hjust = 0.5),
+    legend.title = element_blank(),
+    legend.text = element_text(size = 13),
+    axis.title.y = element_text(size = 13),
+    axis.title.x = element_text(size = 13),
+    axis.text.y = element_text(size = 13),
+    axis.text.x = element_text(size = 13)
   ) +
   labs(
     title = "Ankle",
@@ -58,8 +63,13 @@ BA_GRF_res_back <- cv_res_GRF_models$lower_back %>%
   ) +
   theme_light() +
   theme(
-    plot.title = element_text(hjust = 0.5),
-    legend.title = element_blank()
+    plot.title = element_text(size = 15, hjust = 0.5),
+    legend.title = element_blank(),
+    legend.text = element_text(size = 13),
+    axis.title.y = element_text(size = 13),
+    axis.title.x = element_text(size = 13),
+    axis.text.y = element_text(size = 13),
+    axis.text.x = element_text(size = 13)
   ) +
   labs(
     title = "Lower Back",
@@ -79,8 +89,13 @@ BA_GRF_res_hip <- cv_res_GRF_models$hip %>%
   ) +
   theme_light() +
   theme(
-    plot.title = element_text(hjust = 0.5),
-    legend.title = element_blank()
+    plot.title = element_text(size = 15, hjust = 0.5),
+    legend.title = element_blank(),
+    legend.text = element_text(size = 13),
+    axis.title.y = element_text(size = 13),
+    axis.title.x = element_text(size = 13),
+    axis.text.y = element_text(size = 13),
+    axis.text.x = element_text(size = 13)
   ) +
   labs(
     title = "Hip",
@@ -94,14 +109,19 @@ BA_GRF_ver_ankle <- cv_ver_GRF_models$ankle %>%
   guides(shape = FALSE) +
   scale_color_nejm() +
   scale_y_continuous(
-    limits = c(-900, 600),
[…]

### 2021-03-04 figures/figS4.R

@@ -55,14 +55,24 @@ BA_LR_res_back <- cv_res_LR_models$lower_back %>%
   scale_color_nejm() +
   scale_y_continuous(
     labels = scales::label_number(),
-    limits = c(-30000, 30000),
+    limits = c(-20000, 20000),
     expand = c(0, 0),
-    breaks = seq(-30000, 30000, 10000)
+    breaks = seq(-20000, 20000, 10000)
+  ) +
+  scale_x_continuous(
+    limits = c(2500, 27500),
+    expand = c(0, 0),
+    breaks = seq(5000, 25000, 5000)
   ) +
   theme_light() +
   theme(
-    plot.title = element_text(hjust = 0.5),
-    legend.title = element_blank()
+    plot.title = element_text(size = 15, hjust = 0.5),
+    legend.title = element_blank(),
+    legend.text = element_text(size = 13),
+    axis.title.y = element_text(size = 13),
+    axis.title.x = element_text(size = 13),
+    axis.text.y = element_text(size = 13),
+    axis.text.x = element_text(size = 13)
   ) +
   labs(
     title = "Lower Back",
@@ -77,14 +87,24 @@ BA_LR_res_hip <- cv_res_LR_models$hip %>%
   scale_color_nejm() +
   scale_y_continuous(
     labels = scales::label_number(),
-    limits = c(-30000, 30000),
+    limits = c(-20000, 20000),
     expand = c(0, 0),
-    breaks = seq(-30000, 30000, 10000)
+    breaks = seq(-20000, 20000, 10000)
+  ) +
+  scale_x_continuous(
+    limits = c(2500, 27500),
+    expand = c(0, 0),
+    breaks = seq(5000, 25000, 5000)
   ) +
   theme_light() +
   theme(
-    plot.title = element_text(hjust = 0.5),
-    legend.title = element_blank()
+    plot.title = element_text(size = 15, hjust = 0.5),
+    legend.title = element_blank(),
+    legend.text = element_text(size = 13),
+    axis.title.y = element_text(size = 13),
+    axis.title.x = element_text(size = 13),
+    axis.text.y = element_text(size = 13),
+    axis.text.x = element_text(size = 13)
   ) +
   labs(
     title = "Hip",
@@ -122,14 +142,24 @@ BA_LR_ver_back <- cv_ver_LR_models$lower_back %>%
   scale_color_nejm() +
   scale_y_continuous(
     labels = scales::label_number(),
[…]

### 2021-03-04 code/01_tidy_data.R
Change variable name from ATR to AR

@@ -90,10 +90,10 @@ mechanical_load_rates_dt_wide <- read_csv(
   select(
     id, trial = visita, filename = nome, acc_placement = local_acelerometro,
     speed = velocidade_km_h,
-    pRATR_gs = max_rates_atot_med_atot_g_s,
+    pRAR_gs = max_rates_atot_med_atot_g_s,
     pRLR_Ns = max_rates_ftot_med_ftot_n_s,
     pRLR_BWs = max_rates_ftot_med_ftot_bw,
-    pVATR_gs = max_rates_avt_med_avt_g_s,
+    pVAR_gs = max_rates_avt_med_avt_g_s,
     pVLR_Ns = max_rates_fvt_med_fvt_n_s,
     pVLR_BWs = max_rates_fvt_med_fvt_bw
   ) %>%
@@ -102,7 +102,7 @@ mechanical_load_rates_dt_wide <- read_csv(
 mechanical_load_rates_res <- mechanical_load_rates_dt_wide %>%
   select(
     id, trial, filename, acc_placement, speed,
-    pATR_gs = pRATR_gs,
+    pAR_gs = pRAR_gs,
     pLR_Ns = pRLR_Ns, pLR_BWs = pRLR_BWs
   ) %>%
   mutate(
@@ -113,7 +113,7 @@ mechanical_load_rates_res <- mechanical_load_rates_dt_wide %>%
 mechanical_load_rates_ver <- mechanical_load_rates_dt_wide %>%
   select(
     id, trial, filename, acc_placement, speed,
-    pATR_gs = pVATR_gs,
+    pAR_gs = pVAR_gs,
     pLR_Ns = pVLR_Ns, pLR_BWs = pVLR_BWs
   ) %>%
   mutate(

### 2021-03-04 code/02_explore.R

@@ -66,9 +66,9 @@ ACC_speed_plots <- mechanical_load_data %>%
   geom_boxplot() +
   facet_wrap(~ acc_placement)
 
-# Acceleration transient rate
-ATR_speed_plots <- mechanical_load_data %>%
-  ggplot(aes(x = speed, y = pATR_gs, fill = vector)) +
+# Acceleration rate
+AR_speed_plots <- mechanical_load_data %>%
+  ggplot(aes(x = speed, y = pAR_gs, fill = vector)) +
   geom_boxplot() +
   facet_wrap(~ acc_placement)
 
@@ -82,7 +82,7 @@ correlations <- map2(
   info$vectors, info$placement,
   ~ mechanical_load_data %>%
     filter(vector == .x & acc_placement == .y) %>%
-    select(pGRF_N, pLR_Ns, pACC_g, pATR_gs, body_mass)
+    select(pGRF_N, pLR_Ns, pACC_g, pAR_gs, body_mass)
 ) %>%
   set_names(paste(info$vectors, info$placement, sep = "_")) %>%
   map(correlate)
@@ -95,8 +95,8 @@ GRF_ACC_scatterplot <- plot_scatter(
 ) +
   facet_wrap(~ acc_placement)
 
-# LR x ATR
-LR_ATR_scatterplot <- plot_scatter(
-  mechanical_load_data, pATR_gs, pLR_Ns, color = BMI_cat
+# LR x AR
+LR_AR_scatterplot <- plot_scatter(
+  mechanical_load_data, pAR_gs, pLR_Ns, color = BMI_cat
 ) +
   facet_wrap(~ acc_placement)

### 2021-03-04 code/03_predict.R

@@ -70,7 +70,7 @@ plot_ver_GRF_models <- map(
 # Build LR models ---------------------------------------------------------
 
 LR_formula <- as.formula(
-  "pLR_Ns ~ pATR_gs + body_mass + pATR_gs:body_mass + (1 | subj)"
+  "pLR_Ns ~ pAR_gs + body_mass + pAR_gs:body_mass + (1 | subj)"
 )
 
 # As LR for the ankle placement is missing at the moment, this placement

### 2021-03-04 figures/fig2.R

@@ -54,11 +54,11 @@ scatterplot_LR_res_back <- mechanical_load_data %>%
   filter(vector == "resultant" & acc_placement == "lower_back") %>%
   ggplot() +
   geom_point(
-    aes(x = pATR_gs, y = pLR_Ns, color = BMI_cat, shape = activity),
+    aes(x = pAR_gs, y = pLR_Ns, color = BMI_cat, shape = activity),
     show_guide = FALSE
   ) +
   geom_smooth(
-    aes(x = pATR_gs, y = pLR_Ns, color = BMI_cat),
+    aes(x = pAR_gs, y = pLR_Ns, color = BMI_cat),
     method = "lm", se = FALSE
   ) +
   scale_color_nejm() +
@@ -84,7 +84,7 @@ scatterplot_LR_res_back <- mechanical_load_data %>%
   ) +
   labs(
     title = "Lower Back",
-    x = quote("pRATR" ~ (italic(g) %.% s^-1)),
+    x = quote("pRAR" ~ (italic(g) %.% s^-1)),
     y = quote("pRLR" ~ (N %.% s^-1))
   )
 

### 2021-03-04 figures/figS2.R

@@ -12,18 +12,18 @@ load(here("data", "mechanical_load_data.rda"))
 mechanical_load_data <- mechanical_load_data %>%
   mutate(activity = as.factor(ifelse(speed %in% 1:6, "walking", "running")))
 
-# LR x ATR plots ----------------------------------------------------------
+# LR x AR plots -----------------------------------------------------------
 
 # Resultant: Ankle
 scatterplot_LR_res_ankle <- mechanical_load_data %>%
   filter(vector == "resultant" & acc_placement == "ankle") %>%
   ggplot() +
   geom_point(
-    aes(x = pATR_gs, y = pLR_Ns, color = BMI_cat, shape = activity),
+    aes(x = pAR_gs, y = pLR_Ns, color = BMI_cat, shape = activity),
     show_guide = FALSE
   ) +
   geom_smooth(
-    aes(x = pATR_gs, y = pLR_Ns, color = BMI_cat),
+    aes(x = pAR_gs, y = pLR_Ns, color = BMI_cat),
     method = "lm", se = FALSE
   ) +
   scale_color_nejm() +
@@ -49,7 +49,7 @@ scatterplot_LR_res_ankle <- mechanical_load_data %>%
   ) +
   labs(
     title = "Ankle",
-    x = quote("pRATR" ~ (italic(g) %.% s^-1)),
+    x = quote("pRAR" ~ (italic(g) %.% s^-1)),
     y = quote("pRLR" ~ (N %.% s^-1))
   )
 
@@ -58,11 +58,11 @@ scatterplot_LR_res_back <- mechanical_load_data %>%
   filter(vector == "resultant" & acc_placement == "lower_back") %>%
   ggplot() +
   geom_point(
-    aes(x = pATR_gs, y = pLR_Ns, color = BMI_cat, shape = activity),
+    aes(x = pAR_gs, y = pLR_Ns, color = BMI_cat, shape = activity),
     show_guide = FALSE
   ) +
   geom_smooth(
-    aes(x = pATR_gs, y = pLR_Ns, color = BMI_cat),
+    aes(x = pAR_gs, y = pLR_Ns, color = BMI_cat),
     method = "lm", se = FALSE
   ) +
   scale_color_nejm() +
[…]

### 2021-03-08 figures/fig1.R
Adjust axes scales and export to png as well as tiff

@@ -58,6 +58,17 @@ fig1 <- mechanical_load_data %>%
 
 # Save plot ---------------------------------------------------------------
 
+agg_png(
+  here("figures", "fig1.png"),
+  width = 80,
+  height = 50,
+  units = "cm",
+  res = 100,
+  scaling = 2
+)
+plot(fig1)
+dev.off()
+
 agg_tiff(
   here("figures", "fig1.tiff"),
   width = 80,

### 2021-03-08 figures/fig2.R

@@ -48,10 +48,10 @@ scatterplot_GRF_res_hip <- mechanical_load_data %>%
   ) +
   labs(title = "Hip", x = quote("pRACC" ~ (italic(g))), y = "pRGRF (N)")
 
-# Back pRLR x pRAR plot ---------------------------------------------------
+# Hip pRLR x pRAR plot ----------------------------------------------------
 
-scatterplot_LR_res_back <- mechanical_load_data %>%
-  filter(vector == "resultant" & acc_placement == "lower_back") %>%
+scatterplot_LR_res_hip <- mechanical_load_data %>%
+  filter(vector == "resultant" & acc_placement == "hip") %>%
   ggplot() +
   geom_point(
     aes(x = pAR_gs, y = pLR_Ns, color = BMI_cat, shape = activity),
@@ -63,14 +63,14 @@ scatterplot_LR_res_back <- mechanical_load_data %>%
   ) +
   scale_color_nejm() +
   scale_y_continuous(
-    limits = c(0, 30000),
+    limits = c(0, 60000),
     expand = c(0, 0),
-    breaks = seq(0, 30000, 5000)
+    breaks = seq(0, 60000, 10000)
   ) +
   scale_x_continuous(
-    limits = c(0, 50),
+    limits = c(0, 175),
     expand = c(0, 0),
-    breaks = seq(5, 45, 5)
+    breaks = seq(25, 170, 25)
   ) +
   theme_light() +
   theme(
@@ -83,7 +83,7 @@ scatterplot_LR_res_back <- mechanical_load_data %>%
     axis.text.x = element_text(size = 13)
   ) +
   labs(
-    title = "Lower Back",
+    title = "Hip",
     x = quote("pRAR" ~ (italic(g) %.% s^-1)),
     y = quote("pRLR" ~ (N %.% s^-1))
   )
@@ -91,13 +91,25 @@ scatterplot_LR_res_back <- mechanical_load_data %>%
 # Combine and save plots --------------------------------------------------
 
 fig2 <- scatterplot_GRF_res_hip +
-  scatterplot_LR_res_back +
+  scatterplot_LR_res_hip +
   plot_annotation(tag_levels = "A") +
   plot_layout(guides = "collect") &
   theme(
     legend.position = "bottom",
     plot.tag = element_text(size = 16)
   )
+
+agg_png(
+  here("figures", "fig2.png"),
+  width = 80,
+  height = 25,
[…]

### 2021-03-08 figures/fig3.R

@@ -50,9 +50,9 @@ BA_GRF_res_hip <- cv_res_GRF_models$hip %>%
     y = "Actual - Predicted pRGRF (N)"
   )
 
-# Back pRLR Bland-Altman plot ---------------------------------------------
+# Hip pRLR Bland-Altman plot ----------------------------------------------
 
-BA_LR_res_back <- cv_res_LR_models$lower_back %>%
+BA_LR_res_hip <- cv_res_LR_models$hip %>%
   plot_bland_altman(color = BMI_cat, shape = activity) +
   guides(shape = FALSE) +
   scale_color_nejm() +
@@ -63,9 +63,9 @@ BA_LR_res_back <- cv_res_LR_models$lower_back %>%
     breaks = seq(-20000, 20000, 10000)
   ) +
   scale_x_continuous(
-    limits = c(2500, 27500),
+    limits = c(0, 50000),
     expand = c(0, 0),
-    breaks = seq(5000, 25000, 5000)
+    breaks = seq(5000, 50000, 10000)
   ) +
   theme_light() +
   theme(
@@ -78,7 +78,7 @@ BA_LR_res_back <- cv_res_LR_models$lower_back %>%
     axis.text.x = element_text(size = 13)
   ) +
   labs(
-    title = "Lower Back",
+    title = "Hip",
     x = quote("Mean of Actual and Predicted pRLR" ~ (N %.% s^-1)),
     y = quote("Actual - Predicted pRLR" ~ (N %.% s^-1))
   )
@@ -86,13 +86,25 @@ BA_LR_res_back <- cv_res_LR_models$lower_back %>%
 # Combine and save plots --------------------------------------------------
 
 fig3 <- BA_GRF_res_hip +
-  BA_LR_res_back +
+  BA_LR_res_hip +
   plot_annotation(tag_levels = "A") +
   plot_layout(guides = "collect") &
   theme(
     legend.position = "bottom",
     plot.tag = element_text(size = 16)
   )
+
+agg_png(
+  here("figures", "fig3.png"),
+  width = 80,
+  height = 25,
+  units = "cm",
+  res = 100,
+  scaling = 2
+)
+plot(fig3)
+dev.off()
+
 agg_tiff(
   here("figures", "fig3.tiff"),
   width = 80,

### 2021-03-08 figures/figS1.R

@@ -238,6 +238,18 @@ figS1 <- scatterplot_GRF_res_ankle +
     legend.position = "bottom",
     plot.tag = element_text(size = 16)
   )
+
+agg_png(
+  here("figures", "figS1.png"),
+  width = 120,
+  height = 50,
+  units = "cm",
+  res = 100,
+  scaling = 2
+)
+plot(figS1)
+dev.off()
+
 agg_tiff(
   here("figures", "figS1.tiff"),
   width = 120,

### 2021-03-08 figures/figS2.R

@@ -262,6 +262,18 @@ figS2 <- scatterplot_LR_res_ankle +
     legend.position = "bottom",
     plot.tag = element_text(size = 16)
   )
+
+agg_png(
+  here("figures", "figS2.png"),
+  width = 120,
+  height = 50,
+  units = "cm",
+  res = 100,
+  scaling = 2
+)
+plot(figS2)
+dev.off()
+
 agg_tiff(
   here("figures", "figS2.tiff"),
   width = 120,

### 2021-03-08 figures/figS3.R

@@ -195,6 +195,18 @@ figS3 <- BA_GRF_res_ankle +
     legend.position = "bottom",
     plot.tag = element_text(size = 16)
   )
+
+agg_png(
+  here("figures", "figS3.png"),
+  width = 120,
+  height = 50,
+  units = "cm",
+  res = 100,
+  scaling = 2
+)
+plot(figS3)
+dev.off()
+
 agg_tiff(
   here("figures", "figS3.tiff"),
   width = 120,

### 2021-03-08 figures/figS4.R

@@ -213,6 +213,18 @@ figS4 <- BA_LR_res_ankle +
     legend.position = "bottom",
     plot.tag = element_text(size = 16)
   )
+
+agg_png(
+  here("figures", "figS4.png"),
+  width = 120,
+  height = 50,
+  units = "cm",
+  res = 100,
+  scaling = 2
+)
+plot(figS4)
+dev.off()
+
 agg_tiff(
   here("figures", "figS4.tiff"),
   width = 120,

### 2021-03-08 manuscript/manuscript.Rmd
Remove section numbering and revise Methods section

@@
 output:
-  pdf_document:
+  bookdown::pdf_document2:
+    toc: false
     includes:
@@
 
-# Introduction
+\section*{Introduction}
 
@@
 
-# Methods
+\section*{Methods}
+
+\subsection*{Participants}
 
@@
 
-\medskip
+\subsection*{Protocol}
 
-Data collection took place at Porto Biomechanics Laboratory (LABIOMEP-UP), in an instrumented treadmill with built-in FP (AMTI Corporation, Watertown, USA).
+Data collection took place at Porto Biomechanics Laboratory (LABIOMEP-UP) in an instrumented treadmill with built-in FP (AMTI Corporation, Watertown, USA).
 Before starting the data collection, participants were given time to warm up on the treadmill.
@@
 Data from FP and accelerometers were collected at 1000 Hz and 100 Hz sampling frequency, respectively, and operated through the manufacturer supplied software (for the FP: Netforce, Version 3.5.1; AMTI Corporation, Watertown, USA; for the accelerometers: ActiLife version 6.13.3; Actigraph, Pensacola, USA).
-Raw GRF (expressed in N) and ACC (expressed in gravitational ACC units; 1\textit{g} =
-9.807 $\mathrm{m\cdot s^{-2}}$) from the x, y and z vectors were exported to plain text files.
+Raw GRF (expressed in N) and ACC (expressed in gravitational ACC units; 1\textit{g} = 9.807 $\mathrm{m\cdot s^{-2}}$) from the x, y and z vectors were exported to plain text files.
 
-\medskip
+\subsection*{Data processing}
 
@@
 Briefly, both GRF and ACC signals were filtered using a Butterworth fourth-order low-pass filter, with 20 Hz cut-off frequency, and their resultant vector calculated ($r_i = \sqrt{x_i^2 + y_i^2 + z_i^2}$).
-After that, peak ACC (pACC) values were defined as minimum height of twice the positive ACC average, during any given walking speed, and separated by 0.4 seconds minimum.
-Then, GRF and ACC signals were synchronized using the following sequence: i) signal adjustment through time set by the systems clock; ii) manual correction based on visual inspection for possible errors in time adjustment; iii) determination of pGRF as the highest values between 0.4 seconds before and after each pACC; iv) synchronization by the maximum cross-correlation coefficient of GRF and ACC peak values for both the resultant and its vertical component.
-The rates of change, LR for the GRF and ACC rate (AR) for the ACC, were computed through a second-order centered derivative from the beggining of the foot contact to the pGRF, and pLR and peak AR (pAR) were defined as the maximum value in the array.
+After that, peak ACC (pACC) values were defined as minimum height of twice the positive ACC average, during any given speed, and separated by 0.4 seconds minimum.
+GRF and ACC signals were adjusted through the time set by the systems clock, with manual correction based on visual inspection for possible errors, and then synchronized by the maximum cross-correlation coefficient.
+Peak GRF (pGRF) was determined as the highest value between 0.4 seconds before and after each pACC.
+The rates of change, LR for the GRF and ACC rate (AR) for the ACC, were computed through a second-order centered derivative from the beginning of the foot contact to the pGRF, and pLR and peak AR (pAR) were defined as the maximum value in the array.
 
@@
 
-\medskip
+\subsection*{Statistical analyses}
 
@@
 
-Our pGRF prediction equation was compared with a previously published reference equation [@Neugebauer2014], that was developed using similar methods.
+Finally, our pGRF prediction equation was compared with a previously published reference equation [@Neugebauer2014], that was developed using similar methods.
 This compaarison was performed in three ways using: i) the whole sample; ii) a subsample of walking trials; iii) a subsample of running trials.

### 2021-03-08 figures/fig2.R
Add legend for shape and legend title for both legends

@@ -10,7 +10,12 @@ library(ragg)
 
 load(here("data", "mechanical_load_data.rda"))
 mechanical_load_data <- mechanical_load_data %>%
-  mutate(activity = as.factor(ifelse(speed %in% 1:6, "walking", "running")))
+  mutate(
+    activity = fct_relevel(
+      as.factor(ifelse(speed %in% 1:6, "Walking", "Running")),
+      "Walking"
+    )
+  )
 
 # Hip pRGRF x pRACC plot --------------------------------------------------
 
@@ -18,8 +23,7 @@ scatterplot_GRF_res_hip <- mechanical_load_data %>%
   filter(vector == "resultant" & acc_placement == "hip") %>%
   ggplot() +
   geom_point(
-    aes(x = pACC_g, y = pGRF_N, color = BMI_cat, shape = activity),
-    show.legend = FALSE
+    aes(x = pACC_g, y = pGRF_N, color = BMI_cat, shape = activity)
   ) +
   geom_smooth(
     aes(x = pACC_g, y = pGRF_N, color = BMI_cat),
@@ -39,12 +43,16 @@ scatterplot_GRF_res_hip <- mechanical_load_data %>%
   theme_light() +
   theme(
     plot.title = element_text(size = 15, hjust = 0.5),
-    legend.title = element_blank(),
-    legend.text = element_text(size = 13),
+    legend.title = element_text(size = 13),
+    legend.text = element_text(size = 12),
     axis.title.y = element_text(size = 13),
     axis.title.x = element_text(size = 13),
     axis.text.y = element_text(size = 13),
     axis.text.x = element_text(size = 13)
+  ) +
+  guides(
+    color = guide_legend(title = "Body mass index category:"),
+    shape = guide_legend(title = "Locomotion type:")
   ) +
   labs(title = "Hip", x = quote("pRACC" ~ (italic(g))), y = "pRGRF (N)")
 
@@ -54,8 +62,7 @@ scatterplot_LR_res_hip <- mechanical_load_data %>%
   filter(vector == "resultant" & acc_placement == "hip") %>%
   ggplot() +
   geom_point(
-    aes(x = pAR_gs, y = pLR_Ns, color = BMI_cat, shape = activity),
-    show_guide = FALSE
+    aes(x = pAR_gs, y = pLR_Ns, color = BMI_cat, shape = activity)
   ) +
   geom_smooth(
     aes(x = pAR_gs, y = pLR_Ns, color = BMI_cat),
[…]

### 2021-03-08 figures/fig3.R

@@ -13,13 +13,21 @@ load(here("output", "loocv_data.rda"))
 cv_res_GRF_models <- map(
   cv_res_GRF_models,
   ~ mutate(
-    .x, activity = as.factor(ifelse(speed %in% 1:6, "walking", "running"))
+    .x,
+    activity = fct_relevel(
+      as.factor(ifelse(speed %in% 1:6, "Walking", "Running")),
+      "Walking"
+    )
   )
 )
 cv_res_LR_models <- map(
   cv_res_LR_models,
   ~ mutate(
-    .x, activity = as.factor(ifelse(speed %in% 1:6, "walking", "running"))
+    .x,
+    activity = fct_relevel(
+      as.factor(ifelse(speed %in% 1:6, "Walking", "Running")),
+      "Walking"
+    )
   )
 )
 
@@ -27,7 +35,6 @@ cv_res_LR_models <- map(
 
 BA_GRF_res_hip <- cv_res_GRF_models$hip %>%
   plot_bland_altman(color = BMI_cat, shape = activity) +
-  guides(shape = FALSE) +
   scale_color_nejm() +
   scale_y_continuous(
     limits = c(-600, 600),
@@ -37,12 +44,16 @@ BA_GRF_res_hip <- cv_res_GRF_models$hip %>%
   theme_light() +
   theme(
     plot.title = element_text(size = 15, hjust = 0.5),
-    legend.title = element_blank(),
-    legend.text = element_text(size = 13),
+    legend.title = element_text(size = 13),
+    legend.text = element_text(size = 12),
     axis.title.y = element_text(size = 13),
     axis.title.x = element_text(size = 13),
     axis.text.y = element_text(size = 13),
     axis.text.x = element_text(size = 13)
+  ) +
+  guides(
+    color = guide_legend(title = "Body mass index category:"),
+    shape = guide_legend(title = "Locomotion type:")
   ) +
   labs(
     title = "Hip",
@@ -54,7 +65,6 @@ BA_GRF_res_hip <- cv_res_GRF_models$hip %>%
 
 BA_LR_res_hip <- cv_res_LR_models$hip %>%
   plot_bland_altman(color = BMI_cat, shape = activity) +
-  guides(shape = FALSE) +
   scale_color_nejm() +
   scale_y_continuous(
     labels = scales::label_number(),
@@ -70,12 +80,16 @@ BA_LR_res_hip <- cv_res_LR_models$hip %>%
   theme_light() +
   theme(
     plot.title = element_text(size = 15, hjust = 0.5),
[…]

### 2021-03-08 figures/figS1.R

@@ -10,7 +10,12 @@ library(ragg)
 
 load(here("data", "mechanical_load_data.rda"))
 mechanical_load_data <- mechanical_load_data %>%
-  mutate(activity = as.factor(ifelse(speed %in% 1:6, "walking", "running")))
+  mutate(
+    activity = fct_relevel(
+      as.factor(ifelse(speed %in% 1:6, "Walking", "Running")),
+      "Walking"
+    )
+  )
 
 # GRF x ACC plots ---------------------------------------------------------
 
@@ -19,8 +24,7 @@ scatterplot_GRF_res_ankle <- mechanical_load_data %>%
   filter(vector == "resultant" & acc_placement == "ankle") %>%
   ggplot() +
   geom_point(
-    aes(x = pACC_g, y = pGRF_N, color = BMI_cat, shape = activity),
-    show.legend = FALSE
+    aes(x = pACC_g, y = pGRF_N, color = BMI_cat, shape = activity)
   ) +
   geom_smooth(
     aes(x = pACC_g, y = pGRF_N, color = BMI_cat),
@@ -40,12 +44,16 @@ scatterplot_GRF_res_ankle <- mechanical_load_data %>%
   theme_light() +
   theme(
     plot.title = element_text(size = 15, hjust = 0.5),
-    legend.title = element_blank(),
-    legend.text = element_text(size = 13),
+    legend.title = element_text(size = 13),
+    legend.text = element_text(size = 12),
     axis.title.y = element_text(size = 13),
     axis.title.x = element_text(size = 13),
     axis.text.y = element_text(size = 13),
     axis.text.x = element_text(size = 13)
+  ) +
+  guides(
+    color = guide_legend(title = "Body mass index category:"),
+    shape = guide_legend(title = "Locomotion type:")
   ) +
   labs(title = "Ankle", x = quote("pRACC" ~ (italic(g))), y = "pRGRF (N)")
 
@@ -54,8 +62,7 @@ scatterplot_GRF_res_back <- mechanical_load_data %>%
   filter(vector == "resultant" & acc_placement == "lower_back") %>%
   ggplot() +
   geom_point(
-    aes(x = pACC_g, y = pGRF_N, color = BMI_cat, shape = activity),
-    show.legend = FALSE
+    aes(x = pACC_g, y = pGRF_N, color = BMI_cat, shape = activity)
   ) +
   geom_smooth(
     aes(x = pACC_g, y = pGRF_N, color = BMI_cat),
[…]

### 2021-03-08 figures/figS2.R

@@ -10,7 +10,12 @@ library(ragg)
 
 load(here("data", "mechanical_load_data.rda"))
 mechanical_load_data <- mechanical_load_data %>%
-  mutate(activity = as.factor(ifelse(speed %in% 1:6, "walking", "running")))
+  mutate(
+    activity = fct_relevel(
+      as.factor(ifelse(speed %in% 1:6, "Walking", "Running")),
+      "Walking"
+    )
+  )
 
 # LR x AR plots -----------------------------------------------------------
 
@@ -19,8 +24,7 @@ scatterplot_LR_res_ankle <- mechanical_load_data %>%
   filter(vector == "resultant" & acc_placement == "ankle") %>%
   ggplot() +
   geom_point(
-    aes(x = pAR_gs, y = pLR_Ns, color = BMI_cat, shape = activity),
-    show_guide = FALSE
+    aes(x = pAR_gs, y = pLR_Ns, color = BMI_cat, shape = activity)
   ) +
   geom_smooth(
     aes(x = pAR_gs, y = pLR_Ns, color = BMI_cat),
@@ -40,12 +44,16 @@ scatterplot_LR_res_ankle <- mechanical_load_data %>%
   theme_light() +
   theme(
     plot.title = element_text(size = 15, hjust = 0.5),
-    legend.title = element_blank(),
-    legend.text = element_text(size = 13),
+    legend.title = element_text(size = 13),
+    legend.text = element_text(size = 12),
     axis.title.y = element_text(size = 13),
     axis.title.x = element_text(size = 13),
     axis.text.y = element_text(size = 13),
     axis.text.x = element_text(size = 13)
+  ) +
+  guides(
+    color = guide_legend(title = "Body mass index category:"),
+    shape = guide_legend(title = "Locomotion type:")
   ) +
   labs(
     title = "Ankle",
@@ -58,8 +66,7 @@ scatterplot_LR_res_back <- mechanical_load_data %>%
   filter(vector == "resultant" & acc_placement == "lower_back") %>%
   ggplot() +
   geom_point(
-    aes(x = pAR_gs, y = pLR_Ns, color = BMI_cat, shape = activity),
-    show_guide = FALSE
+    aes(x = pAR_gs, y = pLR_Ns, color = BMI_cat, shape = activity)
   ) +
   geom_smooth(
     aes(x = pAR_gs, y = pLR_Ns, color = BMI_cat),
[…]

### 2021-03-08 figures/figS3.R

@@ -13,13 +13,21 @@ load(here("output", "loocv_data.rda"))
 cv_res_GRF_models <- map(
   cv_res_GRF_models,
   ~ mutate(
-    .x, activity = as.factor(ifelse(speed %in% 1:6, "walking", "running"))
+    .x,
+    activity = fct_relevel(
+      as.factor(ifelse(speed %in% 1:6, "Walking", "Running")),
+      "Walking"
+    )
   )
 )
 cv_ver_GRF_models <- map(
   cv_ver_GRF_models,
   ~ mutate(
-    .x, activity = as.factor(ifelse(speed %in% 1:6, "walking", "running"))
+    .x,
+    activity = fct_relevel(
+      as.factor(ifelse(speed %in% 1:6, "Walking", "Running")),
+      "Walking"
+    )
   )
 )
 
@@ -28,7 +36,6 @@ cv_ver_GRF_models <- map(
 # Resultant: Ankle
 BA_GRF_res_ankle <- cv_res_GRF_models$ankle %>%
   plot_bland_altman(color = BMI_cat, shape = activity) +
-  guides(shape = FALSE) +
   scale_color_nejm() +
   scale_y_continuous(
     limits = c(-600, 600),
@@ -38,12 +45,16 @@ BA_GRF_res_ankle <- cv_res_GRF_models$ankle %>%
   theme_light() +
   theme(
     plot.title = element_text(size = 15, hjust = 0.5),
-    legend.title = element_blank(),
-    legend.text = element_text(size = 13),
+    legend.title = element_text(size = 13),
+    legend.text = element_text(size = 12),
     axis.title.y = element_text(size = 13),
     axis.title.x = element_text(size = 13),
     axis.text.y = element_text(size = 13),
     axis.text.x = element_text(size = 13)
+  ) +
+  guides(
+    color = guide_legend(title = "Body mass index category:"),
+    shape = guide_legend(title = "Locomotion type:")
   ) +
   labs(
     title = "Ankle",
@@ -54,7 +65,6 @@ BA_GRF_res_ankle <- cv_res_GRF_models$ankle %>%
 # Resultant: Lower back
 BA_GRF_res_back <- cv_res_GRF_models$lower_back %>%
   plot_bland_altman(color = BMI_cat, shape = activity) +
-  guides(shape = FALSE) +
   scale_color_nejm() +
   scale_y_continuous(
     limits = c(-600, 600),
[…]

### 2021-03-08 figures/figS4.R

@@ -13,13 +13,21 @@ load(here("output", "loocv_data.rda"))
 cv_res_LR_models <- map(
   cv_res_LR_models,
   ~ mutate(
-    .x, activity = as.factor(ifelse(speed %in% 1:6, "walking", "running"))
+    .x,
+    activity = fct_relevel(
+      as.factor(ifelse(speed %in% 1:6, "Walking", "Running")),
+      "Walking"
+    )
   )
 )
 cv_ver_LR_models <- map(
   cv_ver_LR_models,
   ~ mutate(
-    .x, activity = as.factor(ifelse(speed %in% 1:6, "walking", "running"))
+    .x,
+    activity = fct_relevel(
+      as.factor(ifelse(speed %in% 1:6, "Walking", "Running")),
+      "Walking"
+    )
   )
 )
 
@@ -51,7 +59,6 @@ BA_LR_res_ankle <- ggplot()
 # Resultant: Lower back
 BA_LR_res_back <- cv_res_LR_models$lower_back %>%
   plot_bland_altman(color = BMI_cat, shape = activity) +
-  guides(shape = FALSE) +
   scale_color_nejm() +
   scale_y_continuous(
     labels = scales::label_number(),
@@ -67,12 +74,16 @@ BA_LR_res_back <- cv_res_LR_models$lower_back %>%
   theme_light() +
   theme(
     plot.title = element_text(size = 15, hjust = 0.5),
-    legend.title = element_blank(),
-    legend.text = element_text(size = 13),
+    legend.title = element_text(size = 13),
+    legend.text = element_text(size = 12),
     axis.title.y = element_text(size = 13),
     axis.title.x = element_text(size = 13),
     axis.text.y = element_text(size = 13),
     axis.text.x = element_text(size = 13)
+  ) +
+  guides(
+    color = guide_legend(title = "Body mass index category:"),
+    shape = guide_legend(title = "Locomotion type:")
   ) +
   labs(
     title = "Lower Back",
@@ -83,7 +94,6 @@ BA_LR_res_back <- cv_res_LR_models$lower_back %>%
 # Resultant: Hip
 BA_LR_res_hip <- cv_res_LR_models$hip %>%
   plot_bland_altman(color = BMI_cat, shape = activity) +
-  guides(shape = FALSE) +
   scale_color_nejm() +
   scale_y_continuous(
     labels = scales::label_number(),
@@ -99,12 +109,16 @@ BA_LR_res_hip <- cv_res_LR_models$hip %>%
   theme_light() +
   theme(
[…]

### 2021-03-08 manuscript/manuscript.Rmd
Begin the Results section

@@
 
+\medskip
+
+\section*{Results}
+
+```{r bmi_class}
+BMI_avg <- sample_descriptives$BMI_mean
+BMI_class <- NA
+if (BMI_avg >= 18.5 & BMI_avg < 25) {
+  BMI_class <- "normal weight"
+} else if (BMI_avg >= 25 & BMI_avg < 30) {
+  BMI_class <- "overweight"
+} else if (BMI_avg >= 30 & BMI_avg < 35) {
+  BMI_class <- "obesity class I"
+} else if (BMI_avg >= 35 & BMI_avg < 40) {
+  BMI_class <- "obesity class II"
+} else if (BMI_avg >= 40) {
+  BMI_class <- "obesity class III"
+}
+```
+Our recruited sample presented an average BMI of `r sample_descriptives$BMI_mean` $\mathrm{kg\cdot m^{-2}}$, which is relatively high and classified as `r BMI_class`. 
+It would be expected that participants with higher BMI values, due to low motor ability and physical capacity, are able to perform the trials only at the lowest speeds.
+This is precisely what happened and, as can be observed in Figure 1, from 2 to 5 $\mathrm{km\cdot h^{-1}}$ the BMI median is around 30 $\mathrm{kg\cdot m^{-2}}$, decreasing to around 27 $\mathrm{kg\cdot m^{-2}}$ at 6 $\mathrm{km\cdot h^{-1}}$, which is the fastest walking speed, with a further reduction to less than 25 $\mathrm{kg\cdot m^{-2}}$ at the running speeds.
+Furthermore, only one participant with a BMI higher than 30 $\mathrm{kg\cdot m^{-2}}$ was able to run.
+This implies in a sample size reduction from `r filter(sample_size, acc_placement == "hip" & vector == "resultant" & speed == "2")$pGRF_N` at the slowest speed to `r filter(sample_size, acc_placement == "hip" & vector == "resultant" & speed == "14")$pGRF_N` at the fastest.
+
+\begin{figure}[ht]
+	\captionsetup{singlelinecheck = false, format = hang, justification = raggedright, font = footnotesize, labelsep = space}
+	\centering
+	\begin{measuredfigure}
+		\includegraphics[width=\textwidth]{fig1.png}
+		\caption{Distribution of participant’s body mass index per speed. Dots represent individual participants.}
+	\end{measuredfigure}
+\end{figure}
+
+\medskip
+
+As the main variables collected in this experiment are the peak magnitudes and rates of the GRF and ACC signals, there was a need to first establish their relationship.
+During locomotion, pACC increases linearly as pGRF increases.
+Such relation can be observed in the Figure 2, Panel A, for from hip-worn accelerometers and the resultant vector, and in Supplemental Figure S1 for all placements and the resultant vector and its vertical component.
+Moreover, for the same pACC values, the registered pGRF tended to be higher in subjects with higher BMI.
+It demonstrates that accelerometer data can adequately discriminate between different BMI categories.
+However, data from the pAR x pLR relationship show a greater dispersion, especially at higher values, and cannot discriminate so well the different BMI categories, as can be observed in Figure 2, Panel B, and Supplemental Figure S2.
+
+\begin{figure}[ht]
+	\captionsetup{singlelinecheck = false, format = hang, justification = raggedright, font = footnotesize, labelsep = space}
+	\centering
+	\begin{measuredfigure}
+		\includegraphics[width=\textwidth]{fig2.png}
+		\caption{Relationship between variables collected by force plates and accelerometers worn at hip, for the resultant vector.
+Linear trends for each body mass index category are also depicted.
+Panel A shows peak magnitudes (pRGRF and pRACC) while Panel B shows peak rates (pRLR and pRAR).
+Abbreviations: pRACC, peak resultant acceleration; pRAR, peak resultant acceleration rate; pRGRF, peak resultant ground reaction force; pRLR, peak resultant loading rate.}
+	\end{measuredfigure}
+\end{figure}
+
+\medskip
+
 \newpage
 
-# References
+\section*{References}
+
+<div id="refs"></div>
+
+\newpage
+
+\section*{Supplemental Material}
+
+\subsection*{Supplemental Figure S1}
+\begin{figure}[ht]
+	\captionsetup{singlelinecheck = false, format = hang, justification = raggedright, font = footnotesize, labelsep = space}
+	\centering
+	\begin{measuredfigure}
+		\includegraphics[width=\textwidth]{figS1.png}
+	\end{measuredfigure}
+\end{figure}
+\footnotesize Relationship between peak ground reaction force and acceleration magnitudes.
+Data from accelerometers worn at ankle (Panels A and D), lower back (Panels B and E) and hip (Panels C and F), for both the resultant vector (Panels A to C) and its vertical component (Panels D to F).
+Linear trends for each body mass index category are also depicted.
+Abbreviations: pRACC, peak resultant acceleration; pRGRF, peak resultant ground reaction force; pVACC, peak vertical acceleration; pVGRF, peak vertical ground reaction force.
+
+\medskip
+
+\subsection*{Supplemental Figure S2}
+\begin{figure}[ht]
+	\captionsetup{singlelinecheck = false, format = hang, justification = raggedright, font = footnotesize, labelsep = space}
+	\centering
+	\begin{measuredfigure}
+		\includegraphics[width=\textwidth]{figS2.png}
+	\end{measuredfigure}
+\end{figure}
+\footnotesize Relationship between peak loading and acceleration rates.
+Data from accelerometers worn at ankle (Panels A and D), lower back (Panels B and E) and hip (Panels C and F), for both the resultant vector (Panels A to C) and its vertical component (Panels D to F).
+Linear trends for each body mass index category are also depicted.
+Abbreviations: pRAR, peak resultant acceleration rate; pRLR, peak resultant loading rate; pVAR, peak vertical acceleration rate; pVLR, peak vertical loading rate.

### 2021-03-08 figures/fig1.R
Remove BMI abbreviation from y-axis legend

@@ -53,7 +53,7 @@ fig1 <- mechanical_load_data %>%
   ) +
   labs(
     x = "",
-    y = quote("BMI" ~ (kg %.% m^-2))
+    y = quote("Body mass index" ~ (kg %.% m^-2))
   )
 
 # Save plot ---------------------------------------------------------------

### 2021-03-09 code/funs.R
Change the round digits

@@ -15,9 +15,14 @@ build_formula_table <- function(model, cv, outcome, vector) {
     cv, lvmisc::accuracy, na.rm = TRUE
   )
   R2 <- unname(purrr::map_dbl(model_accuracy, "R2_cond"))
-  MAE <- unname(purrr::map_dbl(model_accuracy, "MAE"))
-  MAPE <- unname(purrr::map_dbl(model_accuracy, "MAPE"))
-  RMSE <- unname(purrr::map_dbl(model_accuracy, "RMSE"))
+  MAE <- broman::myround(unname(purrr::map_dbl(model_accuracy, "MAE")), 1)
+  MAPE <- paste0(
+    broman::myround(
+      unname(purrr::map_dbl(model_accuracy, "MAPE")) * 100, 1
+    ),
+    "\\%"
+  )
+  RMSE <- broman::myround(unname(purrr::map_dbl(model_accuracy, "RMSE")), 1)
 
   tibble::tibble(
     "Accelerometer placement" = c("Ankle", "Lower back", "Hip"),
@@ -40,16 +45,16 @@ build_formula_table <- function(model, cv, outcome, vector) {
 # Returns:
 #    A character vector with the model equation.
 get_equation <- function(model, outcome, vector) {
-  model_coefs <- as.character(round(coef(summary(model))[, 1], 3))
+  model_coefs <- coef(summary(model))[, 1]
   model_coefs[1] <- ifelse(
     model_coefs[1] > 0,
-    model_coefs[1],
-    paste0("- ", abs(as.numeric(model_coefs[1])))
+    broman::myround(model_coefs[1], 3),
+    paste0("- ", broman::myround(abs(as.numeric(model_coefs[1])), 3))
   )
   model_coefs[-1] <- ifelse(
-    as.numeric(model_coefs[-1]) > 0,
-    paste0(" + ", abs(as.numeric(model_coefs[-1]))),
-    paste0(" - ", abs(as.numeric(model_coefs[-1])))
+    model_coefs[-1] > 0,
+    paste0(" + ", broman::myround(abs(as.numeric(model_coefs[-1])), 3)),
+    paste0(" - ", broman::myround(abs(as.numeric(model_coefs[-1])), 3))
   )
 
   if (stringr::str_detect(outcome, "GRF") & vector == "resultant") {

### 2021-03-09 tables/tab1.R
Add a caption and footnote

@@ -20,18 +20,16 @@ ver_GRF_tb <- build_formula_table(
   ver_GRF_models, cv_ver_GRF_models, "pGRF", "vertical"
 )
 
-tab1 <- rbind(res_GRF_tb, ver_GRF_tb) %>%
-  mutate(
-    MAPE = paste0(myround(MAPE * 100, 2), "\\%")
-  )
+tab1 <- rbind(res_GRF_tb, ver_GRF_tb)
 # Reorder table rows
 tab1 <- tab1[c(1, 4, 2, 5, 3, 6), ]
 # Remove placemente name duplicates
 tab1[c(2, 4, 6), 1] <- ""
 
-
-tab1 <- kbl(
+tab1 <- knitr::kable(
   tab1, booktabs = TRUE, escape = FALSE, linesep = "",
+  label = "none",
+  caption = "Regession equations, $R^2$ and accuracy indices",
   col.names = c(
     linebreak("Accelerometer\nplacement"),
     "Regression equations",
@@ -42,6 +40,11 @@ tab1 <- kbl(
   ),
   align = c(rep("l", 2), rep("r", 4))
 ) %>%
+  footnote(
+    general = "Abbreviations: MAE, mean absolute error; MAPE, mean absolute percent error; pRACC, peak resultant acceleration; pRAR, peak resultant acceleration rate; pRGRF, peak resultant ground reaction force; pRLR, peak resultant loading rate; pVACC, peak vertical acceleration; pVAR, peak vertical acceleration rate; pVGRF, peak vertical ground reaction force; pVLR peak vertical loading rate; RMSE, root mean square error",
+    general_title = "",
+    threeparttable = TRUE
+  ) %>%
   kable_styling(latex_options = "scale_down") %>%
   pack_rows(
     "Peak ground reaction force prediction", 1, 6,

### 2021-03-10 tables/tab2.R
Change rounding digits as add caption and footnote

@@ -3,6 +3,7 @@
 library(here)
 library(tidyverse)
 library(kableExtra)
+library(broman)
 source(here("code", "funs.R"))
 
 # Load and prepare data ---------------------------------------------------
@@ -45,12 +46,26 @@ tab2 <- data.frame(
     neug_run_accuracy$RMSE
   )
 ) %>%
+  mutate(
+    across(starts_with("MAPE"), as.numeric),
+    across(starts_with("MAPE"), ~ .x * 100),
+    across(where(is.numeric), myround),
+    across(starts_with("MAPE"), ~ paste0(.x, "\\%"))
+  ) %>%
   kbl(
+    booktabs = TRUE, escape = FALSE,
+    label = "none2",
+    caption = "Accuracy indices of ours and Neugebauer’s equation to predict peak vertical ground reaction force with data from hip-worn accelerometers",
     col.names = c(
       "Prediction",
       "MAE", "MAPE", "RMSE",
       "MAE", "MAPE", "RMSE"
-    ),
-    booktabs = TRUE
+    )
   ) %>%
+  footnote(
+    general = "Abbreviations: MAE, mean absolute error; MAPE, mean absolute percent error; RMSE, root mean square error",
+    general_title = "",
+    threeparttable = TRUE
+  ) %>%
+  kable_styling(position = "center") %>%
   add_header_above(c("", "Our equation" = 3, "Neugebauer equation" = 3))

### 2021-03-10 manuscript/manuscript.Rmd
Finish Results section

@@
 library(here)
+library(broman)
+library(kableExtra)
 source(here("code", "02_explore.R"))
@@
 All statistical analyses were registered in an open platform, where R code utilized and detailed information can be assessed (LINK).
+The statistically significant value was set as $\alpha$ = 0.05.
 
@@
 Bias was computed as the mean of these differences and the limits of agreement were obtained using \pm 1.96 standard deviation of the mean between actual and predicted values [@Bland1986].
+In order to test whether there was a consistent over or underestimation of the acual values, one-sample \textit{t} tests were performed to assess if the difference between actual and predicted values are different than zero.
 Linear regressions were applied to identify if the magnitude of the mean between actual and predicted values influenced the magnitude of their difference.
@@
 
+```{r GRF_accuracy}
+load(here("output", "prediction_models.rda"))
+res_GRF_accuracy <- purrr::map(
+  accuracy_res_GRF_models,
+  ~ dplyr::mutate(.x, MAPE = as.numeric(MAPE))
+) %>%
+  purrr::map_dfr(rbind) %>%
+  dplyr::mutate(
+    placement = c("ankle", "lower back", "hip"),
+    vector = "resultant"
+  )
+ver_GRF_accuracy <- purrr::map(
+  accuracy_ver_GRF_models,
+  ~ dplyr::mutate(.x, MAPE = as.numeric(MAPE))
+) %>%
+  purrr::map_dfr(rbind) %>%
+  dplyr::mutate(
+    placement = c("ankle", "lower back", "hip"),
+    vector = "vertical"
+  )
+GRF_accuracy <- rbind(res_GRF_accuracy, ver_GRF_accuracy)
+hip_GRF_accuracy <- GRF_accuracy %>%
+  filter(placement == "hip" & vector == "resultant")
+```
+
+Regression equations were developed to predict the mechanical loading variables based on accelerometry data.
+In the pGRF prediction equations, body mass and pACC were significant predictors and included in the models.
+Table 1 shows the regression coefficients, $\mathrm{R^2}$ and accuracy indices from models derived from accelerometers placed at ankle, lower back and hip, and from the resultant vector and its vertical component.
+In all pGRF prediction models, $\mathrm{R^2}$ ranged from `r min(GRF_accuracy$R2_cond)` to `r max(GRF_accuracy$R2_cond)`, which shows that the predictor variables are able to explain in a great extent the outcome.
+Also, MAE ranged between `r myround(min(GRF_accuracy$MAE), 1)` N and `r myround(max(GRF_accuracy$MAE), 1)` N, and the MAPE from `r myround(min(GRF_accuracy$MAPE) * 100, 1)`\% to `r myround(max(GRF_accuracy$MAPE) * 100, 1)`\%, which is a smaller error than previously published equations [@Neugebauer2014; @Fortune2014], and similar to the error from equations previously published by out group, but which included walking only [@Veras2020].
+RMSE were also computed and shown to range between `r myround(min(GRF_accuracy$RMSE), 1)` N and `r myround(max(GRF_accuracy$RMSE), 1)` N.
+Of all the pGRF predictions equations developed, the one based on hip-worn accelerometers and the resultant vector has shown the best performance, with an $\mathrm{R^2}$ of `r hip_GRF_accuracy$R2_cond`, and the lowest MAE (`r myround(hip_GRF_accuracy$MAE, 1)` N), MAPE (`r myround(hip_GRF_accuracy$MAPE * 100, 1)`\%) and RMSE (`r myround(hip_GRF_accuracy$RMSE, 1)` N).
+
+\medskip
+
+```{r LR_accuracy}
+load(here("output", "prediction_models.rda"))
+res_LR_accuracy <- purrr::map(
+  accuracy_res_LR_models,
+  ~ dplyr::mutate(.x, MAPE = as.numeric(MAPE))
+) %>%
+  purrr::map_dfr(rbind) %>%
+  dplyr::mutate(
+    placement = c("lower back", "hip"),
+    vector = "resultant"
+  )
+ver_LR_accuracy <- purrr::map(
+  accuracy_ver_LR_models,
+  ~ dplyr::mutate(.x, MAPE = as.numeric(MAPE))
+) %>%
+  purrr::map_dfr(rbind) %>%
+  dplyr::mutate(
+    placement = c("lower back", "hip"),
+    vector = "vertical"
+  )
+LR_accuracy <- rbind(res_LR_accuracy, ver_LR_accuracy)
+back_LR_accuracy <- LR_accuracy %>%
+  filter(placement == "lower back" & vector == "resultant")
+```
+
+Moreover, pLR prediction equations were developed, and these included as predictors the pAR and again body mass.
+Regression coefficients, $\mathrm{R^2}$ and accuracy indices for these equations are also shown in Table 1.
+For these equations, the lowest $\mathrm{R^2}$ found was `r min(LR_accuracy$R2_cond)` and the highest was `r max(LR_accuracy$R2_cond)`, reflecting a moderate prediction ability.
+MAE, MAPE and RMSE values ranged from `r myround(min(LR_accuracy$MAE), 1)` N to `r myround(max(LR_accuracy$MAE), 1)` N, from `r myround(min(LR_accuracy$MAPE) * 100, 1)`\% to `r myround(max(LR_accuracy$MAPE) * 100, 1)`\%, and from `r myround(min(LR_accuracy$RMSE), 1)` N to `r myround(max(LR_accuracy$RMSE), 1)` N, respectively.
+The results found for the pLR prediction equations were all considerably poorer than those for the pGRF equations.
+The accuracy indices values were also somewhat worse than previously published pLR prediction equations [@Veras2020; @Fortune2014], although equations from both studies were only validate for walking.
+The best-performing pLR prediction equation was the one for accelerometers worn at lower back and the resultant vector, with an $\mathrm{R^2}$ of `r back_LR_accuracy$R2_cond` and a MAE, MAPE and RMSE of `r myround(back_LR_accuracy$MAE, 1)` N, `r myround(back_LR_accuracy$MAPE * 100, 1)`\% and `r myround(back_LR_accuracy$RMSE, 1)` N, respectively.
+
+\medskip
+
+```{r accuracy_table, results = "show"}
+source(here("tables", "tab1.R"))
+tab1
+```
+
+```{r ba_regression}
+source(here("code", "04_sub_analyses.R"))
+```
+
+Afterwards, Bland-Altman plots were built to visualize the agreement between the actual and predicted pGRF and pLR, for both the resultant vector and its vertical component and all accelerometer placements.
+A good agreement in a Bland-Altman plot is observed when values tend to cluster around zero, showing no consistent over or underestimation of the actual values.
+Also, the majority of data points should lie within the limits of agreement, which are determined by \pm 1.96 standard deviation of the bias (average difference between actual and predicted values).
+Bland-Altman plots for the pRGRF and pRLR from hip-worn accelerometers can be seen in Figure 3, and for all accelerometers placements and both resultant and vertical vectors in the Supplemental Figures S3 and S4.
+Values for all pGRF and pLR prediction equations developed tended to assemble around zero and between the limits of agreement.
+For the pGRF equations, a tendency to higher dispersion can be observed in the higher magnitudes, which correspond to the running trials, and especially in values derived from ankle-worn accelerometers (Supplemental Figure S3, Panels A and D).
+Prediction equations for pLR showed a more accentuated dispersion, and again in the higher magnitudes.
+None of the regression models presented a bias different than zero ($p > 0.05$), demonstrating that these models do not consistently over or underestimate the pGRF and pLR values.
+Furthermore, linear regressions were applied to identify proportional bias, that is, to test whether pGRF and pLR prediction errors were systematically influenced by their magnitude.
+The pGRF prediction equation based on ankle-worn accelerometers data for the resultant vector showed a significant proportional bias ($p = `r myround(coefficients(ba_regression_res_GRF$ankle)[2, 4], 2)`$), although with a low magnitude ($\mathrm{R^2} = `r myround(ba_regression_res_GRF$ankle["adj.r.squared"], 2)`$).
+
+\begin{figure}[ht]
+	\captionsetup{singlelinecheck = false, format = hang, justification = raggedright, font = footnotesize, labelsep = space}
+	\centering
+	\begin{measuredfigure}
+		\includegraphics[width=\textwidth]{fig3.png}
+		\caption{Bland-Altman plots showing agreement between actual and predicted pRGRF (Panel A) and pRLR (Panel B) for hip-worn accelerometers.
+Continuous thick lines show bias (average of the differences between actual and predicted values) while dashed lines show the limits of agreement (bias \pm 1.96 standard deviation).
+Abbreviations: pRGRF, peak resultant ground reaction force; pRLR, peak resultant loading rate.}
+	\end{measuredfigure}
+\end{figure}
+
+\medskip
+
+Lastly, we compared the performance of our equations with a reference equation previously published.
+As the only comparable equation found was the one to predict pVGRF using hip-worn accelerometer data developed by Neugebauer et al [@Neugebauer2014], we applied it to our data and computed the MAE, MAPE and RMSE.
+Then, these indices were compared to the indices for our equation (Table 1).
+Apart from the overall equations performance, it was done a sub-analysis separating the walking and running trials.
+The results of these comparisons can be observed in the Table 2.
+Accuracy indices of our equations were all remarkably better compared to the reference equation, with a MAPE approximately 2 to 3 times lower.
+Furthermore, our equation performance was constant between walking and running, indicating that it can be used for both activities without loss of accuracy.
+
+```{r comparison_table, results = "show"}
+source(here("tables", "tab2.R"))
+tab2
+```
+
 \newpage
@@
 Abbreviations: pRAR, peak resultant acceleration rate; pRLR, peak resultant loading rate; pVAR, peak vertical acceleration rate; pVLR, peak vertical loading rate.
+
+\newpage
+
+\subsection*{Supplemental Figure S3}
+\begin{figure}[ht]
+	\captionsetup{singlelinecheck = false, format = hang, justification = raggedright, font = footnotesize, labelsep = space}
+	\centering
+	\begin{measuredfigure}
+		\includegraphics[width=\textwidth]{figS3.png}
+	\end{measuredfigure}
+\end{figure}
+\footnotesize Bland-Altman plots showing agreement between actual and predicted pGRF for accelerometers worn at ankle (Panels A and D), lower back (Panels B and E) and hip (Panels C and F), for both the resultant vector (Panels A to C) and its vertical component (Panels D to F).
+Continuous thick lines show bias (average of the differences between actual and predicted values) while dashed lines show the limits of agreement (bias \pm 1.96 standard deviation).
+Abbreviations: pGRF, peak ground reaction force; pRGRF, peak resultant ground reaction force; pVGRF, peak vertical ground reaction force.
+
+\subsection*{Supplemental Figure S4}
+\begin{figure}[ht]
+	\captionsetup{singlelinecheck = false, format = hang, justification = raggedright, font = footnotesize, labelsep = space}
+	\centering
+	\begin{measuredfigure}
+		\includegraphics[width=\textwidth]{figS4.png}
+	\end{measuredfigure}
+\end{figure}
+\footnotesize Bland-Altman plots showing agreement between actual and predicted pLR for accelerometers worn at ankle (Panels A and D), lower back (Panels B and E) and hip (Panels C and F), for both the resultant vector (Panels A to C) and its vertical component (Panels D to F).
+Continuous thick lines show bias (average of the differences between actual and predicted values) while dashed lines show the limits of agreement (bias \pm 1.96 standard deviation).
+Abbreviations: pLR, peak loading rate; pRLR, peak resultant loading rate; pVLR, peak vertical loading rate.

### 2021-03-10 manuscript/manuscript.Rmd
Configure the creation of a docx for the manuscript

@@
     latex_engine: lualatex
+  word_document:
+    reference_docx: style_reference.docx
+always_allow_html: true
 bibliography: bibliography.bib
@@
 
-\section*{Introduction}
+# Introduction {.unnumbered}
 
@@
 
-\section*{Methods}
+# Methods {.unnumbered}
 
-\subsection*{Participants}
+## Participants {.unnumbered}
 
@@
 
-\subsection*{Protocol}
+## Protocol {.unnumbered}
 
@@
 
-\subsection*{Data processing}
+## Data processing {.unnumbered}
 
@@
 
-\subsection*{Statistical analyses}
+## Statistical analyses {.unnumbered}
 
@@
 Final models were chosen according to -2 log-likelihood statistics [@Field2012].
-Traditional coefficient of determination (R\textsuperscript{2}) was represented by conditional R\textsuperscript{2}, that estimates the variance explained by the whole model [@Nakagawa2013].
+Traditional coefficient of determination ($\mathrm{R^2}$) was represented by conditional $\mathrm{R^2}$, that estimates the variance explained by the whole model [@Nakagawa2013].
 
@@
 
-\section*{Results}
+# Results {.unnumbered}
 
@@
 
-\section*{References}
+# References {.unnumbered}
 
@@
 
-\section*{Supplemental Material}
+# Supplemental Material {.unnumbered}
 
-\subsection*{Supplemental Figure S1}
+## Supplemental Figure S1 {.unnumbered}
 \begin{figure}[ht]
@@
 
-\subsection*{Supplemental Figure S2}
+## Supplemental Figure S2 {.unnumbered}
 \begin{figure}[ht]
@@
 
-\subsection*{Supplemental Figure S3}
+## Supplemental Figure S3 {.unnumbered}
 \begin{figure}[ht]
@@
 
-\subsection*{Supplemental Figure S4}
+## Supplemental Figure S4 {.unnumbered}
 \begin{figure}[ht]

### 2021-03-11 manuscript/manuscript.Rmd
Remove pRGRF, pVGRF, pRACC, pVACC, pRLR, pVLR, pRAR, pVAR abbreviations

@@
 
-Regression equations to predict pRGFR, pVGRF, pRLR and pVLR were developed by linear mixed models.
+Regression equations to predict pGFR and pLR for each accelerometer placement (ankle, lower back and hip) and vector (resultant and vertical) were developed by linear mixed models.
 Body mass (kg) was tested as a predictor in all models.
-In addition, in the pRGRF and pVGRF models, pRACC and peak vertical ACC (pVACC; \textit{g}), respectively, were tested as predictors, while in the pRLR and pVLR models, peak resultant ATR (pRATR; $g\phantom{ }\mathrm{\phantom{ }\cdot s^{-1}}$) and peak vertical ATR (pVATR; $g\phantom{ }\mathrm{\phantom{ }\cdot s^{-1}}$), respectively, were tested as predictors.
+In addition, in the pGRF models pACC was tested as predictor, while in the pLR models pAR was tested as predictor.
 All predictors were tested as fixed effects and were shown to be significant.
@@
 
-Model was validated using the leave-one-out cross-validation approach [@Staudenmayer2012].
+Models were validated using the leave-one-out cross-validation approach [@Staudenmayer2012].
 For leave-one-out cross-validation each participant’s data was separated into a testing dataset (one participant at a time) with the remaining data in the training dataset.
@@
 ```
+
 Our recruited sample presented an average BMI of `r sample_descriptives$BMI_mean` $\mathrm{kg\cdot m^{-2}}$, which is relatively high and classified as `r BMI_class`. 
@@
 Linear trends for each body mass index category are also depicted.
-Panel A shows peak magnitudes (pRGRF and pRACC) while Panel B shows peak rates (pRLR and pRAR).
-Abbreviations: pRACC, peak resultant acceleration; pRAR, peak resultant acceleration rate; pRGRF, peak resultant ground reaction force; pRLR, peak resultant loading rate.}
+Panel A shows peak magnitudes (pGRF and pACC) while Panel B shows peak rates (pLR and pAR).
+Abbreviations: pACC, peak acceleration; pAR, peak acceleration rate; pGRF, peak ground reaction force; pLR, peak loading rate.}
 	\end{measuredfigure}
@@
 Also, the majority of data points should lie within the limits of agreement, which are determined by \pm 1.96 standard deviation of the bias (average difference between actual and predicted values).
-Bland-Altman plots for the pRGRF and pRLR from hip-worn accelerometers can be seen in Figure 3, and for all accelerometers placements and both resultant and vertical vectors in the Supplemental Figures S3 and S4.
+Bland-Altman plots for the pGRF and pLR resultant vector from hip-worn accelerometers can be seen in Figure 3, and for all accelerometers placements and both resultant and vertical vectors in the Supplemental Figures S3 and S4.
 Values for all pGRF and pLR prediction equations developed tended to assemble around zero and between the limits of agreement.
@@
 		\includegraphics[width=\textwidth]{fig3.png}
-		\caption{Bland-Altman plots showing agreement between actual and predicted pRGRF (Panel A) and pRLR (Panel B) for hip-worn accelerometers.
+		\caption{Bland-Altman plots showing agreement between actual and predicted peak ground reaction force (Panel A) and loading rate (Panel B) for the resultant vector from hip-worn accelerometers.
 Continuous thick lines show bias (average of the differences between actual and predicted values) while dashed lines show the limits of agreement (bias \pm 1.96 standard deviation).
-Abbreviations: pRGRF, peak resultant ground reaction force; pRLR, peak resultant loading rate.}
+Abbreviations: pGRF, peak ground reaction force; pLR, peak loading rate.}
 	\end{measuredfigure}
@@
 Lastly, we compared the performance of our equations with a reference equation previously published.
-As the only comparable equation found was the one to predict pVGRF using hip-worn accelerometer data developed by Neugebauer et al [@Neugebauer2014], we applied it to our data and computed the MAE, MAPE and RMSE.
+As the only comparable equation found was the one to predict pGRF vertical vector using hip-worn accelerometer data developed by Neugebauer et al [@Neugebauer2014], we applied it to our data and computed the MAE, MAPE and RMSE.
 Then, these indices were compared to the indices for our equation (Table 1).
@@
 Linear trends for each body mass index category are also depicted.
-Abbreviations: pRACC, peak resultant acceleration; pRGRF, peak resultant ground reaction force; pVACC, peak vertical acceleration; pVGRF, peak vertical ground reaction force.
+Abbreviations: pACC, peak acceleration; pGRF, peak ground reaction force.
 
@@
 Linear trends for each body mass index category are also depicted.
-Abbreviations: pRAR, peak resultant acceleration rate; pRLR, peak resultant loading rate; pVAR, peak vertical acceleration rate; pVLR, peak vertical loading rate.
+Abbreviations: pAR, peak acceleration rate; pLR, peak loading rate.
 
@@
 \end{figure}
-\footnotesize Bland-Altman plots showing agreement between actual and predicted pGRF for accelerometers worn at ankle (Panels A and D), lower back (Panels B and E) and hip (Panels C and F), for both the resultant vector (Panels A to C) and its vertical component (Panels D to F).
+\footnotesize Bland-Altman plots showing agreement between actual and predicted peak ground reaction force for accelerometers worn at ankle (Panels A and D), lower back (Panels B and E) and hip (Panels C and F), for both the resultant vector (Panels A to C) and its vertical component (Panels D to F).
 Continuous thick lines show bias (average of the differences between actual and predicted values) while dashed lines show the limits of agreement (bias \pm 1.96 standard deviation).
-Abbreviations: pGRF, peak ground reaction force; pRGRF, peak resultant ground reaction force; pVGRF, peak vertical ground reaction force.
+Abbreviations: pGRF, peak ground reaction force.
 
@@
 \end{figure}
-\footnotesize Bland-Altman plots showing agreement between actual and predicted pLR for accelerometers worn at ankle (Panels A and D), lower back (Panels B and E) and hip (Panels C and F), for both the resultant vector (Panels A to C) and its vertical component (Panels D to F).
+\footnotesize Bland-Altman plots showing agreement between actual and predicted peak loading rate for accelerometers worn at ankle (Panels A and D), lower back (Panels B and E) and hip (Panels C and F), for both the resultant vector (Panels A to C) and its vertical component (Panels D to F).
 Continuous thick lines show bias (average of the differences between actual and predicted values) while dashed lines show the limits of agreement (bias \pm 1.96 standard deviation).
-Abbreviations: pLR, peak loading rate; pRLR, peak resultant loading rate; pVLR, peak vertical loading rate.
+Abbreviations: pLR, peak loading rate.

### 2021-03-11 code/funs.R

@@ -25,9 +25,10 @@ build_formula_table <- function(model, cv, outcome, vector) {
   RMSE <- broman::myround(unname(purrr::map_dbl(model_accuracy, "RMSE")), 1)
 
   tibble::tibble(
+    "Vector" = stringr::str_to_sentence(vector),
     "Accelerometer placement" = c("Ankle", "Lower back", "Hip"),
     "Regression equations" = unname(
-      purrr::map_chr(model, get_equation, outcome, vector)
+      purrr::map_chr(model, get_equation, outcome)
     ),
     "R2" = R2, "MAE" = MAE, "MAPE" = MAPE, "RMSE" = RMSE
   )
@@ -44,7 +45,7 @@ build_formula_table <- function(model, cv, outcome, vector) {
 #
 # Returns:
 #    A character vector with the model equation.
-get_equation <- function(model, outcome, vector) {
+get_equation <- function(model, outcome) {
   model_coefs <- coef(summary(model))[, 1]
   model_coefs[1] <- ifelse(
     model_coefs[1] > 0,
@@ -57,19 +58,12 @@ get_equation <- function(model, outcome, vector) {
     paste0(" - ", broman::myround(abs(as.numeric(model_coefs[-1])), 3))
   )
 
-  if (stringr::str_detect(outcome, "GRF") & vector == "resultant") {
-    model_outcome <- "pRGRF (N)"
-    model_acceleration <- "pRACC"
-  } else if (stringr::str_detect(outcome, "GRF") & vector == "vertical") {
-    model_outcome <- "pVGRF (N)"
-    model_acceleration <- "pVACC"
-  } else if (stringr::str_detect(outcome, "LR") & vector == "resultant") {
-    model_outcome <- "pVLR (Ns)"
-    model_acceleration <- "pRATR"
-  } else if (stringr::str_detect(outcome, "LR") & vector == "resultant") {
-  } else if (stringr::str_detect(outcome, "LR") & vector == "vertical") {
-    model_outcome <- "pVLR (Ns)"
-    model_acceleration <- "pVATR"
+  if (stringr::str_detect(outcome, "GRF")) {
+    model_outcome <- "pGRF (N)"
+    model_acceleration <- "pACC"
+  } else if (stringr::str_detect(outcome, "LR")) {
+    model_outcome <- "pLR (Ns)"
+    model_acceleration <- "pAR"
   }
 
   regression_equation <- glue::glue(

### 2021-03-11 figures/fig2.R

@@ -17,7 +17,7 @@ mechanical_load_data <- mechanical_load_data %>%
     )
   )
 
-# Hip pRGRF x pRACC plot --------------------------------------------------
+# Hip pGRF x pACC plot --------------------------------------------------
 
 scatterplot_GRF_res_hip <- mechanical_load_data %>%
   filter(vector == "resultant" & acc_placement == "hip") %>%
@@ -54,9 +54,13 @@ scatterplot_GRF_res_hip <- mechanical_load_data %>%
     color = guide_legend(title = "Body mass index category:"),
     shape = guide_legend(title = "Locomotion type:")
   ) +
-  labs(title = "Hip", x = quote("pRACC" ~ (italic(g))), y = "pRGRF (N)")
+  labs(
+    title = "Resultant vector - Hip placement",
+    x = quote("pACC" ~ (italic(g))),
+    y = "pGRF (N)"
+  )
 
-# Hip pRLR x pRAR plot ----------------------------------------------------
+# Hip pLR x pAR plot ----------------------------------------------------
 
 scatterplot_LR_res_hip <- mechanical_load_data %>%
   filter(vector == "resultant" & acc_placement == "hip") %>%
@@ -94,9 +98,9 @@ scatterplot_LR_res_hip <- mechanical_load_data %>%
     shape = guide_legend(title = "Locomotion type:")
   ) +
   labs(
-    title = "Hip",
-    x = quote("pRAR" ~ (italic(g) %.% s^-1)),
-    y = quote("pRLR" ~ (N %.% s^-1))
+    title = "Resultant vector - Hip placement",
+    x = quote("pAR" ~ (italic(g) %.% s^-1)),
+    y = quote("pLR" ~ (N %.% s^-1))
   )
 
 # Combine and save plots --------------------------------------------------

### 2021-03-11 figures/fig3.R

@@ -31,7 +31,7 @@ cv_res_LR_models <- map(
   )
 )
 
-# Hip pRGRF Bland-Altman plot ---------------------------------------------
+# Hip pGRF Bland-Altman plot ---------------------------------------------
 
 BA_GRF_res_hip <- cv_res_GRF_models$hip %>%
   plot_bland_altman(color = BMI_cat, shape = activity) +
@@ -56,12 +56,12 @@ BA_GRF_res_hip <- cv_res_GRF_models$hip %>%
     shape = guide_legend(title = "Locomotion type:")
   ) +
   labs(
-    title = "Hip",
-    x = "Mean of Actual and Predicted pRGRF (N)",
-    y = "Actual - Predicted pRGRF (N)"
+    title = "Resultant vector - Hip placement",
+    x = "Mean of Actual and Predicted pGRF (N)",
+    y = "Actual - Predicted pGRF (N)"
   )
 
-# Hip pRLR Bland-Altman plot ----------------------------------------------
+# Hip pLR Bland-Altman plot ----------------------------------------------
 
 BA_LR_res_hip <- cv_res_LR_models$hip %>%
   plot_bland_altman(color = BMI_cat, shape = activity) +
@@ -92,9 +92,9 @@ BA_LR_res_hip <- cv_res_LR_models$hip %>%
     shape = guide_legend(title = "Locomotion type:")
   ) +
   labs(
-    title = "Hip",
-    x = quote("Mean of Actual and Predicted pRLR" ~ (N %.% s^-1)),
-    y = quote("Actual - Predicted pRLR" ~ (N %.% s^-1))
+    title = "Resultant vector - Hip placement",
+    x = quote("Mean of Actual and Predicted pLR" ~ (N %.% s^-1)),
+    y = quote("Actual - Predicted pLR" ~ (N %.% s^-1))
   )
 
 # Combine and save plots --------------------------------------------------

### 2021-03-11 figures/figS1.R

@@ -55,7 +55,11 @@ scatterplot_GRF_res_ankle <- mechanical_load_data %>%
     color = guide_legend(title = "Body mass index category:"),
     shape = guide_legend(title = "Locomotion type:")
   ) +
-  labs(title = "Ankle", x = quote("pRACC" ~ (italic(g))), y = "pRGRF (N)")
+  labs(
+    title = "Resultant vector - Ankle placement",
+    x = quote("pACC" ~ (italic(g))),
+    y = "pGRF (N)"
+  )
 
 # Resultant: Lower back
 scatterplot_GRF_res_back <- mechanical_load_data %>%
@@ -93,7 +97,11 @@ scatterplot_GRF_res_back <- mechanical_load_data %>%
     color = guide_legend(title = "Body mass index category:"),
     shape = guide_legend(title = "Locomotion type:")
   ) +
-  labs(title = "Lower Back", x = quote("pRACC" ~ (italic(g))), y = "pRGRF (N)")
+  labs(
+    title = "Resultant vector - Lower back placement",
+    x = quote("pACC" ~ (italic(g))),
+    y = "pGRF (N)"
+  )
 
 # Resultant: Hip
 scatterplot_GRF_res_hip <- mechanical_load_data %>%
@@ -131,7 +139,11 @@ scatterplot_GRF_res_hip <- mechanical_load_data %>%
     color = guide_legend(title = "Body mass index category:"),
     shape = guide_legend(title = "Locomotion type:")
   ) +
-  labs(title = "Hip", x = quote("pRACC" ~ (italic(g))), y = "pRGRF (N)")
+  labs(
+    title = "Resultant vector - Hip placement",
+    x = quote("pACC" ~ (italic(g))),
+    y = "pGRF (N)"
+  )
 
 # Vertical: Ankle
 scatterplot_GRF_ver_ankle <- mechanical_load_data %>%
@@ -169,7 +181,11 @@ scatterplot_GRF_ver_ankle <- mechanical_load_data %>%
     color = guide_legend(title = "Body mass index category:"),
     shape = guide_legend(title = "Locomotion type:")
   ) +
-  labs(title = "Ankle", x = quote("pVACC" ~ (italic(g))), y = "pVGRF (N)")
+  labs(
+    title = "Vertical vector - Ankle placement",
+    x = quote("pACC" ~ (italic(g))),
+    y = "pGRF (N)"
+  )
 
 # Vertical: Lower back
[…]

### 2021-03-11 figures/figS2.R

@@ -56,9 +56,9 @@ scatterplot_LR_res_ankle <- mechanical_load_data %>%
     shape = guide_legend(title = "Locomotion type:")
   ) +
   labs(
-    title = "Ankle",
-    x = quote("pRAR" ~ (italic(g) %.% s^-1)),
-    y = quote("pRLR" ~ (N %.% s^-1))
+    title = "Resultant vector - Ankle placement",
+    x = quote("pAR" ~ (italic(g) %.% s^-1)),
+    y = quote("pLR" ~ (N %.% s^-1))
   )
 
 # Resultant: Lower back
@@ -98,9 +98,9 @@ scatterplot_LR_res_back <- mechanical_load_data %>%
     shape = guide_legend(title = "Locomotion type:")
   ) +
   labs(
-    title = "Lower Back",
-    x = quote("pRAR" ~ (italic(g) %.% s^-1)),
-    y = quote("pRLR" ~ (N %.% s^-1))
+    title = "Resultant vector - Lower back placement",
+    x = quote("pAR" ~ (italic(g) %.% s^-1)),
+    y = quote("pLR" ~ (N %.% s^-1))
   )
 
 # Resultant: Hip
@@ -140,9 +140,9 @@ scatterplot_LR_res_hip <- mechanical_load_data %>%
     shape = guide_legend(title = "Locomotion type:")
   ) +
   labs(
-    title = "Hip",
-    x = quote("pRAR" ~ (italic(g) %.% s^-1)),
-    y = quote("pRLR" ~ (N %.% s^-1))
+    title = "Resultant vector - Hip placement",
+    x = quote("pAR" ~ (italic(g) %.% s^-1)),
+    y = quote("pLR" ~ (N %.% s^-1))
   )
 
 # Vertical: Ankle
@@ -182,9 +182,9 @@ scatterplot_LR_ver_ankle <- mechanical_load_data %>%
     shape = guide_legend(title = "Locomotion type:")
   ) +
   labs(
-    title = "Ankle",
-    x = quote("pVAR" ~ (italic(g) %.% s^-1)),
-    y = quote("pVLR" ~ (N %.% s^-1))
+    title = "Vertical vector - Ankle placement",
+    x = quote("pAR" ~ (italic(g) %.% s^-1)),
+    y = quote("pLR" ~ (N %.% s^-1))
   )
 
 # Vertical: Lower back
@@ -224,9 +224,9 @@ scatterplot_LR_ver_back <- mechanical_load_data %>%
[…]

### 2021-03-11 figures/figS3.R

@@ -57,9 +57,9 @@ BA_GRF_res_ankle <- cv_res_GRF_models$ankle %>%
     shape = guide_legend(title = "Locomotion type:")
   ) +
   labs(
-    title = "Ankle",
-    x = "Mean of Actual and Predicted pRGRF (N)",
-    y = "Actual - Predicted pRGRF (N)"
+    title = "Resultant vector - Ankle placement",
+    x = "Mean of Actual and Predicted pGRF (N)",
+    y = "Actual - Predicted pGRF (N)"
   )
 
 # Resultant: Lower back
@@ -86,9 +86,9 @@ BA_GRF_res_back <- cv_res_GRF_models$lower_back %>%
     shape = guide_legend(title = "Locomotion type:")
   ) +
   labs(
-    title = "Lower Back",
-    x = "Mean of Actual and Predicted pRGRF (N)",
-    y = "Actual - Predicted pRGRF (N)"
+    title = "Resultant vector - Lower back placement",
+    x = "Mean of Actual and Predicted pGRF (N)",
+    y = "Actual - Predicted pGRF (N)"
   )
 
 # Resultant: Hip
@@ -115,9 +115,9 @@ BA_GRF_res_hip <- cv_res_GRF_models$hip %>%
     shape = guide_legend(title = "Locomotion type:")
   ) +
   labs(
-    title = "Hip",
-    x = "Mean of Actual and Predicted pRGRF (N)",
-    y = "Actual - Predicted pRGRF (N)"
+    title = "Resultant vector - Hip placement",
+    x = "Mean of Actual and Predicted pGRF (N)",
+    y = "Actual - Predicted pGRF (N)"
   )
 
 # Vertical: Ankle
@@ -144,9 +144,9 @@ BA_GRF_ver_ankle <- cv_ver_GRF_models$ankle %>%
     shape = guide_legend(title = "Locomotion type:")
   ) +
   labs(
-    title = "Ankle",
-    x = "Mean of Actual and Predicted pVGRF (N)",
-    y = "Actual - Predicted pVGRF (N)"
+    title = "Vertical vector - Ankle placement",
+    x = "Mean of Actual and Predicted pGRF (N)",
+    y = "Actual - Predicted pGRF (N)"
   )
 
 # Vertical: Lower back
@@ -173,9 +173,9 @@ BA_GRF_ver_back <- cv_ver_GRF_models$lower_back %>%
     shape = guide_legend(title = "Locomotion type:")
   ) +
   labs(
[…]

### 2021-03-11 figures/figS4.R

@@ -52,8 +52,8 @@ BA_LR_res_ankle <- ggplot()
 #   ) +
 #   labs(
 #     title = "Ankle",
-#     x = quote("Mean of Actual and Predicted pRLR" ~ (N %.% s^-1)),
-#     y = quote("Actual - Predicted pRLR" ~ (N %.% s^-1))
+#     x = quote("Mean of Actual and Predicted pLR" ~ (N %.% s^-1)),
+#     y = quote("Actual - Predicted pLR" ~ (N %.% s^-1))
 #   )
 
 # Resultant: Lower back
@@ -86,9 +86,9 @@ BA_LR_res_back <- cv_res_LR_models$lower_back %>%
     shape = guide_legend(title = "Locomotion type:")
   ) +
   labs(
-    title = "Lower Back",
-    x = quote("Mean of Actual and Predicted pRLR" ~ (N %.% s^-1)),
-    y = quote("Actual - Predicted pRLR" ~ (N %.% s^-1))
+    title = "Resultant vector - Lower back placement",
+    x = quote("Mean of Actual and Predicted pLR" ~ (N %.% s^-1)),
+    y = quote("Actual - Predicted pLR" ~ (N %.% s^-1))
   )
 
 # Resultant: Hip
@@ -121,9 +121,9 @@ BA_LR_res_hip <- cv_res_LR_models$hip %>%
     shape = guide_legend(title = "Locomotion type:")
   ) +
   labs(
-    title = "Hip",
-    x = quote("Mean of Actual and Predicted pRLR" ~ (N %.% s^-1)),
-    y = quote("Actual - Predicted pRLR" ~ (N %.% s^-1))
+    title = "Resultant vector - Hip placement",
+    x = quote("Mean of Actual and Predicted pLR" ~ (N %.% s^-1)),
+    y = quote("Actual - Predicted pLR" ~ (N %.% s^-1))
   )
 
 # Vertical: Ankle
@@ -145,8 +145,8 @@ BA_LR_ver_ankle <- ggplot()
 #   ) +
 #   labs(
 #     title = "Ankle",
-#     x = quote("Mean of Actual and Predicted pVLR" ~ (N %.% s^-1)),
-#     y = quote("Actual - Predicted pVLR" ~ (N %.% s^-1))
+#     x = quote("Mean of Actual and Predicted pLR" ~ (N %.% s^-1)),
+#     y = quote("Actual - Predicted pLR" ~ (N %.% s^-1))
 #   )
 
 # Vertical: Lower back
[…]

### 2021-03-11 tables/tab1.R

@@ -21,16 +21,15 @@ ver_GRF_tb <- build_formula_table(
 )
 
 tab1 <- rbind(res_GRF_tb, ver_GRF_tb)
-# Reorder table rows
-tab1 <- tab1[c(1, 4, 2, 5, 3, 6), ]
-# Remove placemente name duplicates
-tab1[c(2, 4, 6), 1] <- ""
+# Remove vector name duplicates
+tab1[c(2, 3, 5, 6), 1] <- ""
 
 tab1 <- knitr::kable(
   tab1, booktabs = TRUE, escape = FALSE, linesep = "",
   label = "none",
   caption = "Regession equations, $R^2$ and accuracy indices",
   col.names = c(
+    "Vector",
     linebreak("Accelerometer\nplacement"),
     "Regression equations",
     "$R^2$",
@@ -38,10 +37,10 @@ tab1 <- knitr::kable(
     "MAPE",
     "RMSE"
   ),
-  align = c(rep("l", 2), rep("r", 4))
+  align = c(rep("l", 3), rep("r", 4))
 ) %>%
   footnote(
-    general = "Abbreviations: MAE, mean absolute error; MAPE, mean absolute percent error; pRACC, peak resultant acceleration; pRAR, peak resultant acceleration rate; pRGRF, peak resultant ground reaction force; pRLR, peak resultant loading rate; pVACC, peak vertical acceleration; pVAR, peak vertical acceleration rate; pVGRF, peak vertical ground reaction force; pVLR peak vertical loading rate; RMSE, root mean square error",
+    general = "Abbreviations: MAE, mean absolute error; MAPE, mean absolute percent error; pACC, peak acceleration; pAR, peak acceleration rate; pGRF, peak ground reaction force; pLR, peak loading rate; RMSE, root mean square error",
     general_title = "",
     threeparttable = TRUE
   ) %>%

### 2021-03-11 manuscript/manuscript.Rmd
Revise text

@@
 
-Data collection took place at Porto Biomechanics Laboratory (LABIOMEP-UP) in an instrumented treadmill with built-in FP (AMTI Corporation, Watertown, USA).
+Data collection took place at Porto Biomechanics Laboratory (LABIOMEP-UP) on an instrumented treadmill with built-in FP (AMTI Corporation, Watertown, USA).
 Before starting the data collection, participants were given time to warm up on the treadmill.
@@
 Activity monitors at waist level were secured on the same elastic belt with a clip and the activity monitor placed on the ankle was secured by an elastic belt fixed with adhesive tape.
-In all placements the accelerometer y-axis (vertical) was aligned with the standing body longitudinal axis.
+In all placements the accelerometer vertical axis was aligned with the standing body longitudinal axis.
 All participants wore the same three activity monitors always positioned at the same places.
@@
 Peak GRF (pGRF) was determined as the highest value between 0.4 seconds before and after each pACC.
-The rates of change, LR for the GRF and ACC rate (AR) for the ACC, were computed through a second-order centered derivative from the beginning of the foot contact to the pGRF, and pLR and peak AR (pAR) were defined as the maximum value in the array.
+The rates of change, LR for the GRF and ACC rate (AR) for the ACC, were computed through a second-order centered derivative from the beginning of the foot contact to the pGRF.
+The pLR and peak AR (pAR) were defined as the maximum value in the array.
 
@@
 Body mass (kg) was tested as a predictor in all models.
-In addition, in the pGRF models pACC was tested as predictor, while in the pLR models pAR was tested as predictor.
+Additionally, pACC and pAR were tested as predictors in the pGRF and pLR models, respectively.
 All predictors were tested as fixed effects and were shown to be significant.
@@
 Models were validated using the leave-one-out cross-validation approach [@Staudenmayer2012].
-For leave-one-out cross-validation each participant’s data was separated into a testing dataset (one participant at a time) with the remaining data in the training dataset.
+For leave-one-out cross-validation each participant’s data were separated into a testing dataset (one participant at a time) with the remaining data in the training dataset.
 New linear mixed models with the same outcomes and predictors as determined for the entire sample were developed using the training dataset and then used to predict the outcome for the participant in the testing dataset.
@@
 Finally, our pGRF prediction equation was compared with a previously published reference equation [@Neugebauer2014], that was developed using similar methods.
-This compaarison was performed in three ways using: i) the whole sample; ii) a subsample of walking trials; iii) a subsample of running trials.
-To assess the prediction accuracy, MAE, MAPE and RMSE were calculated using pVGRF values predicted from both equations.
+This comparison was performed in three ways using: i) the whole sample; ii) a subsample of walking trials; iii) a subsample of running trials.
+To assess the prediction accuracy, MAE, MAPE and RMSE were calculated using pGRF values predicted from both equations.
 No pLR prediction equation suited to be compared with ours was found.
@@
 Table 1 shows the regression coefficients, $\mathrm{R^2}$ and accuracy indices from models derived from accelerometers placed at ankle, lower back and hip, and from the resultant vector and its vertical component.
-In all pGRF prediction models, $\mathrm{R^2}$ ranged from `r min(GRF_accuracy$R2_cond)` to `r max(GRF_accuracy$R2_cond)`, which shows that the predictor variables are able to explain in a great extent the outcome.
-Also, MAE ranged between `r myround(min(GRF_accuracy$MAE), 1)` N and `r myround(max(GRF_accuracy$MAE), 1)` N, and the MAPE from `r myround(min(GRF_accuracy$MAPE) * 100, 1)`\% to `r myround(max(GRF_accuracy$MAPE) * 100, 1)`\%, which is a smaller error than previously published equations [@Neugebauer2014; @Fortune2014], and similar to the error from equations previously published by out group, but which included walking only [@Veras2020].
+In all pGRF prediction models, $\mathrm{R^2}$ ranged from `r min(GRF_accuracy$R2_cond)` to `r max(GRF_accuracy$R2_cond)`, which shows that the predictor variables are able to explain the variability in the outcome varibale in a great extent.
+Also, MAE ranged between `r myround(min(GRF_accuracy$MAE), 1)` N and `r myround(max(GRF_accuracy$MAE), 1)` N, and the MAPE from `r myround(min(GRF_accuracy$MAPE) * 100, 1)`\% to `r myround(max(GRF_accuracy$MAPE) * 100, 1)`\%, which is a smaller error than previously published equations [@Neugebauer2014; @Fortune2014], and similar to the error from equations previously published by our group, but which included walking only [@Veras2020].
 RMSE were also computed and shown to range between `r myround(min(GRF_accuracy$RMSE), 1)` N and `r myround(max(GRF_accuracy$RMSE), 1)` N.
@@
 A good agreement in a Bland-Altman plot is observed when values tend to cluster around zero, showing no consistent over or underestimation of the actual values.
-Also, the majority of data points should lie within the limits of agreement, which are determined by \pm 1.96 standard deviation of the bias (average difference between actual and predicted values).
+Also, the majority of data points should lie within the limits of agreement, which are determined by \pm 1.96 standard deviation of the difference between actual and predicted values.
 Bland-Altman plots for the pGRF and pLR resultant vector from hip-worn accelerometers can be seen in Figure 3, and for all accelerometers placements and both resultant and vertical vectors in the Supplemental Figures S3 and S4.
@@
 Accuracy indices of our equations were all remarkably better compared to the reference equation, with a MAPE approximately 2 to 3 times lower.
-Furthermore, our equation performance was constant between walking and running, indicating that it can be used for both activities without loss of accuracy.
+Furthermore, our equation performance was constant between walking and running, indicating that it can be used for both activities without loss of accuracy, which cannot be observed in the reference equation.
 

### 2021-03-13 tables/tab1.R
Make html version of the table

@@ -20,12 +20,12 @@ ver_GRF_tb <- build_formula_table(
   ver_GRF_models, cv_ver_GRF_models, "pGRF", "vertical"
 )
 
-tab1 <- rbind(res_GRF_tb, ver_GRF_tb)
+tab1_df <- rbind(res_GRF_tb, ver_GRF_tb)
 # Remove vector name duplicates
-tab1[c(2, 3, 5, 6), 1] <- ""
+tab1_df[c(2, 3, 5, 6), 1] <- ""
 
 tab1 <- knitr::kable(
-  tab1, booktabs = TRUE, escape = FALSE, linesep = "",
+  tab1_df, booktabs = TRUE, escape = FALSE, linesep = "",
   label = "none",
   caption = "Regession equations, $R^2$ and accuracy indices",
   col.names = c(
@@ -51,3 +51,12 @@ tab1 <- knitr::kable(
     bold = FALSE
   ) %>%
   landscape()
+
+
+tab1_html <- tab1_df %>%
+  kbl(caption = "Regession equations, $R^2$ and accuracy indices") %>%
+  kable_classic(full_width = F, html_font = "Cambria") %>%
+  footnote(
+    general = "Abbreviations: MAE, mean absolute error; MAPE, mean absolute percent error; pACC, peak acceleration; pAR, peak acceleration rate; pGRF, peak ground reaction force; pLR, peak loading rate; RMSE, root mean square error",
+    general_title = ""
+  )

### 2021-03-13 tables/tab2.R

@@ -13,7 +13,7 @@ load(here("output", "sub_analyses_accuracy.rda"))
 
 # Build table -------------------------------------------------------------
 
-tab2 <- data.frame(
+tab2_df <- data.frame(
   Prediction = c("Overall", "Walking", "Running"),
   MAE_our = c(
     accuracy_ver_GRF_models$hip$MAE,
@@ -51,7 +51,10 @@ tab2 <- data.frame(
     across(starts_with("MAPE"), ~ .x * 100),
     across(where(is.numeric), myround),
     across(starts_with("MAPE"), ~ paste0(.x, "\\%"))
-  ) %>%
+  )
+
+
+tab2 <- tab2_df %>%
   kbl(
     booktabs = TRUE, escape = FALSE,
     label = "none2",
@@ -69,3 +72,12 @@ tab2 <- data.frame(
   ) %>%
   kable_styling(position = "center") %>%
   add_header_above(c("", "Our equation" = 3, "Neugebauer equation" = 3))
+
+
+tab2_html <- tab2_df %>%
+  kbl(caption = "Accuracy indices of ours and Neugebauer’s equation to predict peak vertical ground reaction force with data from hip-worn accelerometers") %>%
+  kable_classic(full_width = F, html_font = "Cambria") %>%
+  footnote(
+    general = "Abbreviations: MAE, mean absolute error; MAPE, mean absolute percent error; RMSE, root mean square error",
+    general_title = ""
+  )

### 2021-03-13 code/03_predict.R
Include ankle accelerometer placement data in the LR analyses

@@ -73,20 +73,15 @@ LR_formula <- as.formula(
   "pLR_Ns ~ pAR_gs + body_mass + pAR_gs:body_mass + (1 | subj)"
 )
 
-# As LR for the ankle placement is missing at the moment, this placement
-# needs to be removed from the data frame list
-res_mechanical_load_data_LR <- res_mechanical_load_data[-1]
-ver_mechanical_load_data_LR <- ver_mechanical_load_data[-1]
-
 # Resultant vector
 # Build models
 res_LR_models <- map(
-  res_mechanical_load_data_LR,
+  res_mechanical_load_data,
   ~ lmer(LR_formula, data = .x)
 )
 # Cross-validate (leave-one-out cross-validation)
 cv_res_LR_models <- map2(
-  res_LR_models, res_mechanical_load_data_LR,
+  res_LR_models, res_mechanical_load_data,
   ~ loo_cv(.x, .y, id = subj)
 )
 # Compute accuracy indices
@@ -99,12 +94,12 @@ plot_res_LR_models <- map(
 # Vertical vector
 # Build models
 ver_LR_models <- map(
-  ver_mechanical_load_data_LR,
+  ver_mechanical_load_data,
   ~ lmer(LR_formula, data = .x)
 )
 # Cross-validate (leave-one-out cross-validation)
 cv_ver_LR_models <- map2(
-  ver_LR_models, ver_mechanical_load_data_LR,
+  ver_LR_models, ver_mechanical_load_data,
   ~ loo_cv(.x, .y, id = subj)
 )
 # Compute accuracy indices

### 2021-03-13 figures/figS4.R
Add plots for ankle accelerometer placement

@@ -34,27 +34,39 @@ cv_ver_LR_models <- map(
 # LR plots ----------------------------------------------------------------
 
 # Resultant: Ankle
-BA_LR_res_ankle <- ggplot()
-# BA_LR_res_ankle <- cv_res_LR_models$ankle %>%
-#   plot_bland_altman(color = BMI_cat, shape = activity) +
-#   guides(shape = FALSE) +
-#   scale_color_nejm() +
-#   scale_y_continuous(
-#     labels = scales::label_number(),
-#     limits = c(-100000, 125000),
-#     expand = c(0, 0),
-#     breaks = seq(-100000, 125000, 50000)
-#   ) +
-#   theme_light() +
-#   theme(
-#     plot.title = element_text(hjust = 0.5),
-#     legend.title = element_blank()
-#   ) +
-#   labs(
-#     title = "Ankle",
-#     x = quote("Mean of Actual and Predicted pLR" ~ (N %.% s^-1)),
-#     y = quote("Actual - Predicted pLR" ~ (N %.% s^-1))
-#   )
+BA_LR_res_ankle <- cv_res_LR_models$ankle %>%
+  plot_bland_altman(color = BMI_cat, shape = activity) +
+  scale_color_nejm() +
+  scale_y_continuous(
+    labels = scales::label_number(),
+    limits = c(-20000, 20000),
+    expand = c(0, 0),
+    breaks = seq(-20000, 20000, 10000)
+  ) +
+  scale_x_continuous(
+    limits = c(2500, 27500),
+    expand = c(0, 0),
+    breaks = seq(5000, 25000, 5000)
+  ) +
+  theme_light() +
+  theme(
+    plot.title = element_text(size = 15, hjust = 0.5),
+    legend.title = element_text(size = 13),
+    legend.text = element_text(size = 12),
+    axis.title.y = element_text(size = 13),
+    axis.title.x = element_text(size = 13),
+    axis.text.y = element_text(size = 13),
+    axis.text.x = element_text(size = 13)
+  ) +
+  guides(
+    color = guide_legend(title = "Body mass index category:"),
+    shape = guide_legend(title = "Locomotion type:")
+  ) +
+  labs(
+    title = "Resultant vector - Ankle placement",
+    x = quote("Mean of Actual and Predicted pLR" ~ (N %.% s^-1)),
[…]

### 2021-03-13 tables/tab1.R
Add LR data to the table

@@ -19,10 +19,16 @@ res_GRF_tb <- build_formula_table(
 ver_GRF_tb <- build_formula_table(
   ver_GRF_models, cv_ver_GRF_models, "pGRF", "vertical"
 )
+res_LR_tb <- build_formula_table(
+  res_LR_models, cv_res_LR_models, "pLR", "resultant"
+)
+ver_LR_tb <- build_formula_table(
+  ver_LR_models, cv_ver_LR_models, "pLR", "vertical"
+)
 
-tab1_df <- rbind(res_GRF_tb, ver_GRF_tb)
+tab1_df <- rbind(res_GRF_tb, ver_GRF_tb, res_LR_tb, ver_LR_tb)
 # Remove vector name duplicates
-tab1_df[c(2, 3, 5, 6), 1] <- ""
+tab1_df[c(2, 3, 5, 6, 8, 9, 11, 12), 1] <- ""
 
 tab1 <- knitr::kable(
   tab1_df, booktabs = TRUE, escape = FALSE, linesep = "",
@@ -47,6 +53,11 @@ tab1 <- knitr::kable(
   kable_styling(latex_options = "scale_down") %>%
   pack_rows(
     "Peak ground reaction force prediction", 1, 6,
+    latex_gap_space = "0.5em",
+    bold = FALSE
+  ) %>%
+  pack_rows(
+    "Peak loading rate prediction", 7, 12,
     latex_gap_space = "0.5em",
     bold = FALSE
   ) %>%

### 2021-03-13 code/funs.R
Write LR unit in LaTeX notation

@@ -62,7 +62,7 @@ get_equation <- function(model, outcome) {
     model_outcome <- "pGRF (N)"
     model_acceleration <- "pACC"
   } else if (stringr::str_detect(outcome, "LR")) {
-    model_outcome <- "pLR (Ns)"
+    model_outcome <- "pLR ($\\mathrm{N\\cdot s^{-1}}$)"
     model_acceleration <- "pAR"
   }
 

### 2021-03-15 manuscript/manuscript.Rmd
Update results

@@
 
-FP and accelerometer data were processed using MATLAB (Version 8.3, Mathworks, Natick, USA), as described elsewhere [@Veras2020].
+FP and accelerometer data were processed using MATLAB (Version 2019a, Mathworks, Natick, USA), as described elsewhere [@Veras2020].
 Briefly, both GRF and ACC signals were filtered using a Butterworth fourth-order low-pass filter, with 20 Hz cut-off frequency, and their resultant vector calculated ($r_i = \sqrt{x_i^2 + y_i^2 + z_i^2}$).
@@
 Table 1 shows the regression coefficients, $\mathrm{R^2}$ and accuracy indices from models derived from accelerometers placed at ankle, lower back and hip, and from the resultant vector and its vertical component.
-In all pGRF prediction models, $\mathrm{R^2}$ ranged from `r min(GRF_accuracy$R2_cond)` to `r max(GRF_accuracy$R2_cond)`, which shows that the predictor variables are able to explain the variability in the outcome varibale in a great extent.
+In all pGRF prediction models, $\mathrm{R^2}$ values were `r max(GRF_accuracy$R2_cond)`, which shows that the predictor variables are able to explain the variability in the outcome varibale to a great extent.
 Also, MAE ranged between `r myround(min(GRF_accuracy$MAE), 1)` N and `r myround(max(GRF_accuracy$MAE), 1)` N, and the MAPE from `r myround(min(GRF_accuracy$MAPE) * 100, 1)`\% to `r myround(max(GRF_accuracy$MAPE) * 100, 1)`\%, which is a smaller error than previously published equations [@Neugebauer2014; @Fortune2014], and similar to the error from equations previously published by our group, but which included walking only [@Veras2020].
@@
   dplyr::mutate(
-    placement = c("lower back", "hip"),
+    placement = c("ankle", "lower back", "hip"),
     vector = "resultant"
@@
   dplyr::mutate(
-    placement = c("lower back", "hip"),
+    placement = c("ankle", "lower back", "hip"),
     vector = "vertical"
@@
 Regression coefficients, $\mathrm{R^2}$ and accuracy indices for these equations are also shown in Table 1.
-For these equations, the lowest $\mathrm{R^2}$ found was `r min(LR_accuracy$R2_cond)` and the highest was `r max(LR_accuracy$R2_cond)`, reflecting a moderate prediction ability.
+For these equations, the lowest $\mathrm{R^2}$ found was `r myround(min(LR_accuracy$R2_cond), 2)` and the highest was `r myround(max(LR_accuracy$R2_cond), 2)`, reflecting a moderate to high prediction ability.
 MAE, MAPE and RMSE values ranged from `r myround(min(LR_accuracy$MAE), 1)` N to `r myround(max(LR_accuracy$MAE), 1)` N, from `r myround(min(LR_accuracy$MAPE) * 100, 1)`\% to `r myround(max(LR_accuracy$MAPE) * 100, 1)`\%, and from `r myround(min(LR_accuracy$RMSE), 1)` N to `r myround(max(LR_accuracy$RMSE), 1)` N, respectively.
@@
 The accuracy indices values were also somewhat worse than previously published pLR prediction equations [@Veras2020; @Fortune2014], although equations from both studies were only validate for walking.
-The best-performing pLR prediction equation was the one for accelerometers worn at lower back and the resultant vector, with an $\mathrm{R^2}$ of `r back_LR_accuracy$R2_cond` and a MAE, MAPE and RMSE of `r myround(back_LR_accuracy$MAE, 1)` N, `r myround(back_LR_accuracy$MAPE * 100, 1)`\% and `r myround(back_LR_accuracy$RMSE, 1)` N, respectively.
+The pLR equation with the lower prediction error was the one for accelerometers worn at lower back and the resultant vector, with an $\mathrm{R^2}$ of `r back_LR_accuracy$R2_cond` and a MAE, MAPE and RMSE of `r myround(back_LR_accuracy$MAE, 1)` N, `r myround(back_LR_accuracy$MAPE * 100, 1)`\% and `r myround(back_LR_accuracy$RMSE, 1)` N, respectively.
 
@@
 source(here("code", "04_sub_analyses.R"))
+ba_regression_ankle_R2 <- c(
+  ba_regression_res_GRF$ankle["adj.r.squared"],
+  ba_regression_ver_GRF$ankle["adj.r.squared"],
+  ba_regression_res_LR$ankle["adj.r.squared"],
+  ba_regression_ver_LR$ankle["adj.r.squared"]
+) %>%
+  as_vector()
 ```
@@
 For the pGRF equations, a tendency to higher dispersion can be observed in the higher magnitudes, which correspond to the running trials, and especially in values derived from ankle-worn accelerometers (Supplemental Figure S3, Panels A and D).
-Prediction equations for pLR showed a more accentuated dispersion, and again in the higher magnitudes.
-None of the regression models presented a bias different than zero ($p > 0.05$), demonstrating that these models do not consistently over or underestimate the pGRF and pLR values.
+Prediction equations for pLR showed a more accentuated dispersion, and again in the higher magnitudes (Supplemental Figure S4).
+The pLR prediction model based on data from accelerometers placed at the ankle and the resultant vector displayed a bias different than zero ($p = `r myround(ba_bias_test_res_LR$ankle["p.value"], 2)`$).
+None of the other models presented a bias different than zero ($p > 0.05$), demonstrating that they do not consistently over or underestimate the pGRF and pLR values.
 Furthermore, linear regressions were applied to identify proportional bias, that is, to test whether pGRF and pLR prediction errors were systematically influenced by their magnitude.
-The pGRF prediction equation based on ankle-worn accelerometers data for the resultant vector showed a significant proportional bias ($p = `r myround(coefficients(ba_regression_res_GRF$ankle)[2, 4], 2)`$), although with a low magnitude ($\mathrm{R^2} = `r myround(ba_regression_res_GRF$ankle["adj.r.squared"], 2)`$).
+All equations based on ankle-worn accelerometer data, for both outcomes (pGRF and pLR) and vectors (resultant and vertical) showed a significant proportional bias ($p < 0.05$).
+Despite the statistical significance, these linear regressions presented a low $\mathrm{R^2}$ (highest $\mathrm{R^2} = `r myround(max(ba_regression_ankle_R2), 2)`$).
 

### 2021-03-15 figures/fig2.R
Change axes limits

@@ -79,9 +79,9 @@ scatterplot_LR_res_hip <- mechanical_load_data %>%
     breaks = seq(0, 60000, 10000)
   ) +
   scale_x_continuous(
-    limits = c(0, 175),
+    limits = c(0, 180),
     expand = c(0, 0),
-    breaks = seq(25, 170, 25)
+    breaks = seq(0, 175, 25)
   ) +
   theme_light() +
   theme(

### 2021-03-15 figures/fig3.R

@@ -73,9 +73,9 @@ BA_LR_res_hip <- cv_res_LR_models$hip %>%
     breaks = seq(-20000, 20000, 10000)
   ) +
   scale_x_continuous(
-    limits = c(0, 50000),
+    limits = c(0, 55000),
     expand = c(0, 0),
-    breaks = seq(5000, 50000, 10000)
+    breaks = seq(0, 50000, 10000)
   ) +
   theme_light() +
   theme(

### 2021-03-15 figures/figS2.R

@@ -32,14 +32,14 @@ scatterplot_LR_res_ankle <- mechanical_load_data %>%
   ) +
   scale_color_nejm() +
   scale_y_continuous(
-    limits = c(0, 30000),
+    limits = c(0, 60000),
     expand = c(0, 0),
-    breaks = seq(0, 30000, 10000)
+    breaks = seq(0, 60000, 10000)
   ) +
   scale_x_continuous(
-    limits = c(0, 50),
+    limits = c(0, 180),
     expand = c(0, 0),
-    breaks = seq(5, 45, 5)
+    breaks = seq(0, 175, 25)
   ) +
   theme_light() +
   theme(
@@ -74,14 +74,14 @@ scatterplot_LR_res_back <- mechanical_load_data %>%
   ) +
   scale_color_nejm() +
   scale_y_continuous(
-    limits = c(0, 30000),
+    limits = c(0, 60000),
     expand = c(0, 0),
-    breaks = seq(0, 30000, 5000)
+    breaks = seq(0, 60000, 10000)
   ) +
   scale_x_continuous(
-    limits = c(0, 50),
+    limits = c(0, 160),
     expand = c(0, 0),
-    breaks = seq(5, 45, 5)
+    breaks = seq(0, 150, 25)
   ) +
   theme_light() +
   theme(
@@ -116,14 +116,14 @@ scatterplot_LR_res_hip <- mechanical_load_data %>%
   ) +
   scale_color_nejm() +
   scale_y_continuous(
-    limits = c(0, 30000),
+    limits = c(0, 60000),
     expand = c(0, 0),
-    breaks = seq(0, 30000, 5000)
+    breaks = seq(0, 60000, 10000)
   ) +
   scale_x_continuous(
-    limits = c(0, 50),
+    limits = c(0, 210),
     expand = c(0, 0),
-    breaks = seq(5, 45, 5)
+    breaks = seq(0, 200, 25)
   ) +
   theme_light() +
   theme(
@@ -158,14 +158,14 @@ scatterplot_LR_ver_ankle <- mechanical_load_data %>%
   ) +
   scale_color_nejm() +
   scale_y_continuous(
-    limits = c(0, 30000),
+    limits = c(0, 60000),
     expand = c(0, 0),
-    breaks = seq(0, 30000, 5000)
+    breaks = seq(0, 60000, 10000)
   ) +
   scale_x_continuous(
-    limits = c(0, 50),
[…]

### 2021-03-15 figures/figS4.R

@@ -44,9 +44,9 @@ BA_LR_res_ankle <- cv_res_LR_models$ankle %>%
     breaks = seq(-20000, 20000, 10000)
   ) +
   scale_x_continuous(
-    limits = c(2500, 27500),
+    limits = c(0, 55000),
     expand = c(0, 0),
-    breaks = seq(5000, 25000, 5000)
+    breaks = seq(0, 50000, 10000)
   ) +
   theme_light() +
   theme(
@@ -79,9 +79,9 @@ BA_LR_res_back <- cv_res_LR_models$lower_back %>%
     breaks = seq(-20000, 20000, 10000)
   ) +
   scale_x_continuous(
-    limits = c(2500, 27500),
+    limits = c(0, 55000),
     expand = c(0, 0),
-    breaks = seq(5000, 25000, 5000)
+    breaks = seq(0, 50000, 10000)
   ) +
   theme_light() +
   theme(
@@ -114,9 +114,9 @@ BA_LR_res_hip <- cv_res_LR_models$hip %>%
     breaks = seq(-20000, 20000, 10000)
   ) +
   scale_x_continuous(
-    limits = c(2500, 27500),
+    limits = c(0, 55000),
     expand = c(0, 0),
-    breaks = seq(5000, 25000, 5000)
+    breaks = seq(0, 50000, 10000)
   ) +
   theme_light() +
   theme(
@@ -149,9 +149,9 @@ BA_LR_ver_ankle <- cv_ver_LR_models$ankle %>%
     breaks = seq(-20000, 20000, 10000)
   ) +
   scale_x_continuous(
-    limits = c(2500, 27500),
+    limits = c(0, 55000),
     expand = c(0, 0),
-    breaks = seq(5000, 25000, 5000)
+    breaks = seq(0, 50000, 10000)
   ) +
   theme_light() +
   theme(
@@ -184,9 +184,9 @@ BA_LR_ver_back <- cv_ver_LR_models$lower_back %>%
     breaks = seq(-20000, 20000, 10000)
   ) +
   scale_x_continuous(
-    limits = c(2500, 27500),
+    limits = c(0, 65000),
     expand = c(0, 0),
-    breaks = seq(5000, 25000, 5000)
+    breaks = seq(0, 60000, 10000)
   ) +
   theme_light() +
   theme(
@@ -219,9 +219,9 @@ BA_LR_ver_hip <- cv_ver_LR_models$hip %>%
[…]

### 2021-03-15 tables/tab1.R
Correct a typo

@@ -33,7 +33,7 @@ tab1_df[c(2, 3, 5, 6, 8, 9, 11, 12), 1] <- ""
 tab1 <- knitr::kable(
   tab1_df, booktabs = TRUE, escape = FALSE, linesep = "",
   label = "none",
-  caption = "Regession equations, $R^2$ and accuracy indices",
+  caption = "Regression equations, $R^2$ and accuracy indices",
   col.names = c(
     "Vector",
     linebreak("Accelerometer\nplacement"),

### 2021-03-15 tables/tab2.R
Correct a typo

@@ -58,7 +58,7 @@ tab2 <- tab2_df %>%
   kbl(
     booktabs = TRUE, escape = FALSE,
     label = "none2",
-    caption = "Accuracy indices of ours and Neugebauer’s equation to predict peak vertical ground reaction force with data from hip-worn accelerometers",
+    caption = "Accuracy indices of ours and Neugebauer’s equations to predict peak vertical ground reaction force with data from hip-worn accelerometers",
     col.names = c(
       "Prediction",
       "MAE", "MAPE", "RMSE",

### 2021-03-16 manuscript/manuscript.Rmd
Add FS suggestions

@@
 Walking and running speeds ranged from 2 to 6 $\mathrm{km\cdot h^{-1}}$ and from 7 to 14 $\mathrm{km\cdot h^{-1}}$, respectively.
-Each walking speed lasted for 1 minute and each running speed for 30 seconds.
+Each walking speed lasted 1 minute and each running speed 30 seconds.
 To ensure single foot contact on each FP during the trial, feedback regarding location on the treadmill was provided by a researcher.
@@
 
-Activity monitors used were the GT9X Link model ($\mathrm{\pm16}$\textit{g} range; ActiGraph, Pensacola, USA) that incorporates primary and secondary triaxial accelerometers.
-As a manufacturer proprietary filter is applied on the primary, but not on the secondary accelerometer raw data, only data from the latter was used in this study, reducing filtering bias and enhancing the data processing replicability.
+Activity monitors used were the GT9X Link model ($\mathrm{\pm16}g$ range; ActiGraph, Pensacola, USA) that incorporates primary and secondary triaxial accelerometers.
+As a manufacturer proprietary filter is applied on the primary, but not on the secondary accelerometer raw data, and therefore only data from the latter was used in this study, reducing filtering bias and enhancing the data processing replicability.
 
@@
 Data from FP and accelerometers were collected at 1000 Hz and 100 Hz sampling frequency, respectively, and operated through the manufacturer supplied software (for the FP: Netforce, Version 3.5.1; AMTI Corporation, Watertown, USA; for the accelerometers: ActiLife version 6.13.3; Actigraph, Pensacola, USA).
-Raw GRF (expressed in N) and ACC (expressed in gravitational ACC units; 1\textit{g} = 9.807 $\mathrm{m\cdot s^{-2}}$) from the x, y and z vectors were exported to plain text files.
+Raw GRF (expressed in N) and ACC (expressed in gravitational ACC units; 1$g$ = 9.807 $\mathrm{m\cdot s^{-2}}$) from the x, y and z vectors were exported to plain text files.
 
@@
 
-Regression equations to predict pGFR and pLR for each accelerometer placement (ankle, lower back and hip) and vector (resultant and vertical) were developed by linear mixed models.
+Regression equations to predict pGFR and pLR in each vector (resultant and vertical) for each accelerometer placement (ankle, lower back and hip) were developed by linear mixed models.
 Body mass (kg) was tested as a predictor in all models.
@@
 Models were validated using the leave-one-out cross-validation approach [@Staudenmayer2012].
-For leave-one-out cross-validation each participant’s data were separated into a testing dataset (one participant at a time) with the remaining data in the training dataset.
+For this, each participant’s data were separated into a testing dataset (one participant at a time) with the remaining data in the training dataset.
 New linear mixed models with the same outcomes and predictors as determined for the entire sample were developed using the training dataset and then used to predict the outcome for the participant in the testing dataset.
@@
 Linear regressions were applied to identify if the magnitude of the mean between actual and predicted values influenced the magnitude of their difference.
-This was done to ascertain if the prediction accuracy was constant throughout the assayed magnitude range [@Giavarina2015].
+This was done to ascertain if the prediction accuracy was constant throughout the assayed magnitude range, which is known as proportional bias [@Giavarina2015].
 These analyses were conducted separately for data from each accelerometer placement and outcome.
@@
 
-```{r bmi_class}
-BMI_avg <- sample_descriptives$BMI_mean
-BMI_class <- NA
-if (BMI_avg >= 18.5 & BMI_avg < 25) {
-  BMI_class <- "normal weight"
-} else if (BMI_avg >= 25 & BMI_avg < 30) {
-  BMI_class <- "overweight"
-} else if (BMI_avg >= 30 & BMI_avg < 35) {
-  BMI_class <- "obesity class I"
-} else if (BMI_avg >= 35 & BMI_avg < 40) {
-  BMI_class <- "obesity class II"
-} else if (BMI_avg >= 40) {
-  BMI_class <- "obesity class III"
-}
-```
-
-Our recruited sample presented an average BMI of `r sample_descriptives$BMI_mean` $\mathrm{kg\cdot m^{-2}}$, which is relatively high and classified as `r BMI_class`. 
+Our recruited sample presented a broad range of body masses, from normal weight to severe obesity.
 It would be expected that participants with higher BMI values, due to low motor ability and physical capacity, are able to perform the trials only at the lowest speeds.
-This is precisely what happened and, as can be observed in Figure 1, from 2 to 5 $\mathrm{km\cdot h^{-1}}$ the BMI median is around 30 $\mathrm{kg\cdot m^{-2}}$, decreasing to around 27 $\mathrm{kg\cdot m^{-2}}$ at 6 $\mathrm{km\cdot h^{-1}}$, which is the fastest walking speed, with a further reduction to less than 25 $\mathrm{kg\cdot m^{-2}}$ at the running speeds.
+This was precisely what happened and, as can be observed in Figure 1, from 2 to 5 $\mathrm{km\cdot h^{-1}}$ the BMI median was around 30 $\mathrm{kg\cdot m^{-2}}$, decreasing to around 27 $\mathrm{kg\cdot m^{-2}}$ at 6 $\mathrm{km\cdot h^{-1}}$, which was the fastest walking speed, with a further reduction to less than 25 $\mathrm{kg\cdot m^{-2}}$ at the running speeds.
 Furthermore, only one participant with a BMI higher than 30 $\mathrm{kg\cdot m^{-2}}$ was able to run.
@@
 It demonstrates that accelerometer data can adequately discriminate between different BMI categories.
-However, data from the pAR x pLR relationship show a greater dispersion, especially at higher values, and cannot discriminate so well the different BMI categories, as can be observed in Figure 2, Panel B, and Supplemental Figure S2.
+However, data from the pAR x pLR relationship show a greater dispersion, especially at higher values, and the different BMI categories cannot be discriminated so well, as can be observed in Figure 2, Panel B, and Supplemental Figure S2.
 
@@
 
-Afterwards, Bland-Altman plots were built to visualize the agreement between the actual and predicted pGRF and pLR, for both the resultant vector and its vertical component and all accelerometer placements.
-A good agreement in a Bland-Altman plot is observed when values tend to cluster around zero, showing no consistent over or underestimation of the actual values.
-Also, the majority of data points should lie within the limits of agreement, which are determined by \pm 1.96 standard deviation of the difference between actual and predicted values.
-Bland-Altman plots for the pGRF and pLR resultant vector from hip-worn accelerometers can be seen in Figure 3, and for all accelerometers placements and both resultant and vertical vectors in the Supplemental Figures S3 and S4.
+Afterwards, Bland-Altman plots were built to visualize the agreement between the actual and predicted pGRF and pLR, for both the resultant vector and its vertical component from all accelerometer placements.
+Bland-Altman plots for the pGRF and pLR resultant vector from hip-worn accelerometers can be seen in Figure 3, and for all accelerometer placements and both resultant and vertical vectors in the Supplemental Figures S3 and S4.
 Values for all pGRF and pLR prediction equations developed tended to assemble around zero and between the limits of agreement.
 For the pGRF equations, a tendency to higher dispersion can be observed in the higher magnitudes, which correspond to the running trials, and especially in values derived from ankle-worn accelerometers (Supplemental Figure S3, Panels A and D).
-Prediction equations for pLR showed a more accentuated dispersion, and again in the higher magnitudes (Supplemental Figure S4).
+Prediction equations for pLR showed a more accentuated dispersion, and again mainly in the higher magnitudes (Supplemental Figure S4).
 The pLR prediction model based on data from accelerometers placed at the ankle and the resultant vector displayed a bias different than zero ($p = `r myround(ba_bias_test_res_LR$ankle["p.value"], 2)`$).
 None of the other models presented a bias different than zero ($p > 0.05$), demonstrating that they do not consistently over or underestimate the pGRF and pLR values.
-Furthermore, linear regressions were applied to identify proportional bias, that is, to test whether pGRF and pLR prediction errors were systematically influenced by their magnitude.
 All equations based on ankle-worn accelerometer data, for both outcomes (pGRF and pLR) and vectors (resultant and vertical) showed a significant proportional bias ($p < 0.05$).
-Despite the statistical significance, these linear regressions presented a low $\mathrm{R^2}$ (highest $\mathrm{R^2} = `r myround(max(ba_regression_ankle_R2), 2)`$).
+Despite the statistical significance, linear regressions presented a low $\mathrm{R^2}$ (highest $\mathrm{R^2} = `r myround(max(ba_regression_ankle_R2), 2)`$).
 
@@
 The results of these comparisons can be observed in the Table 2.
-Accuracy indices of our equations were all remarkably better compared to the reference equation, with a MAPE approximately 2 to 3 times lower.
+Accuracy indices of our equations were all better compared to the reference equation, with a MAPE approximately 2 to 3 times lower.
 Furthermore, our equation performance was constant between walking and running, indicating that it can be used for both activities without loss of accuracy, which cannot be observed in the reference equation.

### 2021-03-16 manuscript/manuscript.Rmd
Change markup characters

@@
 
-A convenience sample of `r max(sample_size$pGRF_N)` adults (`r unname(sample_size_sex["M"])` males; $`r sample_descriptives$age_mean`\pm`r sample_descriptives$age_sd`$ years; $`r sample_descriptives$height_mean`\pm`r sample_descriptives$height_sd`$ cm; $`r sample_descriptives$body_mass_mean`\pm`r sample_descriptives$body_mass_sd`$ kg; $`r sample_descriptives$BMI_mean`\pm`r sample_descriptives$BMI_sd`$ $\mathrm{kg\cdot m^{-2}}$, $\overline{X}\pm SD$) with no self-reported neurological or musculoskeletal limitations was recruited.
+A convenience sample of `r max(sample_size$pGRF_N)` adults (`r unname(sample_size_sex["M"])` males; `r sample_descriptives$age_mean`±`r sample_descriptives$age_sd` years; `r sample_descriptives$height_mean`±`r sample_descriptives$height_sd` cm; `r sample_descriptives$body_mass_mean`±`r sample_descriptives$body_mass_sd` kg; `r sample_descriptives$BMI_mean`±`r sample_descriptives$BMI_sd` kg·m^-2^, X̄±SD) with no self-reported neurological or musculoskeletal limitations was recruited.
 Height and body mass were assessed following standard procedures with a stadiometer and a digital scale, respectively.
@@
 Before starting the data collection, participants were given time to warm up on the treadmill.
-The protocol consisted of walking and running in 13 incremental speeds, with 0º inclination, from 2 to 14 $\mathrm{km\cdot h^{-1}}$ with 1 $\mathrm{km\cdot h^{-1}}$ increment.
-Walking and running speeds ranged from 2 to 6 $\mathrm{km\cdot h^{-1}}$ and from 7 to 14 $\mathrm{km\cdot h^{-1}}$, respectively.
+The protocol consisted of walking and running in 13 incremental speeds, with 0º inclination, from 2 to 14 km·h^-1^ with 1 km·h^-1^ increment.
+Walking and running speeds ranged from 2 to 6 km·h^-1^ and from 7 to 14 km·h^-1^, respectively.
 Each walking speed lasted 1 minute and each running speed 30 seconds.
@@
 
-Activity monitors used were the GT9X Link model ($\mathrm{\pm16}g$ range; ActiGraph, Pensacola, USA) that incorporates primary and secondary triaxial accelerometers.
+Activity monitors used were the GT9X Link model (±16*g* range; ActiGraph, Pensacola, USA) that incorporates primary and secondary triaxial accelerometers.
 As a manufacturer proprietary filter is applied on the primary, but not on the secondary accelerometer raw data, and therefore only data from the latter was used in this study, reducing filtering bias and enhancing the data processing replicability.
@@
 Data from FP and accelerometers were collected at 1000 Hz and 100 Hz sampling frequency, respectively, and operated through the manufacturer supplied software (for the FP: Netforce, Version 3.5.1; AMTI Corporation, Watertown, USA; for the accelerometers: ActiLife version 6.13.3; Actigraph, Pensacola, USA).
-Raw GRF (expressed in N) and ACC (expressed in gravitational ACC units; 1$g$ = 9.807 $\mathrm{m\cdot s^{-2}}$) from the x, y and z vectors were exported to plain text files.
+Raw GRF (expressed in N) and ACC (expressed in gravitational ACC units; 1*g* = 9.807 m·s^-2^) from the x, y and z vectors were exported to plain text files.
 
@@
 Final models were chosen according to -2 log-likelihood statistics [@Field2012].
-Traditional coefficient of determination ($\mathrm{R^2}$) was represented by conditional $\mathrm{R^2}$, that estimates the variance explained by the whole model [@Nakagawa2013].
+Traditional coefficient of determination (R^2^) was represented by conditional R^2^, that estimates the variance explained by the whole model [@Nakagawa2013].
 
@@
 Also, methods agreement was evaluated with Bland-Altman plots by plotting the difference between the actual (FP measured) and predicted values against their mean.
-Bias was computed as the mean of these differences and the limits of agreement were obtained using \pm 1.96 standard deviation of the mean between actual and predicted values [@Bland1986].
+Bias was computed as the mean of these differences and the limits of agreement were obtained using ±1.96 standard deviation of the mean between actual and predicted values [@Bland1986].
 In order to test whether there was a consistent over or underestimation of the acual values, one-sample \textit{t} tests were performed to assess if the difference between actual and predicted values are different than zero.
@@
 It would be expected that participants with higher BMI values, due to low motor ability and physical capacity, are able to perform the trials only at the lowest speeds.
-This was precisely what happened and, as can be observed in Figure 1, from 2 to 5 $\mathrm{km\cdot h^{-1}}$ the BMI median was around 30 $\mathrm{kg\cdot m^{-2}}$, decreasing to around 27 $\mathrm{kg\cdot m^{-2}}$ at 6 $\mathrm{km\cdot h^{-1}}$, which was the fastest walking speed, with a further reduction to less than 25 $\mathrm{kg\cdot m^{-2}}$ at the running speeds.
-Furthermore, only one participant with a BMI higher than 30 $\mathrm{kg\cdot m^{-2}}$ was able to run.
+This was precisely what happened and, as can be observed in Figure 1, from 2 to 5 km·h^-1^ the BMI median was around 30 kg·m^-2^, decreasing to around 27 kg·m^-2^ at 6 km·h^-1^, which was the fastest walking speed, with a further reduction to less than 25 kg·m^-2^ at the running speeds.
+Furthermore, only one participant with a BMI higher than 30 kg·m^-2^ was able to run.
 This implies in a sample size reduction from `r filter(sample_size, acc_placement == "hip" & vector == "resultant" & speed == "2")$pGRF_N` at the slowest speed to `r filter(sample_size, acc_placement == "hip" & vector == "resultant" & speed == "14")$pGRF_N` at the fastest.
@@
 In the pGRF prediction equations, body mass and pACC were significant predictors and included in the models.
-Table 1 shows the regression coefficients, $\mathrm{R^2}$ and accuracy indices from models derived from accelerometers placed at ankle, lower back and hip, and from the resultant vector and its vertical component.
-In all pGRF prediction models, $\mathrm{R^2}$ values were `r max(GRF_accuracy$R2_cond)`, which shows that the predictor variables are able to explain the variability in the outcome varibale to a great extent.
+Table 1 shows the regression coefficients, R^2^ and accuracy indices from models derived from accelerometers placed at ankle, lower back and hip, and from the resultant vector and its vertical component.
+In all pGRF prediction models, R^2^ values were `r max(GRF_accuracy$R2_cond)`, which shows that the predictor variables are able to explain the variability in the outcome varibale to a great extent.
 Also, MAE ranged between `r myround(min(GRF_accuracy$MAE), 1)` N and `r myround(max(GRF_accuracy$MAE), 1)` N, and the MAPE from `r myround(min(GRF_accuracy$MAPE) * 100, 1)`\% to `r myround(max(GRF_accuracy$MAPE) * 100, 1)`\%, which is a smaller error than previously published equations [@Neugebauer2014; @Fortune2014], and similar to the error from equations previously published by our group, but which included walking only [@Veras2020].
 RMSE were also computed and shown to range between `r myround(min(GRF_accuracy$RMSE), 1)` N and `r myround(max(GRF_accuracy$RMSE), 1)` N.
-Of all the pGRF predictions equations developed, the one based on hip-worn accelerometers and the resultant vector has shown the best performance, with an $\mathrm{R^2}$ of `r hip_GRF_accuracy$R2_cond`, and the lowest MAE (`r myround(hip_GRF_accuracy$MAE, 1)` N), MAPE (`r myround(hip_GRF_accuracy$MAPE * 100, 1)`\%) and RMSE (`r myround(hip_GRF_accuracy$RMSE, 1)` N).
+Of all the pGRF predictions equations developed, the one based on hip-worn accelerometers and the resultant vector has shown the best performance, with an R^2^ of `r hip_GRF_accuracy$R2_cond`, and the lowest MAE (`r myround(hip_GRF_accuracy$MAE, 1)` N), MAPE (`r myround(hip_GRF_accuracy$MAPE * 100, 1)`\%) and RMSE (`r myround(hip_GRF_accuracy$RMSE, 1)` N).
 
@@
 Moreover, pLR prediction equations were developed, and these included as predictors the pAR and again body mass.
-Regression coefficients, $\mathrm{R^2}$ and accuracy indices for these equations are also shown in Table 1.
-For these equations, the lowest $\mathrm{R^2}$ found was `r myround(min(LR_accuracy$R2_cond), 2)` and the highest was `r myround(max(LR_accuracy$R2_cond), 2)`, reflecting a moderate to high prediction ability.
+Regression coefficients, R^2^ and accuracy indices for these equations are also shown in Table 1.
+For these equations, the lowest R^2^ found was `r myround(min(LR_accuracy$R2_cond), 2)` and the highest was `r myround(max(LR_accuracy$R2_cond), 2)`, reflecting a moderate to high prediction ability.
 MAE, MAPE and RMSE values ranged from `r myround(min(LR_accuracy$MAE), 1)` N to `r myround(max(LR_accuracy$MAE), 1)` N, from `r myround(min(LR_accuracy$MAPE) * 100, 1)`\% to `r myround(max(LR_accuracy$MAPE) * 100, 1)`\%, and from `r myround(min(LR_accuracy$RMSE), 1)` N to `r myround(max(LR_accuracy$RMSE), 1)` N, respectively.
@@
 The accuracy indices values were also somewhat worse than previously published pLR prediction equations [@Veras2020; @Fortune2014], although equations from both studies were only validate for walking.
-The pLR equation with the lower prediction error was the one for accelerometers worn at lower back and the resultant vector, with an $\mathrm{R^2}$ of `r back_LR_accuracy$R2_cond` and a MAE, MAPE and RMSE of `r myround(back_LR_accuracy$MAE, 1)` N, `r myround(back_LR_accuracy$MAPE * 100, 1)`\% and `r myround(back_LR_accuracy$RMSE, 1)` N, respectively.
+The pLR equation with the lower prediction error was the one for accelerometers worn at lower back and the resultant vector, with an R^2^ of `r back_LR_accuracy$R2_cond` and a MAE, MAPE and RMSE of `r myround(back_LR_accuracy$MAE, 1)` N, `r myround(back_LR_accuracy$MAPE * 100, 1)`\% and `r myround(back_LR_accuracy$RMSE, 1)` N, respectively.
 
@@
 Prediction equations for pLR showed a more accentuated dispersion, and again mainly in the higher magnitudes (Supplemental Figure S4).
-The pLR prediction model based on data from accelerometers placed at the ankle and the resultant vector displayed a bias different than zero ($p = `r myround(ba_bias_test_res_LR$ankle["p.value"], 2)`$).
-None of the other models presented a bias different than zero ($p > 0.05$), demonstrating that they do not consistently over or underestimate the pGRF and pLR values.
-All equations based on ankle-worn accelerometer data, for both outcomes (pGRF and pLR) and vectors (resultant and vertical) showed a significant proportional bias ($p < 0.05$).
-Despite the statistical significance, linear regressions presented a low $\mathrm{R^2}$ (highest $\mathrm{R^2} = `r myround(max(ba_regression_ankle_R2), 2)`$).
+The pLR prediction model based on data from accelerometers placed at the ankle and the resultant vector displayed a bias different than zero (*p* = `r myround(ba_bias_test_res_LR$ankle["p.value"], 2)`).
+None of the other models presented a bias different than zero (*p* > 0.05), demonstrating that they do not consistently over or underestimate the pGRF and pLR values.
+All equations based on ankle-worn accelerometer data, for both outcomes (pGRF and pLR) and vectors (resultant and vertical) showed a significant proportional bias (*p* < 0.05).
+Despite the statistical significance, linear regressions presented a low R^2^ (highest R^2^ = `r myround(max(ba_regression_ankle_R2), 2)`).
 

### 2021-03-19 manuscript/manuscript.Rmd
Include the Discussion section

@@
 
+# Discussion {.unnumbered}
+
+The purpose of this study was to develop pGRF and pLR prediction equations based on accelerometer data able to accurately estimate the mechanical loading during walking and running.
+The predictions were tested with accelerometers placed at different body locations and using both the resultant vector and its vertical component in a set of incremental speeds, ranging from slow walking to running.
+Our results demonstrate that accelerometer-based regression models are a valid and accurate method to estimate pGRF, especially with data from hip-worn accelerometers, during human locomotion.
+The results also indicate that pLR can also be predicted, however with a substantially lower accuracy compared with the pGRF models.
+
+\medskip
+
+All pGRF models showed an accurate prediction, with a MAPE around 5.5\%, using data from hip-worn accelerometers, and around 6.0\% with data from the other placements.
+On the other hand, the pLR prediction models presented considerably higher errors, as shown by their MAPE around 29\%.
+Also, differently than the pGRF equations, the pLR equations that showed the best accuracy are from the lower back placement.
+Moreover, for both outcomes, predictions from the resultant vector tended to be slightly better than from its vertical component.
+These results are also reflected in the Bland-Altman plots, as the pGRF plots presented considerably narrower limits of agreement and lower dispersion compared to the pLR.
+Furthermore, it can be observed that the Bland-Altman plot fot the pGRF equation for the resultant vector and data from accelerometers placed at the hip presents a high degree of clustering around the zero and very few data points outside the limits of agreement, which confirms it to be the prediction equation that better performs.
+
+\medskip
+
+In respect to the various accelerometer placements analysed in this study, the hip placement showed the best results, mainly in the pGRF models.
+This reinforces the hip placement as the established location to collect and analyse accelerometry-measured physical activity data, as this placement is already well accepted for energy expenditure prediction and physical activity intensity classification [@Migueles2017] and has shown a high mechanical loading prediction ability in this and previous studies [@Veras2020; @Neugebauer2018; @Neugebauer2012].
+Besides the various placements, the current study also analysed predictions from both the resultant and vertical vectors.
+Although ACC and GRF from the vertical vector are the most studied and frequently associated with bone health [@Vainionpaa2006; @Heikkinen2007], the resultant vector of both ACC and GRF during locomotion is greatly determined by the vertical vector, and, as can be observed in Supplemental Figure S1, values from both vectors in this study were very similar.
+Also, mechanical loading prediction based on the resultant ACC offers the important advantage of not being dependent on the accelerometer correct orientation, which is hard to assure when measuring daily physical activity [@Welk2005; @Wundersitz2015].
+
+\medskip
+
+Comparing our results with previous studies that used a similar approach to predict pGRF, it can be seen that prediction errors from our models are comparable with some studies whose best results presented a MAPE of approximately 6.0\% but which included only walking in their protocol [@Fortune2014; @Neugebauer2018; @Veras2020].
+Additionally, previous studies that developed pGRF prediction models during both walking and running showed a MAPE of 8.3\% [@Neugebauer2012] and 5.2\% [@Neugebauer2014], which are higher or similar to our results, respectively.
+Nevertheless, models from both studies present a caveat, that is the inclusion of a term in their equations that distinguishes whether the activity is walking or running, which is not easy to be identify from free-living accelerometer data, hindering the application of these equations.
+Also, one of these studies included in the sample only children and adolescents [@Neugebauer2012].
+It is important to mention that the prediction error found in our pGRF models is lower than the error typically found in models developed to estimate energy expenditure [@Diniz-Sousa2020], which is one of the main uses of accelerometer data [@Migueles2017].
+
+\medskip
+
+To the best of our knowledge, only two studies have already attempted to develop pLR prediction models based on accelerometry data, and both studies only included walking activities.
+Accordingly to what we found, these studies also showed that the prediction error for these models is considerably larger than for the pGRF prediction.
+One of these researches investigated the the use of four different accelerometer placements to predict pLR and achieved a MAPE ranging from 23.4\% to 46.1\% [@Fortune2014], the other one found values from 19.5\% to 22.4\% using accelerometers at the lower back and hip [@Veras2020].
+Results from the models developed in both these investigations ranged from lower to higher prediction error compared to our results, which are approximately of 29.0\%.
+Although pLR is an important variable, known for a long time to affect the bone response to mechanical loading [@Bassey1998; @Turner1995], based on this and previous studies [@Fortune2014; @Veras2020], its prediction through accelerometer data does not seem to yield satisfactory results.
+Nevertheless, past research have suggested that the bone response to pGRF and pLR is somewhat similar due to the high correlation between these variables [@Heikkinen2007].
+Therefore, the pGRF prediction by accelerometry data, which is shown to be accurate, already incorporates some of the information about the effect of LR on the bone response to loading.
+
+\medskip
+
+Results from our study offer some advantages compared to previous investigations: i) it includes walking and running, which are among the most usual types of physical activity performed daily; ii) all models developed, especially for pGRF prediction, presented a similar or better accuracy compared to the results from previous publications; iii) the accuracy of our models is constant throughout walking and running, as shown in Table 2, and therefore they can be used adequately for both activities; iv) in order to apply our models there is no need to distinguish the accelerometer data between walking and running, which facilitates its use with daily living accelerometer data; and v) along with the creation of these models, a software was developed in the form of an R package to perform all the computations needed to get from the raw accelerometer data to the mechanical loading prediction.
+
+\medskip
+
+Besides these advantages, our models present some important potential clinical applications.
+For instance, these models can be used to analyse data from large observational studies and make associations between mechanical loading and bone health.
+But instead of analysing accelerometer data averaged in epochs similarly to some previous publications [@Chastin2014; @Stiles2017], which can dilute the higher intensities [@Fabre2020], with our method raw ACC can be analysed at a higher resolution by identifying the pGRF associated with each step taken and establishing, for example, the number of peaks above certain magnitude needed to elicit a positive bone adaptation.
+Another potential application could be the computation of the cumulative mechanical loading by adding up the pGRF of all detected steps.
+This information could be used either for the determination of the relationship between the mechanical loading accumulated in a certain period (e.g., a day) and the bone response to this loading [@Ahola2010], or also in sports training, by monitoring the amount of loading beared by athletes and relate to the osteoarticular safety of the training regimen [@Vanwanseele2020].
+
+\medskip
+
+Despite the validity of the prediction models developed here, some limitations should be mentioned.
+First, although the activities included in our protocol (walking and running) are among the most performed during daily living, our prediction models are not applicable to other activities known to influence the bone adaptation to mechanical stimuli, such as jumping and direction changing.
+Second, in the absence of another sample for model validation, the leave-one-out cross-validation method was used, which is the method of choice in these situations [@Staudenmayer2012].
+Nevertheless, future studies should test the out-of-sample validity of these models.
+Lastly, the protocol was performed on an instrumented treadmill.
+The use of a treadmill, although enhancing the study internal validity by the continuously measuring the GRF [@Ngoh2018] also reduces the external validity due to possible biomechanical changes in the walking and running patterns induced by the treadmill [@Kluitenberg2012].
+
+\medskip
+
+In conclusion, pGRF could be accurately predicted by accelerometry-based regression models.
+These models are valid for the use of both the resultant vector and its vertical component, and using data from accelerometers placed at either the ankle, lower back or hip.
+The prediction models developed in this study have proved to be an efficient method to assess pGRF out of laboratory conditions, with potential applications of monitoring the mechanical loading in several contexts and establishing the relationship of this loading and either a positive or negative osteoarticular adaptation.
+This is also the first work, to the best of our knowledge, to explore the prediction of pLR based on accelerometry data during both walking and running, although the accuracy of these models were considerably lower than of the pGRF models.
+
 \newpage

### 2021-03-19 manuscript/manuscript.Rmd
Add a paragraph describing the R package in the Results section

@@
 
-Lastly, we compared the performance of our equations with a reference equation previously published.
+Furthermore, we compared the performance of our equations with a reference equation previously published.
 As the only comparable equation found was the one to predict pGRF vertical vector using hip-worn accelerometer data developed by Neugebauer et al [@Neugebauer2014], we applied it to our data and computed the MAE, MAPE and RMSE.
@@
 ```
+
+Also, to simplify the use of these equations and enhace their applicability, a software package was developed for the use in the R statistical software (REF).
+Included in this package are functionalities to read raw ACC data, process it and apply the equations to predict both pGRF and pLR.
+In the package website (ADDRESS) there are instructions on how to use it to obtain the predictions, namely in the 'Home' and 'Get started' sections.
 

### 2021-03-23 manuscript/manuscript.Rmd
Add FS suggestions

@@
 To assess the models prediction accuracy, mean absolute error (MAE), mean absolute percent error (MAPE) and root mean square error (RMSE) were calculated.
+Although there is no standard threshold defining an acceptable error for mechanical loading prediction, based on previous studies we considered a MAPE <8\% to be an accurate prediction [@Fortune2014; @Neugebauer2014; @Ancillao2018].
 Also, methods agreement was evaluated with Bland-Altman plots by plotting the difference between the actual (FP measured) and predicted values against their mean.
 Bias was computed as the mean of these differences and the limits of agreement were obtained using ±1.96 standard deviation of the mean between actual and predicted values [@Bland1986].
-In order to test whether there was a consistent over or underestimation of the acual values, one-sample \textit{t} tests were performed to assess if the difference between actual and predicted values are different than zero.
+In order to test whether there was a consistent over or underestimation of the actual values, one-sample *t* tests were performed to assess if the difference between actual and predicted values are different than zero.
 Linear regressions were applied to identify if the magnitude of the mean between actual and predicted values influenced the magnitude of their difference.
@@
 Table 1 shows the regression coefficients, R^2^ and accuracy indices from models derived from accelerometers placed at ankle, lower back and hip, and from the resultant vector and its vertical component.
-In all pGRF prediction models, R^2^ values were `r max(GRF_accuracy$R2_cond)`, which shows that the predictor variables are able to explain the variability in the outcome varibale to a great extent.
+In all pGRF prediction models, R^2^ values were `r max(GRF_accuracy$R2_cond)`, which shows that the predictor variables are able to explain the variability in the outcome variable to a great extent.
 Also, MAE ranged between `r myround(min(GRF_accuracy$MAE), 1)` N and `r myround(max(GRF_accuracy$MAE), 1)` N, and the MAPE from `r myround(min(GRF_accuracy$MAPE) * 100, 1)`\% to `r myround(max(GRF_accuracy$MAPE) * 100, 1)`\%, which is a smaller error than previously published equations [@Neugebauer2014; @Fortune2014], and similar to the error from equations previously published by our group, but which included walking only [@Veras2020].
@@
 
-Also, to simplify the use of these equations and enhace their applicability, a software package was developed for the use in the R statistical software (REF).
+Also, to simplify the use of these equations and enhance their applicability, a software package was developed for the use in the R statistical software (REF).
 Included in this package are functionalities to read raw ACC data, process it and apply the equations to predict both pGRF and pLR.
@@
 The predictions were tested with accelerometers placed at different body locations and using both the resultant vector and its vertical component in a set of incremental speeds, ranging from slow walking to running.
-Our results demonstrate that accelerometer-based regression models are a valid and accurate method to estimate pGRF, especially with data from hip-worn accelerometers, during human locomotion.
-The results also indicate that pLR can also be predicted, however with a substantially lower accuracy compared with the pGRF models.
-
-\medskip
-
-All pGRF models showed an accurate prediction, with a MAPE around 5.5\%, using data from hip-worn accelerometers, and around 6.0\% with data from the other placements.
-On the other hand, the pLR prediction models presented considerably higher errors, as shown by their MAPE around 29\%.
-Also, differently than the pGRF equations, the pLR equations that showed the best accuracy are from the lower back placement.
-Moreover, for both outcomes, predictions from the resultant vector tended to be slightly better than from its vertical component.
-These results are also reflected in the Bland-Altman plots, as the pGRF plots presented considerably narrower limits of agreement and lower dispersion compared to the pLR.
-Furthermore, it can be observed that the Bland-Altman plot fot the pGRF equation for the resultant vector and data from accelerometers placed at the hip presents a high degree of clustering around the zero and very few data points outside the limits of agreement, which confirms it to be the prediction equation that better performs.
-
-\medskip
-
-In respect to the various accelerometer placements analysed in this study, the hip placement showed the best results, mainly in the pGRF models.
-This reinforces the hip placement as the established location to collect and analyse accelerometry-measured physical activity data, as this placement is already well accepted for energy expenditure prediction and physical activity intensity classification [@Migueles2017] and has shown a high mechanical loading prediction ability in this and previous studies [@Veras2020; @Neugebauer2018; @Neugebauer2012].
-Besides the various placements, the current study also analysed predictions from both the resultant and vertical vectors.
-Although ACC and GRF from the vertical vector are the most studied and frequently associated with bone health [@Vainionpaa2006; @Heikkinen2007], the resultant vector of both ACC and GRF during locomotion is greatly determined by the vertical vector, and, as can be observed in Supplemental Figure S1, values from both vectors in this study were very similar.
-Also, mechanical loading prediction based on the resultant ACC offers the important advantage of not being dependent on the accelerometer correct orientation, which is hard to assure when measuring daily physical activity [@Welk2005; @Wundersitz2015].
+Our findings demonstrated that accelerometer-based regression models were a valid and accurate method to estimate pGRF during human locomotion, as all of the pGRF prediction models developed presented a MAPE below the 8\% threshold established by us.
+However, accelerometry-based equations developed to predict pLR did not yield trustworthy results.
 
@@
 
-To the best of our knowledge, only two studies have already attempted to develop pLR prediction models based on accelerometry data, and both studies only included walking activities.
-Accordingly to what we found, these studies also showed that the prediction error for these models is considerably larger than for the pGRF prediction.
-One of these researches investigated the the use of four different accelerometer placements to predict pLR and achieved a MAPE ranging from 23.4\% to 46.1\% [@Fortune2014], the other one found values from 19.5\% to 22.4\% using accelerometers at the lower back and hip [@Veras2020].
-Results from the models developed in both these investigations ranged from lower to higher prediction error compared to our results, which are approximately of 29.0\%.
-Although pLR is an important variable, known for a long time to affect the bone response to mechanical loading [@Bassey1998; @Turner1995], based on this and previous studies [@Fortune2014; @Veras2020], its prediction through accelerometer data does not seem to yield satisfactory results.
-Nevertheless, past research have suggested that the bone response to pGRF and pLR is somewhat similar due to the high correlation between these variables [@Heikkinen2007].
-Therefore, the pGRF prediction by accelerometry data, which is shown to be accurate, already incorporates some of the information about the effect of LR on the bone response to loading.
+Our findings revealed that either ankle, lower back and hip were valid accelerometer placements to accurately predict pGRF.
+However, despite the similarities, hip was the accelerometer placement with the best prediction accuracy, which reinforces it as the preferable location to collect and analyse accelerometry-measured physical activity data.
+This placement is already well accepted for energy expenditure prediction and physical activity intensity classification [@Migueles2017] and has also shown a high mechanical loading prediction ability in previous studies [@Veras2020; @Neugebauer2018; @Neugebauer2012].
+Besides the various placements, the current study also analysed predictions from both the resultant and vertical vectors.
+Although ACC and GRF from the vertical vector are the most studied and frequently associated with bone health [@Vainionpaa2006; @Heikkinen2007], the resultant vector of both ACC and GRF during locomotion is greatly determined by the vertical vector, and, as can be observed in Supplemental Figure S1, values from both vectors in this study were very similar.
+Also, mechanical loading prediction based on the resultant ACC offers the important advantage of not being dependent on the accelerometer correct orientation, which is hard to assure when measuring daily physical activity [@Welk2005; @Wundersitz2015].
 
@@
 
-Results from our study offer some advantages compared to previous investigations: i) it includes walking and running, which are among the most usual types of physical activity performed daily; ii) all models developed, especially for pGRF prediction, presented a similar or better accuracy compared to the results from previous publications; iii) the accuracy of our models is constant throughout walking and running, as shown in Table 2, and therefore they can be used adequately for both activities; iv) in order to apply our models there is no need to distinguish the accelerometer data between walking and running, which facilitates its use with daily living accelerometer data; and v) along with the creation of these models, a software was developed in the form of an R package to perform all the computations needed to get from the raw accelerometer data to the mechanical loading prediction.
+To the best of our knowledge, only two studies have already attempted to develop pLR prediction models based on accelerometry data, and both studies only included walking activities.
+Accordingly to what we found, these studies also showed that the prediction error for these models is considerably larger than for the pGRF prediction.
+One of these researches investigated the use of four different accelerometer placements (ankle, tibia, thigh and waist) to predict pLR and achieved a MAPE ranging from 23.4\% to 46.1\% [@Fortune2014], the other one found values from 19.5\% to 22.4\% using accelerometers at the lower back and hip [@Veras2020].
+Results from the models developed in both these investigations ranged from lower to higher prediction error compared to our results, which are approximately of 29.0\%.
+Although pLR is an important variable, known for a long time to affect the bone response to mechanical loading [@Bassey1998; @Turner1995], based on this and previous studies [@Fortune2014; @Veras2020], its prediction through accelerometer data does not seem to yield trustworthy results.
+Nevertheless, past research has suggested that the bone response to pGRF and pLR is somewhat similar due to the high correlation between these variables [@Heikkinen2007].
+Therefore, the pGRF prediction by accelerometry data, which is shown to be accurate, already incorporates some of the LR effect on the bone response to loading.
+
+\medskip
+
+Results from our study offer some advantages compared to previous investigations: i) it includes walking and running, which are among the most usual types of physical activity performed daily; ii) all models developed, especially for pGRF prediction, presented a similar or better accuracy compared to the results from previous publications; iii) the accuracy of our models is constant throughout walking and running, and therefore they can be used adequately for both activities; iv) in order to apply our models there is no need to distinguish the accelerometer data between walking and running, which facilitates its use with daily living accelerometer data; and v) along with the creation of these models, a software was developed in the form of an R package to perform all the computations needed to get from the raw accelerometer data to the mechanical loading prediction.
 
@@
 The prediction models developed in this study have proved to be an efficient method to assess pGRF out of laboratory conditions, with potential applications of monitoring the mechanical loading in several contexts and establishing the relationship of this loading and either a positive or negative osteoarticular adaptation.
-This is also the first work, to the best of our knowledge, to explore the prediction of pLR based on accelerometry data during both walking and running, although the accuracy of these models were considerably lower than of the pGRF models.
+This is also the first work, to the best of our knowledge, to explore the prediction of pLR based on accelerometry data during both walking and running, although the accuracy of these models were unsatisfactory.
 

### 2021-04-01 manuscript/manuscript.Rmd
Initiate introduction

@@
 
+Exercise-related mechanical loading plays a key role in bone health, by promoting positive effects on bone mass, geometry and microarchitecture [@Turner2003; @Fonseca2014].
+However, not all forms of mechanical stimulation are equally osteogenic [@Turner1998; @Turner2003] and, therefore, there is a need for the correct assessment of exercise-induced mechanical load.
+
+\medskip
+
+Force plates (FP) are the golden standard instrument for mechanical loading measurement [@Medved2000].
+However, their cost, need of expertise to collect and analyze data, of a dedicated indoor space and ongoing calibration and maintenance hinders its application in clinical settings.
+Accelerometers, due to their portability, low-cost and accuracy, are widely used to objectively measure physical activity related to cardiometabolic outcomes [@Troiano2014; @Migueles2017], and, more recently, have been proposed as adequate devices to predict gravitational loading variables [@Neugebauer2012; @Fortune2014; @Neugebauer2014; @Neugebauer2018; @Veras2020].
+Given the need to assess mechanical loading in clincal settings in order to control and monitor exercise prescription for bone health, accurate mechanical loading prediction models need to be developed and validated for the most common activities prescribed to induce bone adaptations, such as walking, running, and jumping.
+These models, apart from being accurate, need to be transparent in their data processing methods, allowing replicability of the results, and ideally offer a tool to facilitate data processing and analysis for those who are not familiar with the employed techniques, such as what is already offered for analysing cardiometabolic variables [@Migueles2019], and thus spread their applicability.
+
+\medskip
+
+<!--
 The aim of this study was to develop pGRF and pLR prediction models based on accelerometer data during walking and running.
+-->
 
@@
 Peak GRF (pGRF) was determined as the highest value between 0.4 seconds before and after each pACC.
-The rates of change, LR for the GRF and ACC rate (AR) for the ACC, were computed through a second-order centered derivative from the beginning of the foot contact to the pGRF.
+The rates of change, LR for the GRF and ACC rate (AR) for the ACC, were computed through a centered derivative from the beginning of the foot contact to the pGRF.
 The pLR and peak AR (pAR) were defined as the maximum value in the array.

### 2021-04-01 manuscript/manuscript.Rmd
Complement the first paragraph of introduction

@@
 Exercise-related mechanical loading plays a key role in bone health, by promoting positive effects on bone mass, geometry and microarchitecture [@Turner2003; @Fonseca2014].
-However, not all forms of mechanical stimulation are equally osteogenic [@Turner1998; @Turner2003] and, therefore, there is a need for the correct assessment of exercise-induced mechanical load.
+However, not all forms of mechanical stimulation are equally osteogenic, and the manipulation of variables such as loading magnitude, rate, frequency and volume can produce different bone responses [@Turner1998; @Turner2003].
+Therefore, the correct assessment of mechanical loading, specially in a clinical environment, is of paramount importance to attain a significant positive bone response to exercise.
+Nonetheless, a limitation in this context is that the mechanical loading evaluation is usually performed only by instruments restrained to laboratory settings.
 

### 2021-04-02 manuscript/manuscript.Rmd
Finish the Introduction

@@
 
-<!--
-The aim of this study was to develop pGRF and pLR prediction models based on accelerometer data during walking and running.
--->
+The GRF is arguably the most studied mechanical loading variable, and its relationship with the bone adaptations to exercise is well established [@Turner2003; @Rantalainen2008; @Rantalainen2009].
+Due to its importance, some studies have already attempted to develop peak GRF (pGRF) prediction equations based on accelerometry data in several populations, such as children [@Neugebauer2012], adults [@Fortune2014; @Neugebauer2014], patients with obesity [@Veras2020] and military [@Neugebauer2018], but with the majority of these studies including only activities with eliciting a low loading magnitude.
+Besides loading magnitude, LR is another variable related with bone health [@Rowlands2014] that could potentially be predicted by accelerometer data, but with apparently lower prediction accuracy compared with pGRF [@Fortune2014; @Veras2020].
+Also, to the best of our knowledge, pLR prediction through accelerometer data has only been tested during walking.
+
+\medskip
+
+Therefore, given the necessity of an accurate and reliable method to assess mechanical loading in clinical settings in order to better monitor and prescribe exercise to obtain positive bone health outcomes, the aim of this study was to develop pGRF and pLR prediction models based on accelerometer data during walking and running.
 

### 2021-04-02 manuscript/manuscript.Rmd
Confirm abbreviatures

@@
 
-The GRF is arguably the most studied mechanical loading variable, and its relationship with the bone adaptations to exercise is well established [@Turner2003; @Rantalainen2008; @Rantalainen2009].
+The ground reaction force (GRF) is arguably the most studied mechanical loading variable, and its relationship with the bone adaptations to exercise is well established [@Turner2003; @Rantalainen2008; @Rantalainen2009].
 Due to its importance, some studies have already attempted to develop peak GRF (pGRF) prediction equations based on accelerometry data in several populations, such as children [@Neugebauer2012], adults [@Fortune2014; @Neugebauer2014], patients with obesity [@Veras2020] and military [@Neugebauer2018], but with the majority of these studies including only activities with eliciting a low loading magnitude.
-Besides loading magnitude, LR is another variable related with bone health [@Rowlands2014] that could potentially be predicted by accelerometer data, but with apparently lower prediction accuracy compared with pGRF [@Fortune2014; @Veras2020].
-Also, to the best of our knowledge, pLR prediction through accelerometer data has only been tested during walking.
+Besides loading magnitude, loading rate (LR) is another variable related with bone health [@Rowlands2014] that could potentially be predicted by accelerometer data, but with apparently lower prediction accuracy compared with pGRF [@Fortune2014; @Veras2020].
+Also, to the best of our knowledge, peak LR (pLR) prediction through accelerometer data has only been tested during walking.
 
@@
 FP and accelerometer data were processed using MATLAB (Version 2019a, Mathworks, Natick, USA), as described elsewhere [@Veras2020].
-Briefly, both GRF and ACC signals were filtered using a Butterworth fourth-order low-pass filter, with 20 Hz cut-off frequency, and their resultant vector calculated ($r_i = \sqrt{x_i^2 + y_i^2 + z_i^2}$).
+Briefly, both GRF and acceleration (ACC signals were filtered using a Butterworth fourth-order low-pass filter, with 20 Hz cut-off frequency, and their resultant vector calculated ($r_i = \sqrt{x_i^2 + y_i^2 + z_i^2}$).
 After that, peak ACC (pACC) values were defined as minimum height of twice the positive ACC average, during any given speed, and separated by 0.4 seconds minimum.
@@
 Our recruited sample presented a broad range of body masses, from normal weight to severe obesity.
-It would be expected that participants with higher BMI values, due to low motor ability and physical capacity, are able to perform the trials only at the lowest speeds.
+It would be expected that participants with higher body mass index (BMI) values, due to low motor ability and physical capacity, are able to perform the trials only at the lowest speeds.
 This was precisely what happened and, as can be observed in Figure 1, from 2 to 5 km·h^-1^ the BMI median was around 30 kg·m^-2^, decreasing to around 27 kg·m^-2^ at 6 km·h^-1^, which was the fastest walking speed, with a further reduction to less than 25 kg·m^-2^ at the running speeds.

### 2021-04-02 manuscript/manuscript.Rmd
Add Acknowledgements section

@@
 
+# Acknowledgments {.unnumbered}
+
+This study was funded by the Foundation for Science and Technology of Portugal (FCT) (grant PTDC/DTP-DES/0968/2014) and by the European Regional Development Fund (ERDF) through the Operational Competitiveness Programme (COMPETE) (grant POCI-01-0145-FEDER-016707).
+The study was developed in the Research Centre in Physical Activity, Health and Leisure (CIAFEL) funded by ERDF through the COMPETE and by the FCT (grant UIDB/00617/2020).
+Lucas Veras was supported by the FCT (grant UI/BD/150673/2020), Florêncio Diniz-Sousa was supported by the FCT (grant SFRH/BD/117622/2016) and Giorjines Boppre was supported by the FCT (grant SFRH/BD/146976/2019).
+The authors thank the participants who took part in this research and all who have collaborated in the project.
+
 \newpage

### 2021-04-02 manuscript/manuscript.Rmd
Move code chunks to the top of the document
Add Abstract

@@
 library(kableExtra)
+library(english)
 source(here("code", "02_explore.R"))
+```
+
+```{r GRF_accuracy}
+load(here("output", "prediction_models.rda"))
+res_GRF_accuracy <- purrr::map(
+  accuracy_res_GRF_models,
+  ~ dplyr::mutate(.x, MAPE = as.numeric(MAPE))
+) %>%
+  purrr::map_dfr(rbind) %>%
+  dplyr::mutate(
+    placement = c("ankle", "lower back", "hip"),
+    vector = "resultant"
+  )
+ver_GRF_accuracy <- purrr::map(
+  accuracy_ver_GRF_models,
+  ~ dplyr::mutate(.x, MAPE = as.numeric(MAPE))
+) %>%
+  purrr::map_dfr(rbind) %>%
+  dplyr::mutate(
+    placement = c("ankle", "lower back", "hip"),
+    vector = "vertical"
+  )
+GRF_accuracy <- rbind(res_GRF_accuracy, ver_GRF_accuracy)
+hip_GRF_accuracy <- GRF_accuracy %>%
+  filter(placement == "hip" & vector == "resultant")
+```
+
+```{r LR_accuracy}
+load(here("output", "prediction_models.rda"))
+res_LR_accuracy <- purrr::map(
+  accuracy_res_LR_models,
+  ~ dplyr::mutate(.x, MAPE = as.numeric(MAPE))
+) %>%
+  purrr::map_dfr(rbind) %>%
+  dplyr::mutate(
+    placement = c("ankle", "lower back", "hip"),
+    vector = "resultant"
+  )
+ver_LR_accuracy <- purrr::map(
+  accuracy_ver_LR_models,
+  ~ dplyr::mutate(.x, MAPE = as.numeric(MAPE))
+) %>%
+  purrr::map_dfr(rbind) %>%
+  dplyr::mutate(
+    placement = c("ankle", "lower back", "hip"),
+    vector = "vertical"
+  )
+LR_accuracy <- rbind(res_LR_accuracy, ver_LR_accuracy)
+back_LR_accuracy <- LR_accuracy %>%
+  filter(placement == "lower back" & vector == "resultant")
+```
+
+```{r ba_regression}
+source(here("code", "04_sub_analyses.R"))
+ba_regression_ankle_R2 <- c(
+  ba_regression_res_GRF$ankle["adj.r.squared"],
+  ba_regression_ver_GRF$ankle["adj.r.squared"],
+  ba_regression_res_LR$ankle["adj.r.squared"],
+  ba_regression_ver_LR$ankle["adj.r.squared"]
+) %>%
+  as_vector()
 ```
@@
 -->
+
+\leftskip=0.5in \rightskip=0.5in
+
+# Abstract {.unnumbered}
+
+**Purpose:** To develop accelerometry-based equations to predict peak ground reaction forces (pGRF) and peak loading rate (pLR) during walking and running.
+**Methods:** `r Words(max(sample_size$pGRF_N))` subjects (`r unname(sample_size_sex["F"])` females; `r sample_descriptives$body_mass_mean`±`r sample_descriptives$body_mass_sd` kg) walked and ran at different speeds (2–14km·h^-1^) on a force plate–equipped treadmill while wearing accelerometers at ankle, lower back and hip.
+Regression equations were developed to predict pGRF and pLR from accelerometry data.
+Leave-one-out cross-validation was used to calculate prediction accuracy and Bland–Altman plots.
+Our pGRF prediction equation was compared with a reference equation previously published.
+**Results:** Body mass and peak acceleration were included for pGRF prediction and body mass and peak acceleration rate for pLR prediction.
+All pGRF equation coefficients of determination were above `r myround(max(GRF_accuracy$R2_cond), 2)`, and a good agreement between actual and predicted pGRF was observed, with a mean absolute percent error (MAPE) below `r myround(max(GRF_accuracy$MAPE) * 100, 1)`\%.
+Accuracy indices from our equations were better than previously developed equations, with a MAPE 2 to 3 times better.
+All pLR prediction equations presented a lower accuracy compared to those developed to predict pGRF.
+**Conclusion:** Walking and running pGRF can be predicted with high accuracy by accelerometry-based equations, representing an easy way to determine mechanical loading in clinical settings.
+The pLR prediction equations yielded an unsatisfactory accuracy.
+
+\leftskip=0in \rightskip=0in
 
@@
 
-```{r GRF_accuracy}
-load(here("output", "prediction_models.rda"))
-res_GRF_accuracy <- purrr::map(
-  accuracy_res_GRF_models,
-  ~ dplyr::mutate(.x, MAPE = as.numeric(MAPE))
-) %>%
-  purrr::map_dfr(rbind) %>%
-  dplyr::mutate(
-    placement = c("ankle", "lower back", "hip"),
-    vector = "resultant"
-  )
-ver_GRF_accuracy <- purrr::map(
-  accuracy_ver_GRF_models,
-  ~ dplyr::mutate(.x, MAPE = as.numeric(MAPE))
-) %>%
-  purrr::map_dfr(rbind) %>%
-  dplyr::mutate(
-    placement = c("ankle", "lower back", "hip"),
-    vector = "vertical"
-  )
-GRF_accuracy <- rbind(res_GRF_accuracy, ver_GRF_accuracy)
-hip_GRF_accuracy <- GRF_accuracy %>%
-  filter(placement == "hip" & vector == "resultant")
-```
-
 Regression equations were developed to predict the mechanical loading variables based on accelerometry data.
@@
 \medskip
-
-```{r LR_accuracy}
-load(here("output", "prediction_models.rda"))
-res_LR_accuracy <- purrr::map(
-  accuracy_res_LR_models,
-  ~ dplyr::mutate(.x, MAPE = as.numeric(MAPE))
-) %>%
-  purrr::map_dfr(rbind) %>%
-  dplyr::mutate(
-    placement = c("ankle", "lower back", "hip"),
-    vector = "resultant"
-  )
-ver_LR_accuracy <- purrr::map(
-  accuracy_ver_LR_models,
-  ~ dplyr::mutate(.x, MAPE = as.numeric(MAPE))
-) %>%
-  purrr::map_dfr(rbind) %>%
-  dplyr::mutate(
-    placement = c("ankle", "lower back", "hip"),
-    vector = "vertical"
-  )
-LR_accuracy <- rbind(res_LR_accuracy, ver_LR_accuracy)
-back_LR_accuracy <- LR_accuracy %>%
-  filter(placement == "lower back" & vector == "resultant")
-```
 
@@
 tab1
-```
-
-```{r ba_regression}
-source(here("code", "04_sub_analyses.R"))
-ba_regression_ankle_R2 <- c(
-  ba_regression_res_GRF$ankle["adj.r.squared"],
-  ba_regression_ver_GRF$ankle["adj.r.squared"],
-  ba_regression_res_LR$ankle["adj.r.squared"],
-  ba_regression_ver_LR$ankle["adj.r.squared"]
-) %>%
-  as_vector()
 ```

### 2021-04-02 manuscript/manuscript.Rmd
Write a title and uncomment authors and affiliations

@@
 ---
-# title: "The document title"
+title: "Mechanical loading prediction though accelerometry data during walking and running"
 output:
@@
 
-<!--
-\begin{center}
+<center>
 Lucas Veras$^1$, Florêncio Diniz-Sousa$^1$, Giorjines Boppre$^1$, Vítor Devezas$^2$, Hugo Santos-Sousa$^2$, John Preto$^2$, João Paulo Vilas-Boas$^{3,4}$, Leandro Machado$^{3,4}$, José Oliveira$^1$, Hélder Fonseca$^1$
-\end{center}
+</center>
 
@@
 4 - Biomechanics Laboratory (LABIOMEP-UP), University of Porto, Porto, Portugal
--->
 
 \leftskip=0.5in \rightskip=0.5in
+
+\bigskip
 

### 2021-04-03 manuscript/manuscript.Rmd
Add keywords
Add FS suggestions

@@
 All pLR prediction equations presented a lower accuracy compared to those developed to predict pGRF.
-**Conclusion:** Walking and running pGRF can be predicted with high accuracy by accelerometry-based equations, representing an easy way to determine mechanical loading in clinical settings.
+**Conclusion:** Walking and running pGRF can be predicted with high accuracy by accelerometry-based equations, representing an easy way to determine mechanical loading in free-living conditions.
 The pLR prediction equations yielded an unsatisfactory accuracy.
+**Key Words:** ACTIVITY MONITOR, FORCE PLATES, LOCOMOTION, RAW ACCELERATION, MECHANICAL LOADING.
 
@@
 
-Exercise-related mechanical loading plays a key role in bone health, by promoting positive effects on bone mass, geometry and microarchitecture [@Turner2003; @Fonseca2014].
+Physical activity-related mechanical loading plays a key role in bone health, by promoting positive effects on bone mass, geometry and microarchitecture [@Turner2003; @Fonseca2014].
 However, not all forms of mechanical stimulation are equally osteogenic, and the manipulation of variables such as loading magnitude, rate, frequency and volume can produce different bone responses [@Turner1998; @Turner2003].
-Therefore, the correct assessment of mechanical loading, specially in a clinical environment, is of paramount importance to attain a significant positive bone response to exercise.
+Therefore, the correct assessment of mechanical loading is of paramount importance to better understand the bone response to physical stimulus.
 Nonetheless, a limitation in this context is that the mechanical loading evaluation is usually performed only by instruments restrained to laboratory settings.
@@
 Force plates (FP) are the golden standard instrument for mechanical loading measurement [@Medved2000].
-However, their cost, need of expertise to collect and analyze data, of a dedicated indoor space and ongoing calibration and maintenance hinders its application in clinical settings.
-Accelerometers, due to their portability, low-cost and accuracy, are widely used to objectively measure physical activity related to cardiometabolic outcomes [@Troiano2014; @Migueles2017], and, more recently, have been proposed as adequate devices to predict gravitational loading variables [@Neugebauer2012; @Fortune2014; @Neugebauer2014; @Neugebauer2018; @Veras2020].
-Given the need to assess mechanical loading in clincal settings in order to control and monitor exercise prescription for bone health, accurate mechanical loading prediction models need to be developed and validated for the most common activities prescribed to induce bone adaptations, such as walking, running, and jumping.
+However, their cost, need of expertise to collect and analyze data, of a dedicated indoor space and ongoing calibration and maintenance hinders its application in free-living conditions.
+Accelerometers, due to their portability, low-cost and reliability, are widely used to objectively measure physical activity related to cardiometabolic outcomes [@Troiano2014; @Migueles2017], and, more recently, have been proposed as adequate devices to predict gravitational loading variables [@Neugebauer2012; @Fortune2014; @Neugebauer2014; @Neugebauer2018; @Veras2020].
+Given the need to assess mechanical loading in out-of-laboratory environments in order to control and monitor exercise prescription for bone health, accurate mechanical loading prediction models need to be developed and validated for the most common activities prescribed to induce bone adaptations, such as walking, running, and jumping.
 These models, apart from being accurate, need to be transparent in their data processing methods, allowing replicability of the results, and ideally offer a tool to facilitate data processing and analysis for those who are not familiar with the employed techniques, such as what is already offered for analysing cardiometabolic variables [@Migueles2019], and thus spread their applicability.
@@
 
-Therefore, given the necessity of an accurate and reliable method to assess mechanical loading in clinical settings in order to better monitor and prescribe exercise to obtain positive bone health outcomes, the aim of this study was to develop pGRF and pLR prediction models based on accelerometer data during walking and running.
+Therefore, given the necessity of an accurate and reliable method to assess mechanical loading in free-living conditions in order to better monitor and prescribe exercise to obtain positive bone health outcomes, the aim of this study was to develop pGRF and pLR prediction models based on accelerometer data during walking and running.
 
@@
 In all placements the accelerometer vertical axis was aligned with the standing body longitudinal axis.
-All participants wore the same three activity monitors always positioned at the same places.
 
@@
 One of these researches investigated the use of four different accelerometer placements (ankle, tibia, thigh and waist) to predict pLR and achieved a MAPE ranging from 23.4\% to 46.1\% [@Fortune2014], the other one found values from 19.5\% to 22.4\% using accelerometers at the lower back and hip [@Veras2020].
-Results from the models developed in both these investigations ranged from lower to higher prediction error compared to our results, which are approximately of 29.0\%.
+Results from the models developed in these investigations have both lower and higher prediction errors compared to our results, which are approximately of 29.0\%.
 Although pLR is an important variable, known for a long time to affect the bone response to mechanical loading [@Bassey1998; @Turner1995], based on this and previous studies [@Fortune2014; @Veras2020], its prediction through accelerometer data does not seem to yield trustworthy results.

### 2021-04-25 manuscript/manuscript.Rmd
Minor corrections

@@
 Data from FP and accelerometers were collected at 1000 Hz and 100 Hz sampling frequency, respectively, and operated through the manufacturer supplied software (for the FP: Netforce, Version 3.5.1; AMTI Corporation, Watertown, USA; for the accelerometers: ActiLife version 6.13.3; Actigraph, Pensacola, USA).
-Raw GRF (expressed in N) and ACC (expressed in gravitational ACC units; 1*g* = 9.807 m·s^-2^) from the x, y and z vectors were exported to plain text files.
+Raw GRF (expressed in N) and acceleration (ACC; expressed in gravitational ACC units; 1*g* = 9.807 m·s^-2^) from the x, y and z vectors were exported to plain text files.
 
@@
 FP and accelerometer data were processed using MATLAB (Version 2019a, Mathworks, Natick, USA), as described elsewhere [@Veras2020].
-Briefly, both GRF and acceleration (ACC signals were filtered using a Butterworth fourth-order low-pass filter, with 20 Hz cut-off frequency, and their resultant vector calculated ($r_i = \sqrt{x_i^2 + y_i^2 + z_i^2}$).
+Briefly, both GRF and acceleration signals were filtered using a Butterworth fourth-order low-pass filter, with 20 Hz cut-off frequency, and their resultant vector calculated ($r_i = \sqrt{x_i^2 + y_i^2 + z_i^2}$).
 After that, peak ACC (pACC) values were defined as minimum height of twice the positive ACC average, during any given speed, and separated by 0.4 seconds minimum.
@@
 Bias was computed as the mean of these differences and the limits of agreement were obtained using ±1.96 standard deviation of the mean between actual and predicted values [@Bland1986].
-In order to test whether there was a consistent over or underestimation of the actual values, one-sample *t* tests were performed to assess if the difference between actual and predicted values are different than zero.
+In order to test whether there was a consistent over or underestimation of the actual values, one-sample *t* tests were performed to assess if the difference between actual and predicted values were different than zero.
 Linear regressions were applied to identify if the magnitude of the mean between actual and predicted values influenced the magnitude of their difference.

### 2021-05-12 manuscript/manuscript.Rmd
Adapt to the new results

@@
 The ground reaction force (GRF) is arguably the most studied mechanical loading variable, and its relationship with the bone adaptations to exercise is well established [@Turner2003; @Rantalainen2008; @Rantalainen2009].
-Due to its importance, some studies have already attempted to develop peak GRF (pGRF) prediction equations based on accelerometry data in several populations, such as children [@Neugebauer2012], adults [@Fortune2014; @Neugebauer2014], patients with obesity [@Veras2020] and military [@Neugebauer2018], but with the majority of these studies including only activities with eliciting a low loading magnitude.
+Due to its importance, some studies have already attempted to develop peak GRF (pGRF) prediction equations based on accelerometry data in several populations, such as children [@Neugebauer2012], adults [@Fortune2014; @Neugebauer2014], patients with obesity [@Veras2020] and military [@Neugebauer2018], but with the majority of these studies including only activities eliciting a low loading magnitude.
 Besides loading magnitude, loading rate (LR) is another variable related with bone health [@Rowlands2014] that could potentially be predicted by accelerometer data, but with apparently lower prediction accuracy compared with pGRF [@Fortune2014; @Veras2020].
@@
 
-Data collection took place at Porto Biomechanics Laboratory (LABIOMEP-UP) on an instrumented treadmill with built-in FP (AMTI Corporation, Watertown, USA).
-Before starting the data collection, participants were given time to warm up on the treadmill.
+Data collection took place at Porto Biomechanics Laboratory (LABIOMEP-UP) on an instrumented front-back split-belt treadmill with built-in FP (AMTI Corporation, Watertown, USA).
+Before starting the data collection, participants were given time to familiarize with the treadmill.
 The protocol consisted of walking and running in 13 incremental speeds, with 0º inclination, from 2 to 14 km·h^-1^ with 1 km·h^-1^ increment.
@@
 Also, methods agreement was evaluated with Bland-Altman plots by plotting the difference between the actual (FP measured) and predicted values against their mean.
-Bias was computed as the mean of these differences and the limits of agreement were obtained using ±1.96 standard deviation of the mean between actual and predicted values [@Bland1986].
+Bias was computed as the mean of these differences and the limits of agreement were obtained using ±1.96 standard deviation of the bias [@Bland1986].
 In order to test whether there was a consistent over or underestimation of the actual values, one-sample *t* tests were performed to assess if the difference between actual and predicted values were different than zero.
@@
 It would be expected that participants with higher body mass index (BMI) values, due to low motor ability and physical capacity, are able to perform the trials only at the lowest speeds.
-This was precisely what happened and, as can be observed in Figure 1, from 2 to 5 km·h^-1^ the BMI median was around 30 kg·m^-2^, decreasing to around 27 kg·m^-2^ at 6 km·h^-1^, which was the fastest walking speed, with a further reduction to less than 25 kg·m^-2^ at the running speeds.
-Furthermore, only one participant with a BMI higher than 30 kg·m^-2^ was able to run.
-This implies in a sample size reduction from `r filter(sample_size, acc_placement == "hip" & vector == "resultant" & speed == "2")$pGRF_N` at the slowest speed to `r filter(sample_size, acc_placement == "hip" & vector == "resultant" & speed == "14")$pGRF_N` at the fastest.
+This was precisely what happened and, as can be observed in Figure 1, the number of data points corresponding to subjects with BMI > 30 kg·m^-2^ greatly decreases in the speeds above 6 km·h^-1^.
 

### 2021-05-15 figures/fig2.R
Adjust plots axes

@@ -36,9 +36,9 @@ scatterplot_GRF_res_hip <- mechanical_load_data %>%
     breaks = seq(0, 3500, 500)
   ) +
   scale_x_continuous(
-    limits = c(0, 7),
+    limits = c(1, 9),
     expand = c(0, 0),
-    breaks = seq(1, 6, 1)
+    breaks = seq(2, 8, 1)
   ) +
   theme_light() +
   theme(

### 2021-05-15 figures/fig3.R

@@ -41,6 +41,11 @@ BA_GRF_res_hip <- cv_res_GRF_models$hip %>%
     expand = c(0, 0),
     breaks = seq(-600, 600, 200)
   ) +
+  scale_x_continuous(
+    limits = c(400, 3100),
+    expand = c(0, 0),
+    breaks = seq(500, 3000, 500)
+  ) +
   theme_light() +
   theme(
     plot.title = element_text(size = 15, hjust = 0.5),
@@ -75,7 +80,7 @@ BA_LR_res_hip <- cv_res_LR_models$hip %>%
   scale_x_continuous(
     limits = c(0, 55000),
     expand = c(0, 0),
-    breaks = seq(0, 50000, 10000)
+    breaks = seq(10000, 50000, 10000)
   ) +
   theme_light() +
   theme(

### 2021-05-15 figures/figS1.R

@@ -37,9 +37,9 @@ scatterplot_GRF_res_ankle <- mechanical_load_data %>%
     breaks = seq(0, 3500, 500)
   ) +
   scale_x_continuous(
-    limits = c(0, 14),
+    limits = c(1, 14),
     expand = c(0, 0),
-    breaks = seq(1, 14, 2)
+    breaks = seq(0, 12, 2)
   ) +
   theme_light() +
   theme(
@@ -79,9 +79,9 @@ scatterplot_GRF_res_back <- mechanical_load_data %>%
     breaks = seq(0, 3500, 500)
   ) +
   scale_x_continuous(
-    limits = c(0, 6),
+    limits = c(1, 9),
     expand = c(0, 0),
-    breaks = seq(1, 6, 1)
+    breaks = seq(2, 8, 1)
   ) +
   theme_light() +
   theme(
@@ -121,9 +121,9 @@ scatterplot_GRF_res_hip <- mechanical_load_data %>%
     breaks = seq(0, 3500, 500)
   ) +
   scale_x_continuous(
-    limits = c(0, 7),
+    limits = c(1, 9),
     expand = c(0, 0),
-    breaks = seq(1, 6, 1)
+    breaks = seq(2, 8, 1)
   ) +
   theme_light() +
   theme(
@@ -163,9 +163,9 @@ scatterplot_GRF_ver_ankle <- mechanical_load_data %>%
     breaks = seq(0, 3500, 500)
   ) +
   scale_x_continuous(
-    limits = c(0, 10),
+    limits = c(1, 10),
     expand = c(0, 0),
-    breaks = seq(1, 10, 2)
+    breaks = seq(2, 9, 1)
   ) +
   theme_light() +
   theme(
@@ -205,9 +205,9 @@ scatterplot_GRF_ver_back <- mechanical_load_data %>%
     breaks = seq(0, 3500, 500)
   ) +
   scale_x_continuous(
-    limits = c(0, 7),
+    limits = c(1, 8),
     expand = c(0, 0),
-    breaks = seq(1, 6, 1)
+    breaks = seq(2, 7, 1)
   ) +
   theme_light() +
   theme(
@@ -247,9 +247,9 @@ scatterplot_GRF_ver_hip <- mechanical_load_data %>%
     breaks = seq(0, 3500, 500)
   ) +
   scale_x_continuous(
-    limits = c(0, 7),
+    limits = c(1, 8),
[…]

### 2021-05-15 figures/figS3.R

@@ -42,6 +42,11 @@ BA_GRF_res_ankle <- cv_res_GRF_models$ankle %>%
     expand = c(0, 0),
     breaks = seq(-600, 600, 300)
   ) +
+  scale_x_continuous(
+    limits = c(400, 3100),
+    expand = c(0, 0),
+    breaks = seq(500, 3000, 500)
+  ) +
   theme_light() +
   theme(
     plot.title = element_text(size = 15, hjust = 0.5),
@@ -70,6 +75,11 @@ BA_GRF_res_back <- cv_res_GRF_models$lower_back %>%
     limits = c(-600, 600),
     expand = c(0, 0),
     breaks = seq(-600, 600, 200)
+  ) +
+  scale_x_continuous(
+    limits = c(400, 3100),
+    expand = c(0, 0),
+    breaks = seq(500, 3000, 500)
   ) +
   theme_light() +
   theme(
@@ -100,6 +110,11 @@ BA_GRF_res_hip <- cv_res_GRF_models$hip %>%
     expand = c(0, 0),
     breaks = seq(-600, 600, 200)
   ) +
+  scale_x_continuous(
+    limits = c(400, 3100),
+    expand = c(0, 0),
+    breaks = seq(500, 3000, 500)
+  ) +
   theme_light() +
   theme(
     plot.title = element_text(size = 15, hjust = 0.5),
@@ -128,6 +143,11 @@ BA_GRF_ver_ankle <- cv_ver_GRF_models$ankle %>%
     limits = c(-600, 600),
     expand = c(0, 0),
     breaks = seq(-600, 600, 300)
+  ) +
+  scale_x_continuous(
+    limits = c(400, 3100),
+    expand = c(0, 0),
+    breaks = seq(500, 3000, 500)
   ) +
   theme_light() +
   theme(
@@ -158,6 +178,11 @@ BA_GRF_ver_back <- cv_ver_GRF_models$lower_back %>%
     expand = c(0, 0),
     breaks = seq(-600, 600, 200)
   ) +
+  scale_x_continuous(
+    limits = c(400, 3100),
+    expand = c(0, 0),
+    breaks = seq(500, 3000, 500)
+  ) +
   theme_light() +
   theme(
     plot.title = element_text(size = 15, hjust = 0.5),
@@ -186,6 +211,11 @@ BA_GRF_ver_hip <- cv_ver_GRF_models$hip %>%
     limits = c(-600, 600),
     expand = c(0, 0),
[…]

### 2021-05-15 figures/figS4.R

@@ -46,7 +46,7 @@ BA_LR_res_ankle <- cv_res_LR_models$ankle %>%
   scale_x_continuous(
     limits = c(0, 55000),
     expand = c(0, 0),
-    breaks = seq(0, 50000, 10000)
+    breaks = seq(10000, 50000, 10000)
   ) +
   theme_light() +
   theme(
@@ -81,7 +81,7 @@ BA_LR_res_back <- cv_res_LR_models$lower_back %>%
   scale_x_continuous(
     limits = c(0, 55000),
     expand = c(0, 0),
-    breaks = seq(0, 50000, 10000)
+    breaks = seq(10000, 50000, 10000)
   ) +
   theme_light() +
   theme(
@@ -116,7 +116,7 @@ BA_LR_res_hip <- cv_res_LR_models$hip %>%
   scale_x_continuous(
     limits = c(0, 55000),
     expand = c(0, 0),
-    breaks = seq(0, 50000, 10000)
+    breaks = seq(10000, 50000, 10000)
   ) +
   theme_light() +
   theme(
@@ -151,7 +151,7 @@ BA_LR_ver_ankle <- cv_ver_LR_models$ankle %>%
   scale_x_continuous(
     limits = c(0, 55000),
     expand = c(0, 0),
-    breaks = seq(0, 50000, 10000)
+    breaks = seq(10000, 50000, 10000)
   ) +
   theme_light() +
   theme(
@@ -186,7 +186,7 @@ BA_LR_ver_back <- cv_ver_LR_models$lower_back %>%
   scale_x_continuous(
     limits = c(0, 65000),
     expand = c(0, 0),
-    breaks = seq(0, 60000, 10000)
+    breaks = seq(10000, 60000, 10000)
   ) +
   theme_light() +
   theme(
@@ -221,7 +221,7 @@ BA_LR_ver_hip <- cv_ver_LR_models$hip %>%
   scale_x_continuous(
     limits = c(0, 55000),
     expand = c(0, 0),
-    breaks = seq(0, 50000, 10000)
+    breaks = seq(10000, 50000, 10000)
   ) +
   theme_light() +
   theme(

### 2021-05-15 manuscript/manuscript.Rmd
Adjust manuscript text

@@
 LR_accuracy <- rbind(res_LR_accuracy, ver_LR_accuracy)
-back_LR_accuracy <- LR_accuracy %>%
-  filter(placement == "lower back" & vector == "resultant")
+hip_LR_accuracy <- LR_accuracy %>%
+  filter(placement == "hip" & vector == "resultant")
 ```
@@
 Table 1 shows the regression coefficients, R^2^ and accuracy indices from models derived from accelerometers placed at ankle, lower back and hip, and from the resultant vector and its vertical component.
-In all pGRF prediction models, R^2^ values were `r max(GRF_accuracy$R2_cond)`, which shows that the predictor variables are able to explain the variability in the outcome variable to a great extent.
+In the pGRF prediction models, the lowest R^2^ value found was `r min(GRF_accuracy$R2_cond)`, which shows that the predictor variables are able to explain the variability in the outcome variable to a great extent.
 Also, MAE ranged between `r myround(min(GRF_accuracy$MAE), 1)` N and `r myround(max(GRF_accuracy$MAE), 1)` N, and the MAPE from `r myround(min(GRF_accuracy$MAPE) * 100, 1)`\% to `r myround(max(GRF_accuracy$MAPE) * 100, 1)`\%, which is a smaller error than previously published equations [@Neugebauer2014; @Fortune2014], and similar to the error from equations previously published by our group, but which included walking only [@Veras2020].
 RMSE were also computed and shown to range between `r myround(min(GRF_accuracy$RMSE), 1)` N and `r myround(max(GRF_accuracy$RMSE), 1)` N.
-Of all the pGRF predictions equations developed, the one based on hip-worn accelerometers and the resultant vector has shown the best performance, with an R^2^ of `r hip_GRF_accuracy$R2_cond`, and the lowest MAE (`r myround(hip_GRF_accuracy$MAE, 1)` N), MAPE (`r myround(hip_GRF_accuracy$MAPE * 100, 1)`\%) and RMSE (`r myround(hip_GRF_accuracy$RMSE, 1)` N).
+Although all pGRF prediction equations have shown very similar results, the one based on hip-worn accelerometers and the resultant vector has shown the best performance, with an R^2^ of `r hip_GRF_accuracy$R2_cond`, and the lowest MAE (`r myround(hip_GRF_accuracy$MAE, 1)` N), MAPE (`r myround(hip_GRF_accuracy$MAPE * 100, 1)`\%) and RMSE (`r myround(hip_GRF_accuracy$RMSE, 1)` N).
 
@@
 The accuracy indices values were also somewhat worse than previously published pLR prediction equations [@Veras2020; @Fortune2014], although equations from both studies were only validate for walking.
-The pLR equation with the lower prediction error was the one for accelerometers worn at lower back and the resultant vector, with an R^2^ of `r back_LR_accuracy$R2_cond` and a MAE, MAPE and RMSE of `r myround(back_LR_accuracy$MAE, 1)` N, `r myround(back_LR_accuracy$MAPE * 100, 1)`\% and `r myround(back_LR_accuracy$RMSE, 1)` N, respectively.
+The pLR equation with the lower prediction error was the one for accelerometers worn at hip and the resultant vector, with an R^2^ of `r hip_LR_accuracy$R2_cond` and a MAE, MAPE and RMSE of `r myround(hip_LR_accuracy$MAE, 1)` N, `r myround(hip_LR_accuracy$MAPE * 100, 1)`\% and `r myround(hip_LR_accuracy$RMSE, 1)` N, respectively.
 
@@
 Prediction equations for pLR showed a more accentuated dispersion, and again mainly in the higher magnitudes (Supplemental Figure S4).
-The pLR prediction model based on data from accelerometers placed at the ankle and the resultant vector displayed a bias different than zero (*p* = `r myround(ba_bias_test_res_LR$ankle["p.value"], 2)`).
-None of the other models presented a bias different than zero (*p* > 0.05), demonstrating that they do not consistently over or underestimate the pGRF and pLR values.
-All equations based on ankle-worn accelerometer data, for both outcomes (pGRF and pLR) and vectors (resultant and vertical) showed a significant proportional bias (*p* < 0.05).
-Despite the statistical significance, linear regressions presented a low R^2^ (highest R^2^ = `r myround(max(ba_regression_ankle_R2), 2)`).
+None of the models presented a bias different than zero (*p* > 0.05), demonstrating that they do not consistently over or underestimate the pGRF and pLR values.
+Although all equations showed a significant proportional bias (*p* < 0.05), the linear regressions presented a low R^2^, with the highest value being `r myround(max(ba_regression_ankle_R2), 2)`.
 
@@
 The results of these comparisons can be observed in the Table 2.
-Accuracy indices of our equations were all better compared to the reference equation, with a MAPE approximately 2 to 3 times lower.
+Accuracy indices of our equations were all better compared to the reference equation.
 Furthermore, our equation performance was constant between walking and running, indicating that it can be used for both activities without loss of accuracy, which cannot be observed in the reference equation.

### 2021-05-15 figures/fig2.R
Add alpha to the geom_point()

@@ -23,7 +23,7 @@ scatterplot_GRF_res_hip <- mechanical_load_data %>%
   filter(vector == "resultant" & acc_placement == "hip") %>%
   ggplot() +
   geom_point(
-    aes(x = pACC_g, y = pGRF_N, color = BMI_cat, shape = activity)
+    aes(x = pACC_g, y = pGRF_N, color = BMI_cat, shape = activity), alpha = 0.5
   ) +
   geom_smooth(
     aes(x = pACC_g, y = pGRF_N, color = BMI_cat),
@@ -66,7 +66,7 @@ scatterplot_LR_res_hip <- mechanical_load_data %>%
   filter(vector == "resultant" & acc_placement == "hip") %>%
   ggplot() +
   geom_point(
-    aes(x = pAR_gs, y = pLR_Ns, color = BMI_cat, shape = activity)
+    aes(x = pAR_gs, y = pLR_Ns, color = BMI_cat, shape = activity), alpha = 0.5
   ) +
   geom_smooth(
     aes(x = pAR_gs, y = pLR_Ns, color = BMI_cat),

### 2021-05-15 figures/fig3.R

@@ -34,7 +34,7 @@ cv_res_LR_models <- map(
 # Hip pGRF Bland-Altman plot ---------------------------------------------
 
 BA_GRF_res_hip <- cv_res_GRF_models$hip %>%
-  plot_bland_altman(color = BMI_cat, shape = activity) +
+  plot_bland_altman(color = BMI_cat, shape = activity, alpha = 0.5) +
   scale_color_nejm() +
   scale_y_continuous(
     limits = c(-600, 600),
@@ -69,7 +69,7 @@ BA_GRF_res_hip <- cv_res_GRF_models$hip %>%
 # Hip pLR Bland-Altman plot ----------------------------------------------
 
 BA_LR_res_hip <- cv_res_LR_models$hip %>%
-  plot_bland_altman(color = BMI_cat, shape = activity) +
+  plot_bland_altman(color = BMI_cat, shape = activity, alpha = 0.5) +
   scale_color_nejm() +
   scale_y_continuous(
     labels = scales::label_number(),

### 2021-05-15 figures/figS1.R

@@ -24,7 +24,7 @@ scatterplot_GRF_res_ankle <- mechanical_load_data %>%
   filter(vector == "resultant" & acc_placement == "ankle") %>%
   ggplot() +
   geom_point(
-    aes(x = pACC_g, y = pGRF_N, color = BMI_cat, shape = activity)
+    aes(x = pACC_g, y = pGRF_N, color = BMI_cat, shape = activity), alpha = 0.5
   ) +
   geom_smooth(
     aes(x = pACC_g, y = pGRF_N, color = BMI_cat),
@@ -66,7 +66,7 @@ scatterplot_GRF_res_back <- mechanical_load_data %>%
   filter(vector == "resultant" & acc_placement == "lower_back") %>%
   ggplot() +
   geom_point(
-    aes(x = pACC_g, y = pGRF_N, color = BMI_cat, shape = activity)
+    aes(x = pACC_g, y = pGRF_N, color = BMI_cat, shape = activity), alpha = 0.5
   ) +
   geom_smooth(
     aes(x = pACC_g, y = pGRF_N, color = BMI_cat),
@@ -108,7 +108,7 @@ scatterplot_GRF_res_hip <- mechanical_load_data %>%
   filter(vector == "resultant" & acc_placement == "hip") %>%
   ggplot() +
   geom_point(
-    aes(x = pACC_g, y = pGRF_N, color = BMI_cat, shape = activity)
+    aes(x = pACC_g, y = pGRF_N, color = BMI_cat, shape = activity), alpha = 0.5
   ) +
   geom_smooth(
     aes(x = pACC_g, y = pGRF_N, color = BMI_cat),
@@ -150,7 +150,7 @@ scatterplot_GRF_ver_ankle <- mechanical_load_data %>%
   filter(vector == "vertical" & acc_placement == "ankle") %>%
   ggplot() +
   geom_point(
-    aes(x = pACC_g, y = pGRF_N, color = BMI_cat, shape = activity)
+    aes(x = pACC_g, y = pGRF_N, color = BMI_cat, shape = activity), alpha = 0.5
   ) +
   geom_smooth(
     aes(x = pACC_g, y = pGRF_N, color = BMI_cat),
@@ -192,7 +192,7 @@ scatterplot_GRF_ver_back <- mechanical_load_data %>%
   filter(vector == "vertical" & acc_placement == "lower_back") %>%
   ggplot() +
   geom_point(
[…]

### 2021-05-15 figures/figS2.R

@@ -24,7 +24,7 @@ scatterplot_LR_res_ankle <- mechanical_load_data %>%
   filter(vector == "resultant" & acc_placement == "ankle") %>%
   ggplot() +
   geom_point(
-    aes(x = pAR_gs, y = pLR_Ns, color = BMI_cat, shape = activity)
+    aes(x = pAR_gs, y = pLR_Ns, color = BMI_cat, shape = activity), alpha = 0.5
   ) +
   geom_smooth(
     aes(x = pAR_gs, y = pLR_Ns, color = BMI_cat),
@@ -66,7 +66,7 @@ scatterplot_LR_res_back <- mechanical_load_data %>%
   filter(vector == "resultant" & acc_placement == "lower_back") %>%
   ggplot() +
   geom_point(
-    aes(x = pAR_gs, y = pLR_Ns, color = BMI_cat, shape = activity)
+    aes(x = pAR_gs, y = pLR_Ns, color = BMI_cat, shape = activity), alpha = 0.5
   ) +
   geom_smooth(
     aes(x = pAR_gs, y = pLR_Ns, color = BMI_cat),
@@ -108,7 +108,7 @@ scatterplot_LR_res_hip <- mechanical_load_data %>%
   filter(vector == "resultant" & acc_placement == "hip") %>%
   ggplot() +
   geom_point(
-    aes(x = pAR_gs, y = pLR_Ns, color = BMI_cat, shape = activity)
+    aes(x = pAR_gs, y = pLR_Ns, color = BMI_cat, shape = activity), alpha = 0.5
   ) +
   geom_smooth(
     aes(x = pAR_gs, y = pLR_Ns, color = BMI_cat),
@@ -150,7 +150,7 @@ scatterplot_LR_ver_ankle <- mechanical_load_data %>%
   filter(vector == "vertical" & acc_placement == "ankle") %>%
   ggplot() +
   geom_point(
-    aes(x = pAR_gs, y = pLR_Ns, color = BMI_cat, shape = activity)
+    aes(x = pAR_gs, y = pLR_Ns, color = BMI_cat, shape = activity), alpha = 0.5
   ) +
   geom_smooth(
     aes(x = pAR_gs, y = pLR_Ns, color = BMI_cat),
@@ -192,7 +192,7 @@ scatterplot_LR_ver_back <- mechanical_load_data %>%
   filter(vector == "vertical" & acc_placement == "lower_back") %>%
[…]

### 2021-05-15 figures/figS3.R

@@ -35,7 +35,7 @@ cv_ver_GRF_models <- map(
 
 # Resultant: Ankle
 BA_GRF_res_ankle <- cv_res_GRF_models$ankle %>%
-  plot_bland_altman(color = BMI_cat, shape = activity) +
+  plot_bland_altman(color = BMI_cat, shape = activity, alpha = 0.5) +
   scale_color_nejm() +
   scale_y_continuous(
     limits = c(-600, 600),
@@ -69,7 +69,7 @@ BA_GRF_res_ankle <- cv_res_GRF_models$ankle %>%
 
 # Resultant: Lower back
 BA_GRF_res_back <- cv_res_GRF_models$lower_back %>%
-  plot_bland_altman(color = BMI_cat, shape = activity) +
+  plot_bland_altman(color = BMI_cat, shape = activity, alpha = 0.5) +
   scale_color_nejm() +
   scale_y_continuous(
     limits = c(-600, 600),
@@ -103,7 +103,7 @@ BA_GRF_res_back <- cv_res_GRF_models$lower_back %>%
 
 # Resultant: Hip
 BA_GRF_res_hip <- cv_res_GRF_models$hip %>%
-  plot_bland_altman(color = BMI_cat, shape = activity) +
+  plot_bland_altman(color = BMI_cat, shape = activity, alpha = 0.5) +
   scale_color_nejm() +
   scale_y_continuous(
     limits = c(-600, 600),
@@ -137,7 +137,7 @@ BA_GRF_res_hip <- cv_res_GRF_models$hip %>%
 
 # Vertical: Ankle
 BA_GRF_ver_ankle <- cv_ver_GRF_models$ankle %>%
-  plot_bland_altman(color = BMI_cat, shape = activity) +
+  plot_bland_altman(color = BMI_cat, shape = activity, alpha = 0.5) +
   scale_color_nejm() +
   scale_y_continuous(
     limits = c(-600, 600),
@@ -171,7 +171,7 @@ BA_GRF_ver_ankle <- cv_ver_GRF_models$ankle %>%
 
 # Vertical: Lower back
 BA_GRF_ver_back <- cv_ver_GRF_models$lower_back %>%
-  plot_bland_altman(color = BMI_cat, shape = activity) +
+  plot_bland_altman(color = BMI_cat, shape = activity, alpha = 0.5) +
   scale_color_nejm() +
   scale_y_continuous(
     limits = c(-600, 600),
[…]

### 2021-05-15 figures/figS4.R

@@ -35,7 +35,7 @@ cv_ver_LR_models <- map(
 
 # Resultant: Ankle
 BA_LR_res_ankle <- cv_res_LR_models$ankle %>%
-  plot_bland_altman(color = BMI_cat, shape = activity) +
+  plot_bland_altman(color = BMI_cat, shape = activity, alpha = 0.5) +
   scale_color_nejm() +
   scale_y_continuous(
     labels = scales::label_number(),
@@ -70,7 +70,7 @@ BA_LR_res_ankle <- cv_res_LR_models$ankle %>%
 
 # Resultant: Lower back
 BA_LR_res_back <- cv_res_LR_models$lower_back %>%
-  plot_bland_altman(color = BMI_cat, shape = activity) +
+  plot_bland_altman(color = BMI_cat, shape = activity, alpha = 0.5) +
   scale_color_nejm() +
   scale_y_continuous(
     labels = scales::label_number(),
@@ -105,7 +105,7 @@ BA_LR_res_back <- cv_res_LR_models$lower_back %>%
 
 # Resultant: Hip
 BA_LR_res_hip <- cv_res_LR_models$hip %>%
-  plot_bland_altman(color = BMI_cat, shape = activity) +
+  plot_bland_altman(color = BMI_cat, shape = activity, alpha = 0.5) +
   scale_color_nejm() +
   scale_y_continuous(
     labels = scales::label_number(),
@@ -140,7 +140,7 @@ BA_LR_res_hip <- cv_res_LR_models$hip %>%
 
 # Vertical: Ankle
 BA_LR_ver_ankle <- cv_ver_LR_models$ankle %>%
-  plot_bland_altman(color = BMI_cat, shape = activity) +
+  plot_bland_altman(color = BMI_cat, shape = activity, alpha = 0.5) +
   scale_color_nejm() +
   scale_y_continuous(
     labels = scales::label_number(),
@@ -175,7 +175,7 @@ BA_LR_ver_ankle <- cv_ver_LR_models$ankle %>%
 
 # Vertical: Lower back
 BA_LR_ver_back <- cv_ver_LR_models$lower_back %>%
-  plot_bland_altman(color = BMI_cat, shape = activity) +
+  plot_bland_altman(color = BMI_cat, shape = activity, alpha = 0.5) +
   scale_color_nejm() +
   scale_y_continuous(
     labels = scales::label_number(),
[…]

### 2021-05-15 manuscript/manuscript.Rmd
Minor corrections

@@
 **Purpose:** To develop accelerometry-based equations to predict peak ground reaction forces (pGRF) and peak loading rate (pLR) during walking and running.
-**Methods:** `r Words(max(sample_size$pGRF_N))` subjects (`r unname(sample_size_sex["F"])` females; `r sample_descriptives$body_mass_mean`±`r sample_descriptives$body_mass_sd` kg) walked and ran at different speeds (2–14km·h^-1^) on a force plate–equipped treadmill while wearing accelerometers at ankle, lower back and hip.
+**Methods:** `r max(sample_size$pGRF_N)` subjects (`r unname(sample_size_sex["F"])` females; `r sample_descriptives$body_mass_mean`±`r sample_descriptives$body_mass_sd` kg) walked and ran at different speeds (2–14km·h^-1^) on a force plate–equipped treadmill while wearing accelerometers at ankle, lower back and hip.
 Regression equations were developed to predict pGRF and pLR from accelerometry data.
@@
 **Results:** Body mass and peak acceleration were included for pGRF prediction and body mass and peak acceleration rate for pLR prediction.
-All pGRF equation coefficients of determination were above `r myround(max(GRF_accuracy$R2_cond), 2)`, and a good agreement between actual and predicted pGRF was observed, with a mean absolute percent error (MAPE) below `r myround(max(GRF_accuracy$MAPE) * 100, 1)`\%.
-Accuracy indices from our equations were better than previously developed equations, with a MAPE 2 to 3 times better.
+All pGRF equation coefficients of determination were above `r myround(min(GRF_accuracy$R2_cond), 2)`, and a good agreement between actual and predicted pGRF was observed, with a mean absolute percent error (MAPE) below `r myround(max(GRF_accuracy$MAPE) * 100, 1)`\%.
+Accuracy indices from our equations were better than previously developed equations.
 All pLR prediction equations presented a lower accuracy compared to those developed to predict pGRF.
@@
 
-A convenience sample of `r max(sample_size$pGRF_N)` adults (`r unname(sample_size_sex["M"])` males; `r sample_descriptives$age_mean`±`r sample_descriptives$age_sd` years; `r sample_descriptives$height_mean`±`r sample_descriptives$height_sd` cm; `r sample_descriptives$body_mass_mean`±`r sample_descriptives$body_mass_sd` kg; `r sample_descriptives$BMI_mean`±`r sample_descriptives$BMI_sd` kg·m^-2^, X̄±SD) with no self-reported neurological or musculoskeletal limitations was recruited.
+A convenience sample of `r max(sample_size$pGRF_N)` adults (`r unname(sample_size_sex["F"])` males; `r sample_descriptives$age_mean`±`r sample_descriptives$age_sd` years; `r sample_descriptives$height_mean`±`r sample_descriptives$height_sd` cm; `r sample_descriptives$body_mass_mean`±`r sample_descriptives$body_mass_sd` kg; `r sample_descriptives$BMI_mean`±`r sample_descriptives$BMI_sd` kg·m^-2^, X̄±SD) with no self-reported neurological or musculoskeletal limitations was recruited.
 Height and body mass were assessed following standard procedures with a stadiometer and a digital scale, respectively.

### 2021-05-29 manuscript/manuscript.Rmd
Add HF suggestions

@@
 **Purpose:** To develop accelerometry-based equations to predict peak ground reaction forces (pGRF) and peak loading rate (pLR) during walking and running.
-**Methods:** `r max(sample_size$pGRF_N)` subjects (`r unname(sample_size_sex["F"])` females; `r sample_descriptives$body_mass_mean`±`r sample_descriptives$body_mass_sd` kg) walked and ran at different speeds (2–14km·h^-1^) on a force plate–equipped treadmill while wearing accelerometers at ankle, lower back and hip.
+**Methods:** `r max(sample_size$pGRF_N)` subjects (`r unname(sample_size_sex["F"])` females; `r sample_descriptives$body_mass_mean`±`r sample_descriptives$body_mass_sd`kg) walked and ran at different speeds (2–14km·h^-1^) on a force plate–equipped treadmill while wearing accelerometers at their ankle, lower back and hip.
 Regression equations were developed to predict pGRF and pLR from accelerometry data.
@@
 
-Physical activity-related mechanical loading plays a key role in bone health, by promoting positive effects on bone mass, geometry and microarchitecture [@Turner2003; @Fonseca2014].
+Physical activity-related mechanical loading plays a key role in bone health, by promoting favorable adaptations on bone mass, geometry and microarchitecture [@Turner2003; @Fonseca2014].
 However, not all forms of mechanical stimulation are equally osteogenic, and the manipulation of variables such as loading magnitude, rate, frequency and volume can produce different bone responses [@Turner1998; @Turner2003].
 Therefore, the correct assessment of mechanical loading is of paramount importance to better understand the bone response to physical stimulus.
-Nonetheless, a limitation in this context is that the mechanical loading evaluation is usually performed only by instruments restrained to laboratory settings.
+Nonetheless, a limitation in this context is that mechanical loading evaluation is usually performed only by instruments restrained to laboratory settings.
 
@@
 
-The ground reaction force (GRF) is arguably the most studied mechanical loading variable, and its relationship with the bone adaptations to exercise is well established [@Turner2003; @Rantalainen2008; @Rantalainen2009].
-Due to its importance, some studies have already attempted to develop peak GRF (pGRF) prediction equations based on accelerometry data in several populations, such as children [@Neugebauer2012], adults [@Fortune2014; @Neugebauer2014], patients with obesity [@Veras2020] and military [@Neugebauer2018], but with the majority of these studies including only activities eliciting a low loading magnitude.
+The ground reaction force (GRF) is arguably the most studied mechanical loading variable, and its relationship with bone adaptations to exercise is well established [@Turner2003; @Rantalainen2008; @Rantalainen2009].
+Due to its importance, some studies have already attempted to develop peak GRF (pGRF) prediction equations based on accelerometry data in several populations, such as children [@Neugebauer2012], adults [@Fortune2014; @Neugebauer2014], patients with obesity [@Veras2020] and military personal [@Neugebauer2018], but with the majority of these studies including only activities eliciting a low loading magnitude.
 Besides loading magnitude, loading rate (LR) is another variable related with bone health [@Rowlands2014] that could potentially be predicted by accelerometer data, but with apparently lower prediction accuracy compared with pGRF [@Fortune2014; @Veras2020].
@@
 A convenience sample of `r max(sample_size$pGRF_N)` adults (`r unname(sample_size_sex["F"])` males; `r sample_descriptives$age_mean`±`r sample_descriptives$age_sd` years; `r sample_descriptives$height_mean`±`r sample_descriptives$height_sd` cm; `r sample_descriptives$body_mass_mean`±`r sample_descriptives$body_mass_sd` kg; `r sample_descriptives$BMI_mean`±`r sample_descriptives$BMI_sd` kg·m^-2^, X̄±SD) with no self-reported neurological or musculoskeletal limitations was recruited.
-Height and body mass were assessed following standard procedures with a stadiometer and a digital scale, respectively.
+Height and body mass were assessed following standard procedures with a stadiometer (model 213; Seca, Hamburg, Germany) and a digital scale (model 899; Seca, Hamburg, Germany), respectively.
 All participants were informed about the experiments purpose and protocol before giving written informed consent.
-The study was approved by the local Ethics Committee (CES 192-14).
+The study protocol was approved by the local Ethics Committee (CES 192-14).
 
-## Protocol {.unnumbered}
+## Data collection protocol {.unnumbered}
 
@@
 Activity monitors used were the GT9X Link model (±16*g* range; ActiGraph, Pensacola, USA) that incorporates primary and secondary triaxial accelerometers.
-As a manufacturer proprietary filter is applied on the primary, but not on the secondary accelerometer raw data, and therefore only data from the latter was used in this study, reducing filtering bias and enhancing the data processing replicability.
+As a manufacturer proprietary filter is applied on the primary, but not on the secondary accelerometer raw data, only data from the latter was used in this study, reducing filtering bias and enhancing the data processing replicability.
 
@@
 Peak GRF (pGRF) was determined as the highest value between 0.4 seconds before and after each pACC.
-The rates of change, LR for the GRF and ACC rate (AR) for the ACC, were computed through a centered derivative from the beginning of the foot contact to the pGRF.
-The pLR and peak AR (pAR) were defined as the maximum value in the array.
+The rates of change were computed through a centered derivative from the beginning of the foot contact to the pGRF.
+The peak loading rate (pLR) and peak acceleration rate (pAR) were defined as the maximum value in the array.
 
@@
 Regions of interest for data analysis were manually selected where locomotion patterns were constant, which corresponded of 20-45 seconds periods, on average.
-Then, pGRF, pACC, pLR and pAR means of the resultant and its vertical component for each participant at each speed were extracted and used in all remaining analyses.
+Then, pGRF, pACC, pLR and pAR means of the resultant and its vertical component for each participant at each speed and for each accelerometer placement were extracted and used in all remaining analyses.
 In walking trials, a median of `r n_peaks_desc[which(n_peaks_desc$locomotion_type == "walking"), "n_peaks_median"]` peaks (interquartile range `r n_peaks_desc[which(n_peaks_desc$locomotion_type == "walking"), "n_peaks_median"] - n_peaks_desc[which(n_peaks_desc$locomotion_type == "walking"), "n_peaks_iqr"]`\textendash`r n_peaks_desc[which(n_peaks_desc$locomotion_type == "walking"), "n_peaks_median"] + n_peaks_desc[which(n_peaks_desc$locomotion_type == "walking"), "n_peaks_iqr"]`) were used to calculate these means, while in running trials it was a median of `r n_peaks_desc[which(n_peaks_desc$locomotion_type == "running"), "n_peaks_median"]` peaks (interquartile range `r n_peaks_desc[which(n_peaks_desc$locomotion_type == "running"), "n_peaks_median"] - n_peaks_desc[which(n_peaks_desc$locomotion_type == "running"), "n_peaks_iqr"]`\textendash`r n_peaks_desc[which(n_peaks_desc$locomotion_type == "running"), "n_peaks_median"] + n_peaks_desc[which(n_peaks_desc$locomotion_type == "running"), "n_peaks_iqr"]`).
@@
 
-Our recruited sample presented a broad range of body masses, from normal weight to severe obesity.
-It would be expected that participants with higher body mass index (BMI) values, due to low motor ability and physical capacity, are able to perform the trials only at the lowest speeds.
+Our recruited sample presented a broad range of body masses (`r myround(min(mechanical_load_data$body_mass))`-`r myround(max(mechanical_load_data$body_mass))`kg; `r myround(min(mechanical_load_data$BMI))`-`r myround(max(mechanical_load_data$BMI))`kg·m^-2^), from normal weight to severe obesity.
+It was expected that participants with higher body mass index (BMI) values, due to their usually lower motor ability and physical capacity, would be able to perform the trials only at the lowest speeds.
 This was precisely what happened and, as can be observed in Figure 1, the number of data points corresponding to subjects with BMI > 30 kg·m^-2^ greatly decreases in the speeds above 6 km·h^-1^.
@@
 During locomotion, pACC increases linearly as pGRF increases.
-Such relation can be observed in the Figure 2, Panel A, for from hip-worn accelerometers and the resultant vector, and in Supplemental Figure S1 for all placements and the resultant vector and its vertical component.
+Such relation can be observed in Figure 2, Panel A, for hip-worn accelerometers and the resultant vector, and in Supplemental Figure S1 for all placements and the resultant vector and its vertical component.
 Moreover, for the same pACC values, the registered pGRF tended to be higher in subjects with higher BMI.
-It demonstrates that accelerometer data can adequately discriminate between different BMI categories.
+This demonstrates that accelerometer data can adequately discriminate between different BMI categories.
 However, data from the pAR x pLR relationship show a greater dispersion, especially at higher values, and the different BMI categories cannot be discriminated so well, as can be observed in Figure 2, Panel B, and Supplemental Figure S2.
@@
 Regression equations were developed to predict the mechanical loading variables based on accelerometry data.
-In the pGRF prediction equations, body mass and pACC were significant predictors and included in the models.
+In the pGRF prediction equations, body mass and pACC were significant predictors and  therefore were included in the models.
 Table 1 shows the regression coefficients, R^2^ and accuracy indices from models derived from accelerometers placed at ankle, lower back and hip, and from the resultant vector and its vertical component.
 In the pGRF prediction models, the lowest R^2^ value found was `r min(GRF_accuracy$R2_cond)`, which shows that the predictor variables are able to explain the variability in the outcome variable to a great extent.
-Also, MAE ranged between `r myround(min(GRF_accuracy$MAE), 1)` N and `r myround(max(GRF_accuracy$MAE), 1)` N, and the MAPE from `r myround(min(GRF_accuracy$MAPE) * 100, 1)`\% to `r myround(max(GRF_accuracy$MAPE) * 100, 1)`\%, which is a smaller error than previously published equations [@Neugebauer2014; @Fortune2014], and similar to the error from equations previously published by our group, but which included walking only [@Veras2020].
+Also, MAE ranged between `r myround(min(GRF_accuracy$MAE), 1)` N and `r myround(max(GRF_accuracy$MAE), 1)` N, and the MAPE from `r myround(min(GRF_accuracy$MAPE) * 100, 1)`\% to `r myround(max(GRF_accuracy$MAPE) * 100, 1)`\%, which is a smaller error than in previously published equations [@Neugebauer2014; @Fortune2014], and similar to the error from equations previously published by our group for walking only [@Veras2020].
 RMSE were also computed and shown to range between `r myround(min(GRF_accuracy$RMSE), 1)` N and `r myround(max(GRF_accuracy$RMSE), 1)` N.
-Although all pGRF prediction equations have shown very similar results, the one based on hip-worn accelerometers and the resultant vector has shown the best performance, with an R^2^ of `r hip_GRF_accuracy$R2_cond`, and the lowest MAE (`r myround(hip_GRF_accuracy$MAE, 1)` N), MAPE (`r myround(hip_GRF_accuracy$MAPE * 100, 1)`\%) and RMSE (`r myround(hip_GRF_accuracy$RMSE, 1)` N).
+Although all pGRF prediction equations showed very similar results, the one based on hip-worn accelerometers and the resultant vector has shown the best performance, with an R^2^ of `r hip_GRF_accuracy$R2_cond`, the lowest MAE (`r myround(hip_GRF_accuracy$MAE, 1)` N), MAPE (`r myround(hip_GRF_accuracy$MAPE * 100, 1)`\%) and RMSE (`r myround(hip_GRF_accuracy$RMSE, 1)` N).
 
@@
 
-Moreover, pLR prediction equations were developed, and these included as predictors the pAR and again body mass.
+Moreover, pLR prediction equations were also developed, and these included as predictors the pAR and again body mass.
 Regression coefficients, R^2^ and accuracy indices for these equations are also shown in Table 1.
 For these equations, the lowest R^2^ found was `r myround(min(LR_accuracy$R2_cond), 2)` and the highest was `r myround(max(LR_accuracy$R2_cond), 2)`, reflecting a moderate to high prediction ability.
-MAE, MAPE and RMSE values ranged from `r myround(min(LR_accuracy$MAE), 1)` N to `r myround(max(LR_accuracy$MAE), 1)` N, from `r myround(min(LR_accuracy$MAPE) * 100, 1)`\% to `r myround(max(LR_accuracy$MAPE) * 100, 1)`\%, and from `r myround(min(LR_accuracy$RMSE), 1)` N to `r myround(max(LR_accuracy$RMSE), 1)` N, respectively.
-The results found for the pLR prediction equations were all considerably poorer than those for the pGRF equations.
-The accuracy indices values were also somewhat worse than previously published pLR prediction equations [@Veras2020; @Fortune2014], although equations from both studies were only validate for walking.
-The pLR equation with the lower prediction error was the one for accelerometers worn at hip and the resultant vector, with an R^2^ of `r hip_LR_accuracy$R2_cond` and a MAE, MAPE and RMSE of `r myround(hip_LR_accuracy$MAE, 1)` N, `r myround(hip_LR_accuracy$MAPE * 100, 1)`\% and `r myround(hip_LR_accuracy$RMSE, 1)` N, respectively.
+MAE, MAPE and RMSE values ranged from `r round(min(LR_accuracy$MAE), 0)` N to `r round(max(LR_accuracy$MAE), 0)` N, from `r myround(min(LR_accuracy$MAPE) * 100, 1)`\% to `r myround(max(LR_accuracy$MAPE) * 100, 1)`\%, and from `r round(min(LR_accuracy$RMSE), 0)` N to `r round(max(LR_accuracy$RMSE), 0)` N, respectively.
+These results show that the pLR prediction equations were all considerably poorer than those for the pGRF equations.
+The accuracy indices values were also somewhat worse than previously published pLR prediction equations [@Veras2020; @Fortune2014], although equations from both studies were only valid for walking activities.
+The pLR equation with the lowest prediction error was the one derived from accelerometers worn at hip and the resultant vector, with an R^2^ of `r hip_LR_accuracy$R2_cond` and a MAE, MAPE and RMSE of `r round(hip_LR_accuracy$MAE, 0)` N, `r myround(hip_LR_accuracy$MAPE * 100, 1)`\% and `r round(hip_LR_accuracy$RMSE, 0)` N, respectively.
 
@@
 Values for all pGRF and pLR prediction equations developed tended to assemble around zero and between the limits of agreement.
-For the pGRF equations, a tendency to higher dispersion can be observed in the higher magnitudes, which correspond to the running trials, and especially in values derived from ankle-worn accelerometers (Supplemental Figure S3, Panels A and D).
-Prediction equations for pLR showed a more accentuated dispersion, and again mainly in the higher magnitudes (Supplemental Figure S4).
-None of the models presented a bias different than zero (*p* > 0.05), demonstrating that they do not consistently over or underestimate the pGRF and pLR values.
+For the pGRF equations, a tendency to higher dispersion can be observed for higher magnitudes, which correspond to the running trials, and especially in values derived from ankle-worn accelerometers (Supplemental Figure S3, Panels A and D).
+Prediction equations for pLR showed a more accentuated dispersion, and again mainly for higher magnitudes (Supplemental Figure S4).
+None of the models presented a bias different from zero (*p* > 0.05), demonstrating that they do not consistently over or underestimate the pGRF and pLR values.
 Although all equations showed a significant proportional bias (*p* < 0.05), the linear regressions presented a low R^2^, with the highest value being `r myround(max(ba_regression_ankle_R2), 2)`.
@@
 Then, these indices were compared to the indices for our equation (Table 1).
-Apart from the overall equations performance, it was done a sub-analysis separating the walking and running trials.
+Apart from the overall equations performance, a sub-analysis was done separating walking and running trials.
 The results of these comparisons can be observed in the Table 2.
@@
 Included in this package are functionalities to read raw ACC data, process it and apply the equations to predict both pGRF and pLR.
-In the package website (ADDRESS) there are instructions on how to use it to obtain the predictions, namely in the 'Home' and 'Get started' sections.
+In the package website (ADDRESS) there are also instructions on how to use it to obtain the predictions, namely in the 'Home' and 'Get started' sections.
 
@@
 
-The purpose of this study was to develop pGRF and pLR prediction equations based on accelerometer data able to accurately estimate the mechanical loading during walking and running.
-The predictions were tested with accelerometers placed at different body locations and using both the resultant vector and its vertical component in a set of incremental speeds, ranging from slow walking to running.
-Our findings demonstrated that accelerometer-based regression models were a valid and accurate method to estimate pGRF during human locomotion, as all of the pGRF prediction models developed presented a MAPE below the 8\% threshold established by us.
+The purpose of this study was to develop pGRF and pLR prediction equations based on accelerometer data able to accurately estimate mechanical loading during walking and running.
+Predictions were tested with accelerometers placed at different body locations, using both the resultant vector and its vertical component in a set of incremental speeds, ranging from slow walking to running.
+Our findings demonstrated that accelerometer-based regression models were a valid and accurate method to estimate pGRF during human locomotion, as all of the pGRF prediction models developed presented a MAPE below the 8\% threshold.
 However, accelerometry-based equations developed to predict pLR did not yield trustworthy results.
@@
 Additionally, previous studies that developed pGRF prediction models during both walking and running showed a MAPE of 8.3\% [@Neugebauer2012] and 5.2\% [@Neugebauer2014], which are higher or similar to our results, respectively.
-Nevertheless, models from both studies present a caveat, that is the inclusion of a term in their equations that distinguishes whether the activity is walking or running, which is not easy to be identify from free-living accelerometer data, hindering the application of these equations.
-Also, one of these studies included in the sample only children and adolescents [@Neugebauer2012].
-It is important to mention that the prediction error found in our pGRF models is lower than the error typically found in models developed to estimate energy expenditure [@Diniz-Sousa2020], which is one of the main uses of accelerometer data [@Migueles2017].
+Nevertheless, models from both studies present a caveat, that is the inclusion of a term in their equations that distinguishes whether the activity is walking or running, which is not easy to be identify from free-living accelerometer data, hindering the real-world application of these previous equations.
+Also, one of these studies included in their sample only children and adolescents [@Neugebauer2012].
+It is also important to mention that the prediction error found in our pGRF models is lower than the error typically found in models developed to estimate energy expenditure [@Diniz-Sousa2020], which is one of the main uses of accelerometer data [@Migueles2017].
 
@@
 
-Our findings revealed that either ankle, lower back and hip were valid accelerometer placements to accurately predict pGRF.
+Our findings revealed that ankle, lower back or hip were all valid accelerometer placements to accurately predict pGRF.
 However, despite the similarities, hip was the accelerometer placement with the best prediction accuracy, which reinforces it as the preferable location to collect and analyse accelerometry-measured physical activity data.
 This placement is already well accepted for energy expenditure prediction and physical activity intensity classification [@Migueles2017] and has also shown a high mechanical loading prediction ability in previous studies [@Veras2020; @Neugebauer2018; @Neugebauer2012].
-Besides the various placements, the current study also analysed predictions from both the resultant and vertical vectors.
-Although ACC and GRF from the vertical vector are the most studied and frequently associated with bone health [@Vainionpaa2006; @Heikkinen2007], the resultant vector of both ACC and GRF during locomotion is greatly determined by the vertical vector, and, as can be observed in Supplemental Figure S1, values from both vectors in this study were very similar.
-Also, mechanical loading prediction based on the resultant ACC offers the important advantage of not being dependent on the accelerometer correct orientation, which is hard to assure when measuring daily physical activity [@Welk2005; @Wundersitz2015].
+Besides various placements, the current study also analysed predictions from both the resultant and vertical vectors.
+Although ACC and GRF from the vertical vector are the most studied and frequently associated with bone health [@Vainionpaa2006; @Heikkinen2007], the resultant vector of both ACC and GRF during locomotion is greatly determined by the vertical vector, and, as can be observed in Supplemental Figure S1, in this study values from both vectors were very similar.
+Also, mechanical loading prediction based on the resultant ACC offers the important advantage of not being dependent on the accelerometer positioning and correct orientation, which is hard to assure when measuring daily physical activity [@Welk2005; @Wundersitz2015].
 
@@
 Accordingly to what we found, these studies also showed that the prediction error for these models is considerably larger than for the pGRF prediction.
-One of these researches investigated the use of four different accelerometer placements (ankle, tibia, thigh and waist) to predict pLR and achieved a MAPE ranging from 23.4\% to 46.1\% [@Fortune2014], the other one found values from 19.5\% to 22.4\% using accelerometers at the lower back and hip [@Veras2020].
-Results from the models developed in these investigations have both lower and higher prediction errors compared to our results, which are approximately of 29.0\%.
-Although pLR is an important variable, known for a long time to affect the bone response to mechanical loading [@Bassey1998; @Turner1995], based on this and previous studies [@Fortune2014; @Veras2020], its prediction through accelerometer data does not seem to yield trustworthy results.
+One of these researches investigated the use of four different accelerometer placements (ankle, tibia, thigh and waist) to predict pLR and achieved a MAPE ranging from 23.4\% to 46.1\% [@Fortune2014], while the other obtained values from 19.5\% to 22.4\% using accelerometers at the lower back and hip [@Veras2020].
+Results from the models developed in these previous investigations have both lower and higher prediction errors compared to our results, which were approximately of 29.0\%.
+Although pLR is an important variable, known for a long time to determine to a great extent the bone formation response to mechanical loading [@Bassey1998; @Turner1995], based on our findings and in previous studies [@Fortune2014; @Veras2020], its prediction through accelerometer data does not seem to yield trustworthy results.
 Nevertheless, past research has suggested that the bone response to pGRF and pLR is somewhat similar due to the high correlation between these variables [@Heikkinen2007].
@@
 
-Results from our study offer some advantages compared to previous investigations: i) it includes walking and running, which are among the most usual types of physical activity performed daily; ii) all models developed, especially for pGRF prediction, presented a similar or better accuracy compared to the results from previous publications; iii) the accuracy of our models is constant throughout walking and running, and therefore they can be used adequately for both activities; iv) in order to apply our models there is no need to distinguish the accelerometer data between walking and running, which facilitates its use with daily living accelerometer data; and v) along with the creation of these models, a software was developed in the form of an R package to perform all the computations needed to get from the raw accelerometer data to the mechanical loading prediction.
+Results from our study offer some advantages compared to previous investigations: i) it includes walking and running, which are among the most usual types of physical activity performed daily; ii) all models developed, especially for pGRF prediction, presented a similar or better accuracy compared to the results from previous publications; iii) the accuracy of our models is constant throughout walking and running, and therefore they can be used adequately for both activities; iv) in order to apply our models there is no need to manually distinguish the accelerometer data between walking and running, as in previously developed equations [@Neugebauer2014], which facilitates its use with daily living accelerometer data; and v) along with the creation of these models, a software was developed in the form of an R package to perform all the computations needed to get from the raw accelerometer data to the mechanical loading prediction.
 
@@
 For instance, these models can be used to analyse data from large observational studies and make associations between mechanical loading and bone health.
-But instead of analysing accelerometer data averaged in epochs similarly to some previous publications [@Chastin2014; @Stiles2017], which can dilute the higher intensities [@Fabre2020], with our method raw ACC can be analysed at a higher resolution by identifying the pGRF associated with each step taken and establishing, for example, the number of peaks above certain magnitude needed to elicit a positive bone adaptation.
+But instead of analysing accelerometer data averaged in epochs similarly to some previous publications [@Chastin2014; @Stiles2017], which can dilute the higher intensities [@Fabre2020], with our method raw ACC can be analysed at a higher resolution by identifying the pGRF associated with each step taken and establishing, for example, the number of peaks above a certain magnitude threshold needed to elicit a positive bone adaptation.
 Another potential application could be the computation of the cumulative mechanical loading by adding up the pGRF of all detected steps.
-This information could be used either for the determination of the relationship between the mechanical loading accumulated in a certain period (e.g., a day) and the bone response to this loading [@Ahola2010], or also in sports training, by monitoring the amount of loading beared by athletes and relate to the osteoarticular safety of the training regimen [@Vanwanseele2020].
+This information could be used either for the determination of the relationship between the mechanical loading accumulated in a certain period (e.g., a day) and the bone response to this loading [@Ahola2010], or also in sports training, by monitoring the amount of loading beared by athletes and relate this information with osteoarticular safety of the training regimen [@Vanwanseele2020].
 
@@
 Despite the validity of the prediction models developed here, some limitations should be mentioned.
-First, although the activities included in our protocol (walking and running) are among the most performed during daily living, our prediction models are not applicable to other activities known to influence the bone adaptation to mechanical stimuli, such as jumping and direction changing.
+First, although the activities included in our protocol (walking and running) are among the most frequently performed motor activities during daily living, our prediction models are not applicable to other activities known to influence bone adaptation to mechanical stimuli, such as jumping and direction changing.
 Second, in the absence of another sample for model validation, the leave-one-out cross-validation method was used, which is the method of choice in these situations [@Staudenmayer2012].
@@
 These models are valid for the use of both the resultant vector and its vertical component, and using data from accelerometers placed at either the ankle, lower back or hip.
-The prediction models developed in this study have proved to be an efficient method to assess pGRF out of laboratory conditions, with potential applications of monitoring the mechanical loading in several contexts and establishing the relationship of this loading and either a positive or negative osteoarticular adaptation.
+The prediction models developed in this study have proved to be an efficient method to assess pGRF out of laboratory conditions, with potential applications for monitoring the mechanical loading in several contexts and establishing a relationship between mechanican loading and either beneficial or pathological osteoarticular adaptations.
 This is also the first work, to the best of our knowledge, to explore the prediction of pLR based on accelerometry data during both walking and running, although the accuracy of these models were unsatisfactory.
@@
 
-This study was funded by the Foundation for Science and Technology of Portugal (FCT) (grant PTDC/DTP-DES/0968/2014) and by the European Regional Development Fund (ERDF) through the Operational Competitiveness Programme (COMPETE) (grant POCI-01-0145-FEDER-016707).
+This study was funded by the Fundação para a Ciência e a Tecnologia (FCT) (grant PTDC/DTP-DES/0968/2014) and by the European Regional Development Fund (ERDF) through the Operational Competitiveness Programme (COMPETE) (grant POCI-01-0145-FEDER-016707).
 The study was developed in the Research Centre in Physical Activity, Health and Leisure (CIAFEL) funded by ERDF through the COMPETE and by the FCT (grant UIDB/00617/2020).
-Lucas Veras was supported by the FCT (grant UI/BD/150673/2020), Florêncio Diniz-Sousa was supported by the FCT (grant SFRH/BD/117622/2016) and Giorjines Boppre was supported by the FCT (grant SFRH/BD/146976/2019).
+Lucas Veras, Florêncio Diniz-Sousa and Giorjines Bopper were supported by the FCT grants UI/BD/150673/2020, SFRH/BD/117622/2016, SFRH/BD/146976/2019, respectively.
 The authors thank the participants who took part in this research and all who have collaborated in the project.
@@
 \footnotesize Bland-Altman plots showing agreement between actual and predicted peak loading rate for accelerometers worn at ankle (Panels A and D), lower back (Panels B and E) and hip (Panels C and F), for both the resultant vector (Panels A to C) and its vertical component (Panels D to F).
-Continuous thick lines show bias (average of the differences between actual and predicted values) while dashed lines show the limits of agreement (bias \pm 1.96 standard deviation).
+Continuous thick lines show bias (average of the differences between actual and predicted values) while dashed lines show the limits of agreement (bias ± 1.96 standard deviation).
 Abbreviations: pLR, peak loading rate.

### 2021-05-31 manuscript/manuscript.Rmd
Minor corrections

@@
 Comparing our results with previous studies that used a similar approach to predict pGRF, it can be seen that prediction errors from our models are comparable with some studies whose best results presented a MAPE of approximately 6.0\% but which included only walking in their protocol [@Fortune2014; @Neugebauer2018; @Veras2020].
-Additionally, previous studies that developed pGRF prediction models during both walking and running showed a MAPE of 8.3\% [@Neugebauer2012] and 5.2\% [@Neugebauer2014], which are higher or similar to our results, respectively.
+Additionally, previous studies that developed pGRF prediction models during both walking and running showed a MAPE of 8.3\% [@Neugebauer2014], which is higher than our results, and 5.2\% [@Neugebauer2012] with no cross-validation, which is similar to our results when no cross-validation method is applied (data not shown).
 Nevertheless, models from both studies present a caveat, that is the inclusion of a term in their equations that distinguishes whether the activity is walking or running, which is not easy to be identify from free-living accelerometer data, hindering the real-world application of these previous equations.
@@
 One of these researches investigated the use of four different accelerometer placements (ankle, tibia, thigh and waist) to predict pLR and achieved a MAPE ranging from 23.4\% to 46.1\% [@Fortune2014], while the other obtained values from 19.5\% to 22.4\% using accelerometers at the lower back and hip [@Veras2020].
-Results from the models developed in these previous investigations have both lower and higher prediction errors compared to our results, which were approximately of 29.0\%.
+Results from the models developed in these previous investigations have both lower and higher prediction errors compared to our results, which were approximately of 35\%.
 Although pLR is an important variable, known for a long time to determine to a great extent the bone formation response to mechanical loading [@Bassey1998; @Turner1995], based on our findings and in previous studies [@Fortune2014; @Veras2020], its prediction through accelerometer data does not seem to yield trustworthy results.
-Nevertheless, past research has suggested that the bone response to pGRF and pLR is somewhat similar due to the high correlation between these variables [@Heikkinen2007].
+Nevertheless, past research has suggested that the bone response to pGRF and pLR is somewhat similar during walking, running and jumping activities due to the high correlation between these variables [@Heikkinen2007].
 Therefore, the pGRF prediction by accelerometry data, which is shown to be accurate, already incorporates some of the LR effect on the bone response to loading.
@@
 
-Results from our study offer some advantages compared to previous investigations: i) it includes walking and running, which are among the most usual types of physical activity performed daily; ii) all models developed, especially for pGRF prediction, presented a similar or better accuracy compared to the results from previous publications; iii) the accuracy of our models is constant throughout walking and running, and therefore they can be used adequately for both activities; iv) in order to apply our models there is no need to manually distinguish the accelerometer data between walking and running, as in previously developed equations [@Neugebauer2014], which facilitates its use with daily living accelerometer data; and v) along with the creation of these models, a software was developed in the form of an R package to perform all the computations needed to get from the raw accelerometer data to the mechanical loading prediction.
+Results from our study offer some advantages compared to previous investigations: i) it includes walking and running, which are among the most usual types of physical activity performed daily; ii) all models developed, especially for pGRF prediction, presented a similar or better accuracy compared to the results from previous publications; iii) the accuracy of our models is constant throughout walking and running, notably for the pGRF prediction models, and therefore they can be used adequately for both activities; iv) in order to apply our models there is no need to manually distinguish the accelerometer data between walking and running, as in previously developed equations [@Neugebauer2014], which facilitates its use with daily living accelerometer data; and v) along with the creation of these models, a software was developed in the form of an R package to perform all the computations needed to get from the raw accelerometer data to the mechanical loading prediction.
 

### 2021-06-03 tables/tab1.R
Round LR variables to 0 decimals

@@ -21,10 +21,12 @@ ver_GRF_tb <- build_formula_table(
 )
 res_LR_tb <- build_formula_table(
   res_LR_models, cv_res_LR_models, "pLR", "resultant"
-)
+) %>%
+  mutate(MAE = round(as.numeric(MAE), 0), RMSE = round(as.numeric(RMSE), 0))
 ver_LR_tb <- build_formula_table(
   ver_LR_models, cv_ver_LR_models, "pLR", "vertical"
-)
+) %>%
+  mutate(MAE = round(as.numeric(MAE), 0), RMSE = round(as.numeric(RMSE), 0))
 
 tab1_df <- rbind(res_GRF_tb, ver_GRF_tb, res_LR_tb, ver_LR_tb)
 # Remove vector name duplicates

### 2021-06-03 manuscript/manuscript.Rmd
Add more HF suggestions

@@
   as_vector()
+```
+
+```{r ba_agreement}
+load(here("output", "loocv_data.rda"))
+percent_inside_loa <- function(data) {
+  lower <- lvmisc::loa(
+    data[[".actual"]], data[[".predicted"]], na.rm = TRUE
+  )$lower
+  upper <- lvmisc::loa(
+    data[[".actual"]], data[[".predicted"]], na.rm = TRUE
+  )$upper
+  diff <- dplyr::transmute(
+    data, diff = .data[[".actual"]] - .data[[".predicted"]]
+  )
+  diff <- dplyr::mutate(
+    diff,
+    inside = as.factor(ifelse(diff > lower & diff < upper, "yes", "no"))
+  )
+  round(prop.table(table(diff$inside))[[2]] * 100, 0)
+}
+
+ba_agreement <- min(
+  map_dbl(cv_res_GRF_models, percent_inside_loa) %>% min(),
+  map_dbl(cv_ver_GRF_models, percent_inside_loa) %>% min(),
+  map_dbl(cv_res_LR_models, percent_inside_loa) %>% min(),
+  map_dbl(cv_ver_LR_models, percent_inside_loa) %>% min()
+)
 ```
@@
 Bland-Altman plots for the pGRF and pLR resultant vector from hip-worn accelerometers can be seen in Figure 3, and for all accelerometer placements and both resultant and vertical vectors in the Supplemental Figures S3 and S4.
-Values for all pGRF and pLR prediction equations developed tended to assemble around zero and between the limits of agreement.
+Values for all pGRF and pLR prediction equations developed tended to assemble around zero and between the limits of agreement, as shown by the fact that, for all models, at least `r ba_agreement`\% of the data points lie within the limits of agreement.
 For the pGRF equations, a tendency to higher dispersion can be observed for higher magnitudes, which correspond to the running trials, and especially in values derived from ankle-worn accelerometers (Supplemental Figure S3, Panels A and D).
@@
 Nevertheless, models from both studies present a caveat, that is the inclusion of a term in their equations that distinguishes whether the activity is walking or running, which is not easy to be identify from free-living accelerometer data, hindering the real-world application of these previous equations.
-Also, one of these studies included in their sample only children and adolescents [@Neugebauer2012].
+Also, one of these studies included in their sample only children and adolescents [@Neugebauer2012], that present a higher variability in their gait pattern compared to the adults [@Gouelle2016], making this equation unsuitable to use in an adult population.
 It is also important to mention that the prediction error found in our pGRF models is lower than the error typically found in models developed to estimate energy expenditure [@Diniz-Sousa2020], which is one of the main uses of accelerometer data [@Migueles2017].
@@
 Nevertheless, future studies should test the out-of-sample validity of these models.
-Lastly, the protocol was performed on an instrumented treadmill.
+Also, the protocol was performed on an instrumented treadmill.
 The use of a treadmill, although enhancing the study internal validity by the continuously measuring the GRF [@Ngoh2018] also reduces the external validity due to possible biomechanical changes in the walking and running patterns induced by the treadmill [@Kluitenberg2012].
+Lastly, bone adaptation to mechanical stimulation depends not only on external loads, such as GRF, but also on internal loads generated by muscle forces [@Robling2009].
+As accelerometers cannot reliably assess muscle force generation, they may have a limited capacity in estimating the bone response to loading.
 
@@
 The prediction models developed in this study have proved to be an efficient method to assess pGRF out of laboratory conditions, with potential applications for monitoring the mechanical loading in several contexts and establishing a relationship between mechanican loading and either beneficial or pathological osteoarticular adaptations.
-This is also the first work, to the best of our knowledge, to explore the prediction of pLR based on accelerometry data during both walking and running, although the accuracy of these models were unsatisfactory.
+This is also the first work, to the best of our knowledge, to explore the prediction of pLR based on accelerometry data during both walking and running, although the accuracy of these models was unsatisfactory.
 

### 2021-06-09 manuscript/manuscript.Rmd
Add more HF suggetions

@@
 
+```{r accuracy_no_cv}
+accuracy_no_cv <- map_dfr(res_GRF_models, accuracy) %>%
+  select(MAPE) %>%
+  mutate(MAPE = as.numeric(MAPE))
+accuracy_no_cv <- myround(max(accuracy_no_cv) * 100, 1)
+```
+
 <center>
@@
 
-Data from FP and accelerometers were collected at 1000 Hz and 100 Hz sampling frequency, respectively, and operated through the manufacturer supplied software (for the FP: Netforce, Version 3.5.1; AMTI Corporation, Watertown, USA; for the accelerometers: ActiLife version 6.13.3; Actigraph, Pensacola, USA).
+Data from FP were collected at 1000 Hz and from accelerometers at 100 Hz, which is the GT9X Link maximum sampling frequency, and operated through the manufacturer supplied software (for the FP: Netforce, Version 3.5.1; AMTI Corporation, Watertown, USA; for the accelerometers: ActiLife version 6.13.3; Actigraph, Pensacola, USA).
 Raw GRF (expressed in N) and acceleration (ACC; expressed in gravitational ACC units; 1*g* = 9.807 m·s^-2^) from the x, y and z vectors were exported to plain text files.
@@
 Peak GRF (pGRF) was determined as the highest value between 0.4 seconds before and after each pACC.
-The rates of change were computed through a centered derivative from the beginning of the foot contact to the pGRF.
-The peak loading rate (pLR) and peak acceleration rate (pAR) were defined as the maximum value in the array.
+The rates of change were computed through a centered derivative from the beginning of the foot contact to the peak (Equation 1).
+The peak loading rate (pLR) and peak acceleration rate (pAR) were defined as the maximum value in the array (Equation 2).
 
@@
 To assess the prediction accuracy, MAE, MAPE and RMSE were calculated using pGRF values predicted from both equations.
-No pLR prediction equation suited to be compared with ours was found.
+No pLR prediction equation developed for both walking and running was found to be compared with ours.
 
@@
 
-Furthermore, we compared the performance of our equations with a reference equation previously published.
+Furthermore, we compared the performance of our equations with a reference equation previously published [@Neugebauer2014].
 As the only comparable equation found was the one to predict pGRF vertical vector using hip-worn accelerometer data developed by Neugebauer et al [@Neugebauer2014], we applied it to our data and computed the MAE, MAPE and RMSE.
@@
 Comparing our results with previous studies that used a similar approach to predict pGRF, it can be seen that prediction errors from our models are comparable with some studies whose best results presented a MAPE of approximately 6.0\% but which included only walking in their protocol [@Fortune2014; @Neugebauer2018; @Veras2020].
-Additionally, previous studies that developed pGRF prediction models during both walking and running showed a MAPE of 8.3\% [@Neugebauer2014], which is higher than our results, and 5.2\% [@Neugebauer2012] with no cross-validation, which is similar to our results when no cross-validation method is applied (data not shown).
-Nevertheless, models from both studies present a caveat, that is the inclusion of a term in their equations that distinguishes whether the activity is walking or running, which is not easy to be identify from free-living accelerometer data, hindering the real-world application of these previous equations.
+Additionally, previous studies that developed pGRF prediction models during both walking and running showed a MAPE of 8.3\% [@Neugebauer2014], which is higher than our results, and 5.2\% [@Neugebauer2012] with no model cross-validation, which is similar to our results when no cross-validation method is applied (highest value of `r accuracy_no_cv`\% in our models).
+Nevertheless, models from both studies present a caveat, that is the inclusion of a term in their equations that distinguishes whether the activity is walking or running, which is not easy to be identified from free-living accelerometer data, hindering the real-world application of these previous equations.
 Also, one of these studies included in their sample only children and adolescents [@Neugebauer2012], that present a higher variability in their gait pattern compared to the adults [@Gouelle2016], making this equation unsuitable to use in an adult population.
+Furthermore, it was observed that the reference equation [@Neugebauer2014] used to compare with ours has a much lower accuracy during walking compared with running (16.6\% vs.
+9.7\%).
+As this reference equation was not validated in a sample including people with obesitu, it performs poorly when applied to data from people with high body mass, as is the case in our sample that performed the walking trials, compared with running trials mostly performed by non-obese people.
 It is also important to mention that the prediction error found in our pGRF models is lower than the error typically found in models developed to estimate energy expenditure [@Diniz-Sousa2020], which is one of the main uses of accelerometer data [@Migueles2017].
@@
 Although pLR is an important variable, known for a long time to determine to a great extent the bone formation response to mechanical loading [@Bassey1998; @Turner1995], based on our findings and in previous studies [@Fortune2014; @Veras2020], its prediction through accelerometer data does not seem to yield trustworthy results.
-Nevertheless, past research has suggested that the bone response to pGRF and pLR is somewhat similar during walking, running and jumping activities due to the high correlation between these variables [@Heikkinen2007].
-Therefore, the pGRF prediction by accelerometry data, which is shown to be accurate, already incorporates some of the LR effect on the bone response to loading.
-
 \medskip
 
-Results from our study offer some advantages compared to previous investigations: i) it includes walking and running, which are among the most usual types of physical activity performed daily; ii) all models developed, especially for pGRF prediction, presented a similar or better accuracy compared to the results from previous publications; iii) the accuracy of our models is constant throughout walking and running, notably for the pGRF prediction models, and therefore they can be used adequately for both activities; iv) in order to apply our models there is no need to manually distinguish the accelerometer data between walking and running, as in previously developed equations [@Neugebauer2014], which facilitates its use with daily living accelerometer data; and v) along with the creation of these models, a software was developed in the form of an R package to perform all the computations needed to get from the raw accelerometer data to the mechanical loading prediction.
+Results from our study offer some advantages compared to previous investigations: i) it includes walking and running, which are among the most usual types of physical activity performed daily; ii) all models developed, especially for pGRF prediction, presented a similar or better accuracy compared to the results from previous publications; iii) the accuracy of our models is fairly constant throughout walking and running, notably for the pGRF prediction models, with a MAPE increase from 6\% to 8\%, showing, therefore, that they can be used adequately for both activities; iv) in order to apply our models there is no need to manually distinguish the accelerometer data between walking and running, as in previously developed equations [@Neugebauer2014], which facilitates its use with daily living accelerometer data; and v) along with the creation of these models, a software was developed in the form of an R package to perform all the computations needed to get from the raw accelerometer data to the mechanical loading prediction.
 
@@
 Lastly, bone adaptation to mechanical stimulation depends not only on external loads, such as GRF, but also on internal loads generated by muscle forces [@Robling2009].
-As accelerometers cannot reliably assess muscle force generation, they may have a limited capacity in estimating the bone response to loading.
+As accelerometers cannot reliably assess muscle force generation, they may have a limited capacity in estimating the bone response to loading in situations where there are no major gravitational loadings involved.
 

### 2021-06-21 manuscript/manuscript.Rmd
Add JO and GB suggestions

@@
 ---
-title: "Mechanical loading prediction though accelerometry data during walking and running"
+title: "Mechanical loading prediction through accelerometry data during walking and running"
 output:
@@
 The pLR prediction equations yielded an unsatisfactory accuracy.
-**Key Words:** ACTIVITY MONITOR, FORCE PLATES, LOCOMOTION, RAW ACCELERATION, MECHANICAL LOADING.
+**Key Words:** activity monitor; force plates; locomotion; raw acceleration; mechanical loading.
 
@@
 However, not all forms of mechanical stimulation are equally osteogenic, and the manipulation of variables such as loading magnitude, rate, frequency and volume can produce different bone responses [@Turner1998; @Turner2003].
-Therefore, the correct assessment of mechanical loading is of paramount importance to better understand the bone response to physical stimulus.
+Therefore, the accurate assessment of mechanical loading is of paramount importance to better understand the bone response to physical stimulus.
 Nonetheless, a limitation in this context is that mechanical loading evaluation is usually performed only by instruments restrained to laboratory settings.
@@
 Accelerometers, due to their portability, low-cost and reliability, are widely used to objectively measure physical activity related to cardiometabolic outcomes [@Troiano2014; @Migueles2017], and, more recently, have been proposed as adequate devices to predict gravitational loading variables [@Neugebauer2012; @Fortune2014; @Neugebauer2014; @Neugebauer2018; @Veras2020].
-Given the need to assess mechanical loading in out-of-laboratory environments in order to control and monitor exercise prescription for bone health, accurate mechanical loading prediction models need to be developed and validated for the most common activities prescribed to induce bone adaptations, such as walking, running, and jumping.
-These models, apart from being accurate, need to be transparent in their data processing methods, allowing replicability of the results, and ideally offer a tool to facilitate data processing and analysis for those who are not familiar with the employed techniques, such as what is already offered for analysing cardiometabolic variables [@Migueles2019], and thus spread their applicability.
+Given the need to assess mechanical loading in out-of-laboratory environments to control and monitor exercise prescription for bone health, accurate mechanical loading prediction models need to be developed and validated for the most common physical activities prescribed to induce bone adaptations, such as walking, running, and jumping.
+These models, apart from being accurate, need to be transparent in their data processing methods, allowing reproducibility of the results, and ideally offer a tool to facilitate data processing and analysis for those who are not familiar with the employed techniques, such as what is already offered for analysing cardiometabolic variables [@Migueles2019], and therefore increase their applicability.
 
@@
 The ground reaction force (GRF) is arguably the most studied mechanical loading variable, and its relationship with bone adaptations to exercise is well established [@Turner2003; @Rantalainen2008; @Rantalainen2009].
-Due to its importance, some studies have already attempted to develop peak GRF (pGRF) prediction equations based on accelerometry data in several populations, such as children [@Neugebauer2012], adults [@Fortune2014; @Neugebauer2014], patients with obesity [@Veras2020] and military personal [@Neugebauer2018], but with the majority of these studies including only activities eliciting a low loading magnitude.
-Besides loading magnitude, loading rate (LR) is another variable related with bone health [@Rowlands2014] that could potentially be predicted by accelerometer data, but with apparently lower prediction accuracy compared with pGRF [@Fortune2014; @Veras2020].
-Also, to the best of our knowledge, peak LR (pLR) prediction through accelerometer data has only been tested during walking.
+Due to its importance, some studies have already attempted to develop peak GRF (pGRF) prediction equations based on accelerometry data in different populations, such as children [@Neugebauer2012], adults [@Fortune2014; @Neugebauer2014], patients with obesity [@Veras2020] and military personal [@Neugebauer2018], but with the majority of these studies including only activities eliciting a low loading magnitude.
+Besides loading magnitude, loading rate (LR) is another variable related to bone health [@Rowlands2014] that could potentially be predicted by accelerometer data, but with apparently lower prediction accuracy compared with pGRF [@Fortune2014; @Veras2020].
+Furthermore, to the best of our knowledge, peak LR (pLR) prediction through accelerometer data has only been tested during walking.
 
@@
 
-Therefore, given the necessity of an accurate and reliable method to assess mechanical loading in free-living conditions in order to better monitor and prescribe exercise to obtain positive bone health outcomes, the aim of this study was to develop pGRF and pLR prediction models based on accelerometer data during walking and running.
+Therefore, given the necessity of an accurate and reliable method to assess mechanical loading in free-living conditions to better monitor and prescribe exercise to achieve positive bone health outcomes, this study aimed to develop pGRF and pLR prediction models based on accelerometer data during walking and running.
 
@@
 Before starting the data collection, participants were given time to familiarize with the treadmill.
-The protocol consisted of walking and running in 13 incremental speeds, with 0º inclination, from 2 to 14 km·h^-1^ with 1 km·h^-1^ increment.
+The protocol consisted of walking and running, from 2 to 14 km·h^-1^ with 1 km·h^-1^ increment and 0º slope.
 Walking and running speeds ranged from 2 to 6 km·h^-1^ and from 7 to 14 km·h^-1^, respectively.
 Each walking speed lasted 1 minute and each running speed 30 seconds.
-To ensure single foot contact on each FP during the trial, feedback regarding location on the treadmill was provided by a researcher.
+To ensure single foot contact on each FP during the trial, feedback regarding the location on the treadmill was provided by a researcher.
 Participants wore their own sports shoes throughout the trials.
@@
 Activity monitors at waist level were secured on the same elastic belt with a clip and the activity monitor placed on the ankle was secured by an elastic belt fixed with adhesive tape.
-In all placements the accelerometer vertical axis was aligned with the standing body longitudinal axis.
+In all placements, the accelerometer vertical axis was aligned with the standing body longitudinal axis.
 
@@
 
-Activity monitors used were the GT9X Link model (±16*g* range; ActiGraph, Pensacola, USA) that incorporates primary and secondary triaxial accelerometers.
+Motion sensors used were the GT9X Link model (±16*g* range; ActiGraph, Pensacola, USA) that incorporates primary and secondary triaxial accelerometers.
 As a manufacturer proprietary filter is applied on the primary, but not on the secondary accelerometer raw data, only data from the latter was used in this study, reducing filtering bias and enhancing the data processing replicability.
@@
 
-Regions of interest for data analysis were manually selected where locomotion patterns were constant, which corresponded of 20-45 seconds periods, on average.
+Regions of interest for data analysis were manually selected where locomotion patterns were constant, representing 20-45 seconds periods, on average.
 Then, pGRF, pACC, pLR and pAR means of the resultant and its vertical component for each participant at each speed and for each accelerometer placement were extracted and used in all remaining analyses.
-In walking trials, a median of `r n_peaks_desc[which(n_peaks_desc$locomotion_type == "walking"), "n_peaks_median"]` peaks (interquartile range `r n_peaks_desc[which(n_peaks_desc$locomotion_type == "walking"), "n_peaks_median"] - n_peaks_desc[which(n_peaks_desc$locomotion_type == "walking"), "n_peaks_iqr"]`\textendash`r n_peaks_desc[which(n_peaks_desc$locomotion_type == "walking"), "n_peaks_median"] + n_peaks_desc[which(n_peaks_desc$locomotion_type == "walking"), "n_peaks_iqr"]`) were used to calculate these means, while in running trials it was a median of `r n_peaks_desc[which(n_peaks_desc$locomotion_type == "running"), "n_peaks_median"]` peaks (interquartile range `r n_peaks_desc[which(n_peaks_desc$locomotion_type == "running"), "n_peaks_median"] - n_peaks_desc[which(n_peaks_desc$locomotion_type == "running"), "n_peaks_iqr"]`\textendash`r n_peaks_desc[which(n_peaks_desc$locomotion_type == "running"), "n_peaks_median"] + n_peaks_desc[which(n_peaks_desc$locomotion_type == "running"), "n_peaks_iqr"]`).
+In walking and running trials, `r round(n_peaks_desc[which(n_peaks_desc$locomotion_type == "walking"), "n_peaks_mean"], 0)`±`r round(n_peaks_desc[which(n_peaks_desc$locomotion_type == "walking"), "n_peaks_sd"], 0)` and `r round(n_peaks_desc[which(n_peaks_desc$locomotion_type == "running"), "n_peaks_mean"], 0)`±`r round(n_peaks_desc[which(n_peaks_desc$locomotion_type == "running"), "n_peaks_sd"], 0)` peaks were used to calculate these means, respectively.
 
@@
 Final models were chosen according to -2 log-likelihood statistics [@Field2012].
-Traditional coefficient of determination (R^2^) was represented by conditional R^2^, that estimates the variance explained by the whole model [@Nakagawa2013].
+Traditional coefficient of determination (R^2^) was represented by conditional R^2^, which estimates the variance explained by the whole model [@Nakagawa2013].
 
@@
 Bias was computed as the mean of these differences and the limits of agreement were obtained using ±1.96 standard deviation of the bias [@Bland1986].
-In order to test whether there was a consistent over or underestimation of the actual values, one-sample *t* tests were performed to assess if the difference between actual and predicted values were different than zero.
+To test whether there was a consistent over or underestimation of the actual values, one-sample *t*-tests were performed to assess if the difference between actual and predicted values were different than zero.
 Linear regressions were applied to identify if the magnitude of the mean between actual and predicted values influenced the magnitude of their difference.
@@
 Finally, our pGRF prediction equation was compared with a previously published reference equation [@Neugebauer2014], that was developed using similar methods.
-This comparison was performed in three ways using: i) the whole sample; ii) a subsample of walking trials; iii) a subsample of running trials.
+This comparison was performed in three ways, using: i) the whole sample; ii) a subsample of walking trials; iii) a subsample of running trials.
 To assess the prediction accuracy, MAE, MAPE and RMSE were calculated using pGRF values predicted from both equations.
@@
 
-Our recruited sample presented a broad range of body masses (`r myround(min(mechanical_load_data$body_mass))`-`r myround(max(mechanical_load_data$body_mass))`kg; `r myround(min(mechanical_load_data$BMI))`-`r myround(max(mechanical_load_data$BMI))`kg·m^-2^), from normal weight to severe obesity.
-It was expected that participants with higher body mass index (BMI) values, due to their usually lower motor ability and physical capacity, would be able to perform the trials only at the lowest speeds.
-This was precisely what happened and, as can be observed in Figure 1, the number of data points corresponding to subjects with BMI > 30 kg·m^-2^ greatly decreases in the speeds above 6 km·h^-1^.
+The sample presented a broad range of body masses (`r myround(min(mechanical_load_data$body_mass))`-`r myround(max(mechanical_load_data$body_mass))`kg; `r myround(min(mechanical_load_data$BMI))`-`r myround(max(mechanical_load_data$BMI))`kg·m^-2^), from normal weight to severe obesity.
+It was expected that participants with higher body mass index (BMI) , due to their probable lower motor ability and physical capacity, would be able to perform the trials only at the lowest speeds.
+This was precisely what happened and, as can be observed in Figure 1, the number of data points corresponding to subjects with BMI > 30 kg·m^-2^ greatly decreases in speeds above 6 km·h^-1^.
 
@@
 Regression equations were developed to predict the mechanical loading variables based on accelerometry data.
-In the pGRF prediction equations, body mass and pACC were significant predictors and  therefore were included in the models.
+In the pGRF prediction equations, body mass and pACC were significant predictors and, therefore, included in the models.
 Table 1 shows the regression coefficients, R^2^ and accuracy indices from models derived from accelerometers placed at ankle, lower back and hip, and from the resultant vector and its vertical component.
@@
 Also, MAE ranged between `r myround(min(GRF_accuracy$MAE), 1)` N and `r myround(max(GRF_accuracy$MAE), 1)` N, and the MAPE from `r myround(min(GRF_accuracy$MAPE) * 100, 1)`\% to `r myround(max(GRF_accuracy$MAPE) * 100, 1)`\%, which is a smaller error than in previously published equations [@Neugebauer2014; @Fortune2014], and similar to the error from equations previously published by our group for walking only [@Veras2020].
-RMSE were also computed and shown to range between `r myround(min(GRF_accuracy$RMSE), 1)` N and `r myround(max(GRF_accuracy$RMSE), 1)` N.
+RMSE was also computed and shown to range between `r myround(min(GRF_accuracy$RMSE), 1)` N and `r myround(max(GRF_accuracy$RMSE), 1)` N.
 Although all pGRF prediction equations showed very similar results, the one based on hip-worn accelerometers and the resultant vector has shown the best performance, with an R^2^ of `r hip_GRF_accuracy$R2_cond`, the lowest MAE (`r myround(hip_GRF_accuracy$MAE, 1)` N), MAPE (`r myround(hip_GRF_accuracy$MAPE * 100, 1)`\%) and RMSE (`r myround(hip_GRF_accuracy$RMSE, 1)` N).
@@
 Apart from the overall equations performance, a sub-analysis was done separating walking and running trials.
-The results of these comparisons can be observed in the Table 2.
+The results of these comparisons can be observed in Table 2.
 Accuracy indices of our equations were all better compared to the reference equation.
@@
 
-Also, to simplify the use of these equations and enhance their applicability, a software package was developed for the use in the R statistical software (REF).
+Also, to simplify the use of these equations and enhance their applicability, a software package was developed for use in the R statistical software (REF).
 Included in this package are functionalities to read raw ACC data, process it and apply the equations to predict both pGRF and pLR.
-In the package website (ADDRESS) there are also instructions on how to use it to obtain the predictions, namely in the 'Home' and 'Get started' sections.
+In the package website (LINK) there are also instructions on how to use it to obtain the predictions, namely in the 'Home' and 'Get started' sections.
 
@@
 
-Comparing our results with previous studies that used a similar approach to predict pGRF, it can be seen that prediction errors from our models are comparable with some studies whose best results presented a MAPE of approximately 6.0\% but which included only walking in their protocol [@Fortune2014; @Neugebauer2018; @Veras2020].
+Comparing our results with previous studies that used a similar approach to predict pGRF, it can be seen that prediction errors from our models are comparable with some studies whose best results presented a MAPE of approximately 6.0\%, but which included only walking in their protocol [@Fortune2014; @Neugebauer2018; @Veras2020].
 Additionally, previous studies that developed pGRF prediction models during both walking and running showed a MAPE of 8.3\% [@Neugebauer2014], which is higher than our results, and 5.2\% [@Neugebauer2012] with no model cross-validation, which is similar to our results when no cross-validation method is applied (highest value of `r accuracy_no_cv`\% in our models).
 Nevertheless, models from both studies present a caveat, that is the inclusion of a term in their equations that distinguishes whether the activity is walking or running, which is not easy to be identified from free-living accelerometer data, hindering the real-world application of these previous equations.
-Also, one of these studies included in their sample only children and adolescents [@Neugebauer2012], that present a higher variability in their gait pattern compared to the adults [@Gouelle2016], making this equation unsuitable to use in an adult population.
-Furthermore, it was observed that the reference equation [@Neugebauer2014] used to compare with ours has a much lower accuracy during walking compared with running (16.6\% vs.
+Also, one of these studies included in their sample only children and adolescents [@Neugebauer2012], which present a higher variability in their gait pattern compared to the adults [@Gouelle2016], making this equation unsuitable to use in an adult population.
+Furthermore, it was observed that the reference equation [@Neugebauer2014] used as reference has a much lower accuracy during walking compared with running (16.6\% vs.
 9.7\%).
-As this reference equation was not validated in a sample including people with obesitu, it performs poorly when applied to data from people with high body mass, as is the case in our sample that performed the walking trials, compared with running trials mostly performed by non-obese people.
+As this reference equation was not validated in a sample including people with obesity, it performs poorly when applied to data from people with high body mass, as is the case in our sample that performed the walking trials, compared with running trials mostly performed by non-obese people.
 It is also important to mention that the prediction error found in our pGRF models is lower than the error typically found in models developed to estimate energy expenditure [@Diniz-Sousa2020], which is one of the main uses of accelerometer data [@Migueles2017].
@@
 Despite the validity of the prediction models developed here, some limitations should be mentioned.
-First, although the activities included in our protocol (walking and running) are among the most frequently performed motor activities during daily living, our prediction models are not applicable to other activities known to influence bone adaptation to mechanical stimuli, such as jumping and direction changing.
+First, although the activities included in our protocol (walking and running) are among the most frequently performed motor activities during daily living, our prediction models do not apply to other activities known to influence bone adaptation to mechanical stimuli, such as jumping and direction changing.
 Second, in the absence of another sample for model validation, the leave-one-out cross-validation method was used, which is the method of choice in these situations [@Staudenmayer2012].
 Nevertheless, future studies should test the out-of-sample validity of these models.
-Also, the protocol was performed on an instrumented treadmill.
+Moreover, the protocol was performed on an instrumented treadmill.
 The use of a treadmill, although enhancing the study internal validity by the continuously measuring the GRF [@Ngoh2018] also reduces the external validity due to possible biomechanical changes in the walking and running patterns induced by the treadmill [@Kluitenberg2012].

### 2021-06-21 code/02_explore.R
Change median and iqr for mean and sd to describe the number of peaks

@@ -43,7 +43,7 @@ n_peaks_desc <- mechanical_load_data %>%
   ) %>%
   group_by(locomotion_type) %>%
   summarise(
-    n_peaks_median = median(n_peaks), n_peaks_iqr = IQR(n_peaks)
+    n_peaks_mean = mean(n_peaks), n_peaks_sd = sd(n_peaks)
   )
 
 # GRF and ACC magnitude and rate per running speeed -----------------------

### 2021-06-22 manuscript/manuscript.Rmd
Update authors and affiliations

@@
 <center>
-Lucas Veras$^1$, Florêncio Diniz-Sousa$^1$, Giorjines Boppre$^1$, Vítor Devezas$^2$, Hugo Santos-Sousa$^2$, John Preto$^2$, João Paulo Vilas-Boas$^{3,4}$, Leandro Machado$^{3,4}$, José Oliveira$^1$, Hélder Fonseca$^1$
+Lucas Veras$^{1,2}$, Florêncio Diniz-Sousa$^{1,2}$, Giorjines Boppre$^{1,2}$, Edgar Moutinho-Ribeiro$^{1,2}$, Ana Resende-Coelho$^{1,2}$, Vítor Devezas$^3$, Hugo Santos-Sousa$^3$, John Preto$^3$, João Paulo Vilas-Boas$^{4,5}$, Leandro Machado$^{4,5}$, José Oliveira$^{1,2}$, Hélder Fonseca$^{1,2}$
 </center>
@@
 
-2 - General Surgery Department, São João Medical Center, Porto, Portugal
+2 - Laboratory for Integrative and Translational Research in Population Health (ITR), University of Porto, Porto, Portugal
 
-3 - Center of Research, Education, Innovation and Intervention in Sport (CIFI2D), Faculty of Sport, University of Porto, Porto, Portugal
+3 - General Surgery Department, São João Medical Center, Porto, Portugal
 
-4 - Biomechanics Laboratory (LABIOMEP-UP), University of Porto, Porto, Portugal
+4 - Center of Research, Education, Innovation and Intervention in Sport (CIFI2D), Faculty of Sport, University of Porto, Porto, Portugal
+
+5 - Biomechanics Laboratory (LABIOMEP-UP), University of Porto, Porto, Portugal
 
@@
 This study was funded by the Fundação para a Ciência e a Tecnologia (FCT) (grant PTDC/DTP-DES/0968/2014) and by the European Regional Development Fund (ERDF) through the Operational Competitiveness Programme (COMPETE) (grant POCI-01-0145-FEDER-016707).
-The study was developed in the Research Centre in Physical Activity, Health and Leisure (CIAFEL) funded by ERDF through the COMPETE and by the FCT (grant UIDB/00617/2020).
-Lucas Veras, Florêncio Diniz-Sousa and Giorjines Bopper were supported by the FCT grants UI/BD/150673/2020, SFRH/BD/117622/2016, SFRH/BD/146976/2019, respectively.
+The study was developed in the Research Centre in Physical Activity, Health and Leisure (CIAFEL) funded by ERDF through the COMPETE and by the FCT (grant UIDB/00617/2020), and Laboratory for Integrative and Translational Research in Population Health (ITR).
+Lucas Veras, Florêncio Diniz-Sousa and Giorjines Boppre were supported by the FCT grants UI/BD/150673/2020, SFRH/BD/117622/2016, SFRH/BD/146976/2019, respectively.
 The authors thank the participants who took part in this research and all who have collaborated in the project.

### 2021-06-28 manuscript/manuscript.Rmd
Add reference

@@
 Statistical analyses were conducted using R statistical software (version `r paste(R.Version()$major, R.Version()$minor, sep = ".")`, R Foundation for Statistical Computing, Vienna, Austria).
-All statistical analyses were registered in an open platform, where R code utilized and detailed information can be assessed (LINK).
+All statistical analyses were registered in an open platform, where R code utilized can be assessed (https://bit.ly/3A6P3Tx).
 The statistically significant value was set as $\alpha$ = 0.05.
@@
 
-Also, to simplify the use of these equations and enhance their applicability, a software package was developed for use in the R statistical software (REF).
+Also, to simplify the use of these equations and enhance their applicability, a software package was developed for use in the R statistical software [@Veras2021a].
 Included in this package are functionalities to read raw ACC data, process it and apply the equations to predict both pGRF and pLR.
-In the package website (LINK) there are also instructions on how to use it to obtain the predictions, namely in the 'Home' and 'Get started' sections.
+In the package website (https://verasls.github.io/impactr) there are also instructions on how to use it to obtain the predictions, namely in the 'Home' and 'Get started' sections.
 

### 2021-06-28 figures/fig3.R
Remove legend for alpha

@@ -58,7 +58,8 @@ BA_GRF_res_hip <- cv_res_GRF_models$hip %>%
   ) +
   guides(
     color = guide_legend(title = "Body mass index category:"),
-    shape = guide_legend(title = "Locomotion type:")
+    shape = guide_legend(title = "Locomotion type:"),
+    alpha = FALSE
   ) +
   labs(
     title = "Resultant vector - Hip placement",
@@ -94,7 +95,8 @@ BA_LR_res_hip <- cv_res_LR_models$hip %>%
   ) +
   guides(
     color = guide_legend(title = "Body mass index category:"),
-    shape = guide_legend(title = "Locomotion type:")
+    shape = guide_legend(title = "Locomotion type:"),
+    alpha = FALSE
   ) +
   labs(
     title = "Resultant vector - Hip placement",

### 2021-06-28 figures/figS3.R

@@ -59,7 +59,8 @@ BA_GRF_res_ankle <- cv_res_GRF_models$ankle %>%
   ) +
   guides(
     color = guide_legend(title = "Body mass index category:"),
-    shape = guide_legend(title = "Locomotion type:")
+    shape = guide_legend(title = "Locomotion type:"),
+    alpha = FALSE
   ) +
   labs(
     title = "Resultant vector - Ankle placement",
@@ -93,7 +94,8 @@ BA_GRF_res_back <- cv_res_GRF_models$lower_back %>%
   ) +
   guides(
     color = guide_legend(title = "Body mass index category:"),
-    shape = guide_legend(title = "Locomotion type:")
+    shape = guide_legend(title = "Locomotion type:"),
+    alpha = FALSE
   ) +
   labs(
     title = "Resultant vector - Lower back placement",
@@ -127,7 +129,8 @@ BA_GRF_res_hip <- cv_res_GRF_models$hip %>%
   ) +
   guides(
     color = guide_legend(title = "Body mass index category:"),
-    shape = guide_legend(title = "Locomotion type:")
+    shape = guide_legend(title = "Locomotion type:"),
+    alpha = FALSE
   ) +
   labs(
     title = "Resultant vector - Hip placement",
@@ -161,7 +164,8 @@ BA_GRF_ver_ankle <- cv_ver_GRF_models$ankle %>%
   ) +
   guides(
     color = guide_legend(title = "Body mass index category:"),
-    shape = guide_legend(title = "Locomotion type:")
+    shape = guide_legend(title = "Locomotion type:"),
+    alpha = FALSE
   ) +
   labs(
     title = "Vertical vector - Ankle placement",
@@ -195,7 +199,8 @@ BA_GRF_ver_back <- cv_ver_GRF_models$lower_back %>%
   ) +
   guides(
     color = guide_legend(title = "Body mass index category:"),
-    shape = guide_legend(title = "Locomotion type:")
+    shape = guide_legend(title = "Locomotion type:"),
+    alpha = FALSE
   ) +
   labs(
     title = "Vertical vector - Lower back placement",
@@ -229,7 +234,8 @@ BA_GRF_ver_hip <- cv_ver_GRF_models$hip %>%
   ) +
   guides(
     color = guide_legend(title = "Body mass index category:"),
-    shape = guide_legend(title = "Locomotion type:")
+    shape = guide_legend(title = "Locomotion type:"),
+    alpha = FALSE
   ) +
   labs(
     title = "Vertical vector - Hip placement",

### 2021-06-28 figures/figS4.R

@@ -60,7 +60,8 @@ BA_LR_res_ankle <- cv_res_LR_models$ankle %>%
   ) +
   guides(
     color = guide_legend(title = "Body mass index category:"),
-    shape = guide_legend(title = "Locomotion type:")
+    shape = guide_legend(title = "Locomotion type:"),
+    alpha = FALSE
   ) +
   labs(
     title = "Resultant vector - Ankle placement",
@@ -95,7 +96,8 @@ BA_LR_res_back <- cv_res_LR_models$lower_back %>%
   ) +
   guides(
     color = guide_legend(title = "Body mass index category:"),
-    shape = guide_legend(title = "Locomotion type:")
+    shape = guide_legend(title = "Locomotion type:"),
+    alpha = FALSE
   ) +
   labs(
     title = "Resultant vector - Lower back placement",
@@ -130,7 +132,8 @@ BA_LR_res_hip <- cv_res_LR_models$hip %>%
   ) +
   guides(
     color = guide_legend(title = "Body mass index category:"),
-    shape = guide_legend(title = "Locomotion type:")
+    shape = guide_legend(title = "Locomotion type:"),
+    alpha = FALSE
   ) +
   labs(
     title = "Resultant vector - Hip placement",
@@ -165,7 +168,8 @@ BA_LR_ver_ankle <- cv_ver_LR_models$ankle %>%
   ) +
   guides(
     color = guide_legend(title = "Body mass index category:"),
-    shape = guide_legend(title = "Locomotion type:")
+    shape = guide_legend(title = "Locomotion type:"),
+    alpha = FALSE
   ) +
   labs(
     title = "Vertical vector - Ankle placement",
@@ -200,7 +204,8 @@ BA_LR_ver_back <- cv_ver_LR_models$lower_back %>%
   ) +
   guides(
     color = guide_legend(title = "Body mass index category:"),
-    shape = guide_legend(title = "Locomotion type:")
+    shape = guide_legend(title = "Locomotion type:"),
+    alpha = FALSE
   ) +
   labs(
     title = "Vertical vector - Lower back placement",
@@ -235,7 +240,8 @@ BA_LR_ver_hip <- cv_ver_LR_models$hip %>%
   ) +
   guides(
     color = guide_legend(title = "Body mass index category:"),
-    shape = guide_legend(title = "Locomotion type:")
+    shape = guide_legend(title = "Locomotion type:"),
+    alpha = FALSE
   ) +
   labs(
     title = "Vertical vector - Hip placement",

### 2021-09-24 manuscript/manuscript.Rmd
Add new study from Florencio in the discussion

@@
 Predictions were tested with accelerometers placed at different body locations, using both the resultant vector and its vertical component in a set of incremental speeds, ranging from slow walking to running.
-Our findings demonstrated that accelerometer-based regression models were a valid and accurate method to estimate pGRF during human locomotion, as all of the pGRF prediction models developed presented a MAPE below the 8\% threshold.
+Our findings demonstrated that accelerometer-based regression models were a valid and accurate method to estimate pGRF during human locomotion, as all pGRF prediction models developed presented a MAPE below the 8\% threshold.
 However, accelerometry-based equations developed to predict pLR did not yield trustworthy results.
@@
 But instead of analysing accelerometer data averaged in epochs similarly to some previous publications [@Chastin2014; @Stiles2017], which can dilute the higher intensities [@Fabre2020], with our method raw ACC can be analysed at a higher resolution by identifying the pGRF associated with each step taken and establishing, for example, the number of peaks above a certain magnitude threshold needed to elicit a positive bone adaptation.
-Another potential application could be the computation of the cumulative mechanical loading by adding up the pGRF of all detected steps.
-This information could be used either for the determination of the relationship between the mechanical loading accumulated in a certain period (e.g., a day) and the bone response to this loading [@Ahola2010], or also in sports training, by monitoring the amount of loading beared by athletes and relate this information with osteoarticular safety of the training regimen [@Vanwanseele2020].
+Another potential application could be the computation of the cumulative mechanical loading by adding up the pGRF of all detected steps [@Diniz-Sousa2021].
+This information could be used either for the determination of the relationship between the mechanical loading accumulated in a certain period (e.g., a day) and the bone response to this loading [@Ahola2010], or to investigate the causes of bone loss in populations with bone health issues, such as bariatric patients [@Diniz-Sousa2021].
+This approach has already been applied with equations developed in a previous research from our group [@Veras2020], but now can be extended to include both types of locomotion (walking and running) contemplated in the equations from the current study.
+Another possible application is in sports training, by monitoring the amount of loading beared by athletes and relate this information with osteoarticular safety of the training regimen [@Vanwanseele2020].
 
@@
 These models are valid for the use of both the resultant vector and its vertical component, and using data from accelerometers placed at either the ankle, lower back or hip.
-The prediction models developed in this study have proved to be an efficient method to assess pGRF out of laboratory conditions, with potential applications for monitoring the mechanical loading in several contexts and establishing a relationship between mechanican loading and either beneficial or pathological osteoarticular adaptations.
+The prediction models developed in this study have proved to be an efficient method to assess pGRF out of laboratory conditions, with potential applications for monitoring the mechanical loading in several contexts and establishing a relationship between mechanical loading and either beneficial or pathological osteoarticular adaptations.
 This is also the first work, to the best of our knowledge, to explore the prediction of pLR based on accelerometry data during both walking and running, although the accuracy of these models was unsatisfactory.

### 2021-09-24 manuscript/manuscript.Rmd
Do some small corrections in the text

@@
 Given the need to assess mechanical loading in out-of-laboratory environments to control and monitor exercise prescription for bone health, accurate mechanical loading prediction models need to be developed and validated for the most common physical activities prescribed to induce bone adaptations, such as walking, running, and jumping.
-These models, apart from being accurate, need to be transparent in their data processing methods, allowing reproducibility of the results, and ideally offer a tool to facilitate data processing and analysis for those who are not familiar with the employed techniques, such as what is already offered for analysing cardiometabolic variables [@Migueles2019], and therefore increase their applicability.
+These models, apart from being accurate, need to be transparent in their data processing methods, allowing reproducibility of the results, and ideally offer a tool to facilitate data processing and analysis for those who are not familiar with the employed techniques, such as what is already offered for analyzing cardiometabolic variables [@Migueles2019], and therefore increase their applicability.
 
@@
 Table 1 shows the regression coefficients, R^2^ and accuracy indices from models derived from accelerometers placed at ankle, lower back and hip, and from the resultant vector and its vertical component.
-In the pGRF prediction models, the lowest R^2^ value found was `r min(GRF_accuracy$R2_cond)`, which shows that the predictor variables are able to explain the variability in the outcome variable to a great extent.
+In the pGRF prediction models, the lowest R^2^ value found was `r min(GRF_accuracy$R2_cond)`, which shows that the predictor variables can explain the variability in the outcome variable to a great extent.
 Also, MAE ranged between `r myround(min(GRF_accuracy$MAE), 1)` N and `r myround(max(GRF_accuracy$MAE), 1)` N, and the MAPE from `r myround(min(GRF_accuracy$MAPE) * 100, 1)`\% to `r myround(max(GRF_accuracy$MAPE) * 100, 1)`\%, which is a smaller error than in previously published equations [@Neugebauer2014; @Fortune2014], and similar to the error from equations previously published by our group for walking only [@Veras2020].
@@
 
-Comparing our results with previous studies that used a similar approach to predict pGRF, it can be seen that prediction errors from our models are comparable with some studies whose best results presented a MAPE of approximately 6.0\%, but which included only walking in their protocol [@Fortune2014; @Neugebauer2018; @Veras2020].
+Comparing our results with previous studies that used a similar approach to predict pGRF, prediction errors from our models are comparable with some studies whose best results presented a MAPE of approximately 6.0\%, but which included only walking in their protocol [@Fortune2014; @Neugebauer2018; @Veras2020].
 Additionally, previous studies that developed pGRF prediction models during both walking and running showed a MAPE of 8.3\% [@Neugebauer2014], which is higher than our results, and 5.2\% [@Neugebauer2012] with no model cross-validation, which is similar to our results when no cross-validation method is applied (highest value of `r accuracy_no_cv`\% in our models).
@@
 Our findings revealed that ankle, lower back or hip were all valid accelerometer placements to accurately predict pGRF.
-However, despite the similarities, hip was the accelerometer placement with the best prediction accuracy, which reinforces it as the preferable location to collect and analyse accelerometry-measured physical activity data.
+However, despite the similarities, hip was the accelerometer placement with the best prediction accuracy, which reinforces it as the preferable location to collect and analyze accelerometry-measured physical activity data.
 This placement is already well accepted for energy expenditure prediction and physical activity intensity classification [@Migueles2017] and has also shown a high mechanical loading prediction ability in previous studies [@Veras2020; @Neugebauer2018; @Neugebauer2012].
-Besides various placements, the current study also analysed predictions from both the resultant and vertical vectors.
+Besides various placements, the current study also analyzed predictions from both the resultant and vertical vectors.
 Although ACC and GRF from the vertical vector are the most studied and frequently associated with bone health [@Vainionpaa2006; @Heikkinen2007], the resultant vector of both ACC and GRF during locomotion is greatly determined by the vertical vector, and, as can be observed in Supplemental Figure S1, in this study values from both vectors were very similar.
@@
 Besides these advantages, our models present some important potential clinical applications.
-For instance, these models can be used to analyse data from large observational studies and make associations between mechanical loading and bone health.
-But instead of analysing accelerometer data averaged in epochs similarly to some previous publications [@Chastin2014; @Stiles2017], which can dilute the higher intensities [@Fabre2020], with our method raw ACC can be analysed at a higher resolution by identifying the pGRF associated with each step taken and establishing, for example, the number of peaks above a certain magnitude threshold needed to elicit a positive bone adaptation.
+For instance, these models can be used to analyze data from large observational studies and make associations between mechanical loading and bone health.
+But instead of analyzing accelerometer data averaged in epochs similarly to some previous publications [@Chastin2014; @Stiles2017], which can dilute the higher intensities [@Fabre2020], with our method raw ACC can be analyzed at a higher resolution by identifying the pGRF associated with each step taken and establishing, for example, the number of peaks above a certain magnitude threshold needed to elicit a positive bone adaptation.
 Another potential application could be the computation of the cumulative mechanical loading by adding up the pGRF of all detected steps [@Diniz-Sousa2021].

### 2021-10-01 manuscript/manuscript.Rmd
Add HF suggestions

@@
 
-3 - General Surgery Department, São João Medical Center, Porto, Portugal
+3 - Centro de Responsabilidade Integrado de Obesidade (CRIO), Centro Hospitalar Universitário de São João, Porto, Portugal
 
@@
 Additionally, previous studies that developed pGRF prediction models during both walking and running showed a MAPE of 8.3\% [@Neugebauer2014], which is higher than our results, and 5.2\% [@Neugebauer2012] with no model cross-validation, which is similar to our results when no cross-validation method is applied (highest value of `r accuracy_no_cv`\% in our models).
-Nevertheless, models from both studies present a caveat, that is the inclusion of a term in their equations that distinguishes whether the activity is walking or running, which is not easy to be identified from free-living accelerometer data, hindering the real-world application of these previous equations.
-Also, one of these studies included in their sample only children and adolescents [@Neugebauer2012], which present a higher variability in their gait pattern compared to the adults [@Gouelle2016], making this equation unsuitable to use in an adult population.
-Furthermore, it was observed that the reference equation [@Neugebauer2014] used as reference has a much lower accuracy during walking compared with running (16.6\% vs.
+Nevertheless, models from both studies present a caveat, that is the inclusion of a term in their equations that implies distinguishing whether the activity is walking or running.
+This is not easy to be identified from free-living accelerometer data, hindering real-world application of these previous equations.
+Also, one of these studies included only children and adolescents in their sample [@Neugebauer2012], which present a higher variability in terms of gait pattern compared to adults [@Gouelle2016], making, therefore, this equation unsuitable to use in an adult population.
+Furthermore, it was observed that the equation used as reference in this study [@Neugebauer2014] has a much lower accuracy during walking compared with running (16.6\% vs.
 9.7\%).
@@
 However, despite the similarities, hip was the accelerometer placement with the best prediction accuracy, which reinforces it as the preferable location to collect and analyze accelerometry-measured physical activity data.
-This placement is already well accepted for energy expenditure prediction and physical activity intensity classification [@Migueles2017] and has also shown a high mechanical loading prediction ability in previous studies [@Veras2020; @Neugebauer2018; @Neugebauer2012].
-Besides various placements, the current study also analyzed predictions from both the resultant and vertical vectors.
-Although ACC and GRF from the vertical vector are the most studied and frequently associated with bone health [@Vainionpaa2006; @Heikkinen2007], the resultant vector of both ACC and GRF during locomotion is greatly determined by the vertical vector, and, as can be observed in Supplemental Figure S1, in this study values from both vectors were very similar.
+This placement is already well accepted for energy expenditure prediction and physical activity intensity classification purposes [@Migueles2017] and has also shown a high mechanical loading prediction ability in previous studies [@Veras2020; @Neugebauer2018; @Neugebauer2012].
+In addition to testing various accelerometer placements, the current study also investigated the prediction ability of both the resultant and vertical vectors.
+Although ACC and GRF from the vertical vector are the most studied and frequently associated with bone health variables [@Vainionpaa2006; @Heikkinen2007], the resultant vector of both ACC and GRF during locomotion is greatly determined by the vertical vector.
+In fact, as can be observed in Supplemental Figure S1, in this study, values from both vectors were very similar.
 Also, mechanical loading prediction based on the resultant ACC offers the important advantage of not being dependent on the accelerometer positioning and correct orientation, which is hard to assure when measuring daily physical activity [@Welk2005; @Wundersitz2015].
@@
 To the best of our knowledge, only two studies have already attempted to develop pLR prediction models based on accelerometry data, and both studies only included walking activities.
-Accordingly to what we found, these studies also showed that the prediction error for these models is considerably larger than for the pGRF prediction.
-One of these researches investigated the use of four different accelerometer placements (ankle, tibia, thigh and waist) to predict pLR and achieved a MAPE ranging from 23.4\% to 46.1\% [@Fortune2014], while the other obtained values from 19.5\% to 22.4\% using accelerometers at the lower back and hip [@Veras2020].
-Results from the models developed in these previous investigations have both lower and higher prediction errors compared to our results, which were approximately of 35\%.
-Although pLR is an important variable, known for a long time to determine to a great extent the bone formation response to mechanical loading [@Bassey1998; @Turner1995], based on our findings and in previous studies [@Fortune2014; @Veras2020], its prediction through accelerometer data does not seem to yield trustworthy results.
+In accordance with our own findings, these studies also showed that the prediction error for these models is considerably larger than that for the pGRF prediction.
+One of these studies investigated the use of four different accelerometer placements (ankle, tibia, thigh and waist) to predict pLR and achieved a MAPE ranging from 23.4\% to 46.1\% [@Fortune2014], while the other obtained values from 19.5\% to 22.4\% using accelerometers at the lower back and hip [@Veras2020].
+Results from the models developed in these previous studies have both lower and higher prediction errors compared to our own results, which were of approximately 35\%.
+Although pLR is an important variable, known for a long time to determine to a great extent the bone formation response to mechanical loading [@Bassey1998; @Turner1995], based on our findings and in previous studies [@Fortune2014; @Veras2020], its prediction through accelerometer data, at the current state of technical accelerometer development, does not seem to yield trustworthy results that can be used in models to predict the bone anabolic response to mechanical loading.
+
 \medskip
@@
 
-Besides these advantages, our models present some important potential clinical applications.
-For instance, these models can be used to analyze data from large observational studies and make associations between mechanical loading and bone health.
-But instead of analyzing accelerometer data averaged in epochs similarly to some previous publications [@Chastin2014; @Stiles2017], which can dilute the higher intensities [@Fabre2020], with our method raw ACC can be analyzed at a higher resolution by identifying the pGRF associated with each step taken and establishing, for example, the number of peaks above a certain magnitude threshold needed to elicit a positive bone adaptation.
+In addition to these technical advantages, our models present some important potential clinical applications.
+For instance, these models can be used to analyze data from large observational studies and to investigate associations between objectively mesured mechanical loading and bone health.
+Nevertheless, instead of analyzing accelerometer data averaged in epochs as it has been performed previously [@Chastin2014; @Stiles2017], which can have a dilution effect and hinder the detection of higher intensities [@Fabre2020], with our method, raw ACC can be analyzed at a higher resolution by identifying the pGRF associated with each step taken and establishing, for instance, the number of peaks above a certain magnitude threshold needed to elicit a positive bone adaptation.
 Another potential application could be the computation of the cumulative mechanical loading by adding up the pGRF of all detected steps [@Diniz-Sousa2021].
-This information could be used either for the determination of the relationship between the mechanical loading accumulated in a certain period (e.g., a day) and the bone response to this loading [@Ahola2010], or to investigate the causes of bone loss in populations with bone health issues, such as bariatric patients [@Diniz-Sousa2021].
-This approach has already been applied with equations developed in a previous research from our group [@Veras2020], but now can be extended to include both types of locomotion (walking and running) contemplated in the equations from the current study.
-Another possible application is in sports training, by monitoring the amount of loading beared by athletes and relate this information with osteoarticular safety of the training regimen [@Vanwanseele2020].
+This information could be used either for the determination of the relationship between mechanical loading accumulated within a certain period (e.g., a day) and the bone response to this loading [@Ahola2010], or to investigate the causes of bone loss in populations with bone health issues, such as post-bariatric surgery patients [@Diniz-Sousa2021].
+This approach has already been followed by using equations previously developed by our group [@Veras2020].
+However, these new equations described in the current study have the advantage of being able to be applied to both types of locomotion (walking and running) and not just walking.
+Another possible application is in sports training, by monitoring the amount of mechanical loading sustained by athletes and using this information for training loading monitoring and for establishing adequate measures to prevent osteoarticular overload and reduce the risk of osteoarthritis [@Vanwanseele2020].
 
@@
 Lastly, bone adaptation to mechanical stimulation depends not only on external loads, such as GRF, but also on internal loads generated by muscle forces [@Robling2009].
-As accelerometers cannot reliably assess muscle force generation, they may have a limited capacity in estimating the bone response to loading in situations where there are no major gravitational loadings involved.
+As accelerometers cannot reliably assess muscle force generation, they may have a limited capacity in estimating the bone response to loading in situations where there are no major gravitational loadings involved, such as during resistance training.
 
@@
 In conclusion, pGRF could be accurately predicted by accelerometry-based regression models.
-These models are valid for the use of both the resultant vector and its vertical component, and using data from accelerometers placed at either the ankle, lower back or hip.
-The prediction models developed in this study have proved to be an efficient method to assess pGRF out of laboratory conditions, with potential applications for monitoring the mechanical loading in several contexts and establishing a relationship between mechanical loading and either beneficial or pathological osteoarticular adaptations.
-This is also the first work, to the best of our knowledge, to explore the prediction of pLR based on accelerometry data during both walking and running, although the accuracy of these models was unsatisfactory.
+These models are valid for the use of both the resultant vector and its vertical component, and for using data derived from accelerometers placed at either the ankle, lower back or hip.
+The prediction models developed in this study have proved to be an efficient method to assess pGRF out of laboratory conditions, with potential applications for monitoring mechanical loading in several contexts and for establishing a relationship between mechanical loading and either beneficial or pathological skeletal and osteoarticular adaptations.
+This is also the first work, to the best of our knowledge, to explore the prediction of pLR based on accelerometry data during both walking and running, although the accuracy of these models was found to be unsatisfactory.
 
@@
 The study was developed in the Research Centre in Physical Activity, Health and Leisure (CIAFEL) funded by ERDF through the COMPETE and by the FCT (grant UIDB/00617/2020), and Laboratory for Integrative and Translational Research in Population Health (ITR).
-Lucas Veras, Florêncio Diniz-Sousa and Giorjines Boppre were supported by the FCT grants UI/BD/150673/2020, SFRH/BD/117622/2016, SFRH/BD/146976/2019, respectively.
-The authors thank the participants who took part in this research and all who have collaborated in the project.
+Lucas Veras, Florêncio Diniz-Sousa and Giorjines Boppre are supported by the FCT grants UI/BD/150673/2020, SFRH/BD/117622/2016 and SFRH/BD/146976/2019, respectively.
+The authors thank the participants who took part in this research and all who have collaborated in the research project.
 

### 2021-10-01 manuscript/manuscript.Rmd
Format manuscript to submit to JSHS

@@
 bibliography: bibliography.bib
-csl: nlm.csl
+csl: ama.csl
 ---
@@
 
-# Introduction {.unnumbered}
+# Introduction
 
@@
 
-# Methods {.unnumbered}
+# Methods
 
-## Participants {.unnumbered}
+## Participants
 
@@
 
-## Data collection protocol {.unnumbered}
+## Data collection protocol
 
@@
 
-## Data processing {.unnumbered}
+## Data processing
 
@@
 
-## Statistical analyses {.unnumbered}
+## Statistical analyses
 
@@
 
-# Results {.unnumbered}
+# Results
 
@@
 
-# Discussion {.unnumbered}
+# Discussion
 
@@
 
-\medskip
+# Conclusion
 

### 2022-03-08 code/03_predict.R
Include speed as a random factor in the loading rate models

@@ -70,7 +70,7 @@ plot_ver_GRF_models <- map(
 # Build LR models ---------------------------------------------------------
 
 LR_formula <- as.formula(
-  "pLR_Ns ~ pAR_gs + body_mass + pAR_gs:body_mass + (1 | subj)"
+  "pLR_Ns ~ pAR_gs + body_mass + pAR_gs:body_mass + (1 | subj) + (1 | speed)"
 )
 
 # Resultant vector

### 2022-03-08 manuscript/manuscript.Rmd
Revise manuscript

@@
 **Purpose:** To develop accelerometry-based equations to predict peak ground reaction forces (pGRF) and peak loading rate (pLR) during walking and running.
-**Methods:** `r max(sample_size$pGRF_N)` subjects (`r unname(sample_size_sex["F"])` females; `r sample_descriptives$body_mass_mean`±`r sample_descriptives$body_mass_sd`kg) walked and ran at different speeds (2–14km·h^-1^) on a force plate–equipped treadmill while wearing accelerometers at their ankle, lower back and hip.
+**Methods:** `r max(sample_size$pGRF_N)` subjects (`r unname(sample_size_sex["F"])` females; `r sample_descriptives$body_mass_mean`±`r sample_descriptives$body_mass_sd`kg) walked and ran at different speeds (2–14km·h^-1^) on a force plate–instrumented treadmill while wearing accelerometers at their ankle, lower back and hip.
 Regression equations were developed to predict pGRF and pLR from accelerometry data.
@@
 Besides loading magnitude, loading rate (LR) is another variable related to bone health [@Rowlands2014] that could potentially be predicted by accelerometer data, but with apparently lower prediction accuracy compared with pGRF [@Fortune2014; @Veras2020].
-Furthermore, to the best of our knowledge, peak LR (pLR) prediction through accelerometer data has only been tested during walking.
+Furthermore, to the best of our knowledge, peak LR (pLR) prediction through accelerometer data has only been tested during walking [@Fortune2014; @Veras2020].
 
@@
 Regression equations to predict pGFR and pLR in each vector (resultant and vertical) for each accelerometer placement (ankle, lower back and hip) were developed by linear mixed models.
-Body mass (kg) was tested as a predictor in all models.
+Body mass (kg), body mass index (BMI; kg·m^2^), and BMI category were tested as a predictor in all models.
 Additionally, pACC and pAR were tested as predictors in the pGRF and pLR models, respectively.
-All predictors were tested as fixed effects and were shown to be significant.
+All predictors were tested as fixed effects, and body mass, pACC and pAR were shown to be significant predictors.
 Subject and speed were tested as random effects, and the inclusion of both has showed to improve the models.
@@
 The sample presented a broad range of body masses (`r myround(min(mechanical_load_data$body_mass))`-`r myround(max(mechanical_load_data$body_mass))`kg; `r myround(min(mechanical_load_data$BMI))`-`r myround(max(mechanical_load_data$BMI))`kg·m^-2^), from normal weight to severe obesity.
-It was expected that participants with higher body mass index (BMI) , due to their probable lower motor ability and physical capacity, would be able to perform the trials only at the lowest speeds.
+It was expected that participants with higher BMI, due to their probable lower motor ability and physical capacity, would be able to perform the trials only at the lowest speeds.
 This was precisely what happened and, as can be observed in Figure 1, the number of data points corresponding to subjects with BMI > 30 kg·m^-2^ greatly decreases in speeds above 6 km·h^-1^.
@@
 As this reference equation was not validated in a sample including people with obesity, it performs poorly when applied to data from people with high body mass, as is the case in our sample that performed the walking trials, compared with running trials mostly performed by non-obese people.
-It is also important to mention that the prediction error found in our pGRF models is lower than the error typically found in models developed to estimate energy expenditure [@Diniz-Sousa2020], which is one of the main uses of accelerometer data [@Migueles2017].
 

### 2022-03-09 figures/fig2.R
Update figure

@@ -2,6 +2,7 @@
 
 library(here)
 library(tidyverse)
+library(lvmisc)
 library(ggsci)
 library(patchwork)
 library(ragg)
@@ -16,6 +17,28 @@ mechanical_load_data <- mechanical_load_data %>%
       "Walking"
     )
   )
+
+load(here("output", "loocv_data.rda"))
+cv_res_GRF_models <- map(
+  cv_res_GRF_models,
+  ~ mutate(
+    .x,
+    activity = fct_relevel(
+      as.factor(ifelse(speed %in% 1:6, "Walking", "Running")),
+      "Walking"
+    )
+  )
+)
+cv_res_LR_models <- map(
+  cv_res_LR_models,
+  ~ mutate(
+    .x,
+    activity = fct_relevel(
+      as.factor(ifelse(speed %in% 1:6, "Walking", "Running")),
+      "Walking"
+    )
+  )
+)
 
 # Hip pGRF x pACC plot --------------------------------------------------
 
@@ -38,17 +61,21 @@ scatterplot_GRF_res_hip <- mechanical_load_data %>%
   scale_x_continuous(
     limits = c(1, 9),
     expand = c(0, 0),
-    breaks = seq(2, 8, 1)
+    breaks = seq(1, 9, 1)
   ) +
   theme_light() +
   theme(
-    plot.title = element_text(size = 15, hjust = 0.5),
-    legend.title = element_text(size = 13),
-    legend.text = element_text(size = 12),
-    axis.title.y = element_text(size = 13),
-    axis.title.x = element_text(size = 13),
-    axis.text.y = element_text(size = 13),
-    axis.text.x = element_text(size = 13)
+    legend.position = "none",
+    plot.title = element_text(size = 17, hjust = 0.5),
+    # legend.title = element_text(size = 15),
+    # legend.title = element_blank(),
+    # legend.text = element_text(size = 14),
+    # legend.text = element_blank(),
+    axis.title.y = element_text(size = 15),
+    axis.title.x = element_text(size = 15),
+    axis.text.y = element_text(size = 15),
+    axis.text.x = element_text(size = 15, vjust = 0.2),
+    plot.margin = margin(r = 1, unit = "cm")
   ) +
   guides(
     color = guide_legend(title = "Body mass index category:"),
@@ -74,6 +101,7 @@ scatterplot_LR_res_hip <- mechanical_load_data %>%
   ) +
   scale_color_nejm() +
   scale_y_continuous(
[…]

### 2022-03-09 figures/fig3.R
Delete file

@@ -1,138 +1,0 @@
-# Load packages -----------------------------------------------------------
-
-library(here)
-library(tidyverse)
-library(lvmisc)
-library(ggsci)
-library(patchwork)
-library(ragg)
-
-# Load data ---------------------------------------------------------------
-
-load(here("output", "loocv_data.rda"))
-cv_res_GRF_models <- map(
-  cv_res_GRF_models,
-  ~ mutate(
-    .x,
-    activity = fct_relevel(
-      as.factor(ifelse(speed %in% 1:6, "Walking", "Running")),
-      "Walking"
-    )
-  )
-)
-cv_res_LR_models <- map(
-  cv_res_LR_models,
-  ~ mutate(
-    .x,
-    activity = fct_relevel(
-      as.factor(ifelse(speed %in% 1:6, "Walking", "Running")),
-      "Walking"
-    )
-  )
-)
-
-# Hip pGRF Bland-Altman plot ---------------------------------------------
-
-BA_GRF_res_hip <- cv_res_GRF_models$hip %>%
-  plot_bland_altman(color = BMI_cat, shape = activity, alpha = 0.5) +
-  scale_color_nejm() +
-  scale_y_continuous(
-    limits = c(-600, 600),
-    expand = c(0, 0),
-    breaks = seq(-600, 600, 200)
-  ) +
-  scale_x_continuous(
-    limits = c(400, 3100),
-    expand = c(0, 0),
-    breaks = seq(500, 3000, 500)
-  ) +
-  theme_light() +
-  theme(
-    plot.title = element_text(size = 15, hjust = 0.5),
-    legend.title = element_text(size = 13),
-    legend.text = element_text(size = 12),
-    axis.title.y = element_text(size = 13),
-    axis.title.x = element_text(size = 13),
-    axis.text.y = element_text(size = 13),
-    axis.text.x = element_text(size = 13)
-  ) +
-  guides(
-    color = guide_legend(title = "Body mass index category:"),
-    shape = guide_legend(title = "Locomotion type:"),
-    alpha = FALSE
-  ) +
-  labs(
-    title = "Resultant vector - Hip placement",
-    x = "Mean of Actual and Predicted pGRF (N)",
-    y = "Actual - Predicted pGRF (N)"
-  )
-
-# Hip pLR Bland-Altman plot ----------------------------------------------
-
-BA_LR_res_hip <- cv_res_LR_models$hip %>%
[…]

### 2022-03-09 figures/fig2.R
Delete commented out code

@@ -67,10 +67,6 @@ scatterplot_GRF_res_hip <- mechanical_load_data %>%
   theme(
     legend.position = "none",
     plot.title = element_text(size = 17, hjust = 0.5),
-    # legend.title = element_text(size = 15),
-    # legend.title = element_blank(),
-    # legend.text = element_text(size = 14),
-    # legend.text = element_blank(),
     axis.title.y = element_text(size = 15),
     axis.title.x = element_text(size = 15),
     axis.text.y = element_text(size = 15),
@@ -115,10 +111,6 @@ scatterplot_LR_res_hip <- mechanical_load_data %>%
   theme(
     legend.position = "none",
     plot.title = element_text(size = 17, hjust = 0.5),
-    # legend.title = element_text(size = 15),
-    # legend.title = element_blank(),
-    # legend.text = element_text(size = 14),
-    # legend.text = element_blank(),
     axis.title.y = element_text(size = 15),
     axis.title.x = element_text(size = 15),
     axis.text.y = element_text(size = 15),
@@ -224,11 +216,7 @@ BA_plots <- BA_GRF_res_hip +
 
 fig2 <- scatterplots / BA_plots +
   plot_annotation(tag_levels = "A") &
-  # plot_layout(guides = "collect") &
-  theme(
-    # legend.position = "bottom",
-    plot.tag = element_text(size = 16)
-  )
+  theme(plot.tag = element_text(size = 16))
 
 agg_tiff(
   here("figures", "fig2.tiff"),

### 2022-03-09 figures/figS1.R
Increase size of axes and labels text

@@ -28,7 +28,7 @@ scatterplot_GRF_res_ankle <- mechanical_load_data %>%
   ) +
   geom_smooth(
     aes(x = pACC_g, y = pGRF_N, color = BMI_cat),
-    method = "lm", se = FALSE
+    method = "lm", se = FALSE, show.legend = FALSE
   ) +
   scale_color_nejm() +
   scale_y_continuous(
@@ -37,23 +37,25 @@ scatterplot_GRF_res_ankle <- mechanical_load_data %>%
     breaks = seq(0, 3500, 500)
   ) +
   scale_x_continuous(
-    limits = c(1, 14),
+    limits = c(0, 14),
     expand = c(0, 0),
-    breaks = seq(0, 12, 2)
+    breaks = seq(0, 14, 2)
   ) +
   theme_light() +
   theme(
-    plot.title = element_text(size = 15, hjust = 0.5),
-    legend.title = element_text(size = 13),
-    legend.text = element_text(size = 12),
-    axis.title.y = element_text(size = 13),
-    axis.title.x = element_text(size = 13),
-    axis.text.y = element_text(size = 13),
-    axis.text.x = element_text(size = 13)
+    plot.title = element_text(size = 17, hjust = 0.5),
+    legend.title = element_text(size = 15),
+    legend.text = element_text(size = 15),
+    axis.title.y = element_text(size = 15),
+    axis.title.x = element_text(size = 15),
+    axis.text.y = element_text(size = 15),
+    axis.text.x = element_text(size = 15, vjust = 0.2),
+    plot.margin = margin(r = 1, unit = "cm")
   ) +
   guides(
     color = guide_legend(title = "Body mass index category:"),
-    shape = guide_legend(title = "Locomotion type:")
+    shape = guide_legend(title = "Locomotion type:"),
+    alpha = FALSE
   ) +
   labs(
     title = "Resultant vector - Ankle placement",
@@ -70,7 +72,7 @@ scatterplot_GRF_res_back <- mechanical_load_data %>%
   ) +
   geom_smooth(
     aes(x = pACC_g, y = pGRF_N, color = BMI_cat),
-    method = "lm", se = FALSE
+    method = "lm", se = FALSE, show.legend = FALSE
   ) +
   scale_color_nejm() +
   scale_y_continuous(
@@ -81,21 +83,23 @@ scatterplot_GRF_res_back <- mechanical_load_data %>%
   scale_x_continuous(
[…]

### 2022-03-09 figures/figS2.R

@@ -28,10 +28,11 @@ scatterplot_LR_res_ankle <- mechanical_load_data %>%
   ) +
   geom_smooth(
     aes(x = pAR_gs, y = pLR_Ns, color = BMI_cat),
-    method = "lm", se = FALSE
+    method = "lm", se = FALSE, show.legend = FALSE
   ) +
   scale_color_nejm() +
   scale_y_continuous(
+    labels = scales::label_number(),
     limits = c(0, 60000),
     expand = c(0, 0),
     breaks = seq(0, 60000, 10000)
@@ -39,17 +40,18 @@ scatterplot_LR_res_ankle <- mechanical_load_data %>%
   scale_x_continuous(
     limits = c(0, 400),
     expand = c(0, 0),
-    breaks = seq(50, 350, 50)
+    breaks = seq(0, 400, 50)
   ) +
   theme_light() +
   theme(
-    plot.title = element_text(size = 15, hjust = 0.5),
-    legend.title = element_text(size = 13),
-    legend.text = element_text(size = 12),
-    axis.title.y = element_text(size = 13),
-    axis.title.x = element_text(size = 13),
-    axis.text.y = element_text(size = 13),
-    axis.text.x = element_text(size = 13)
+    plot.title = element_text(size = 17, hjust = 0.5),
+    legend.title = element_text(size = 15),
+    legend.text = element_text(size = 15),
+    axis.title.y = element_text(size = 15),
+    axis.title.x = element_text(size = 15),
+    axis.text.y = element_text(size = 15),
+    axis.text.x = element_text(size = 15, vjust = 0.2),
+    plot.margin = margin(r = 1, unit = "cm")
   ) +
   guides(
     color = guide_legend(title = "Body mass index category:"),
@@ -70,10 +72,11 @@ scatterplot_LR_res_back <- mechanical_load_data %>%
   ) +
   geom_smooth(
     aes(x = pAR_gs, y = pLR_Ns, color = BMI_cat),
-    method = "lm", se = FALSE
+    method = "lm", se = FALSE, show.legend = FALSE
   ) +
   scale_color_nejm() +
   scale_y_continuous(
+    labels = scales::label_number(),
     limits = c(0, 60000),
     expand = c(0, 0),
     breaks = seq(0, 60000, 10000)
@@ -81,17 +84,18 @@ scatterplot_LR_res_back <- mechanical_load_data %>%
   scale_x_continuous(
     limits = c(0, 250),
     expand = c(0, 0),
-    breaks = seq(50, 200, 50)
[…]

### 2022-03-09 figures/figS3.R

@@ -43,19 +43,20 @@ BA_GRF_res_ankle <- cv_res_GRF_models$ankle %>%
     breaks = seq(-600, 600, 300)
   ) +
   scale_x_continuous(
-    limits = c(400, 3100),
+    limits = c(500, 3000),
     expand = c(0, 0),
     breaks = seq(500, 3000, 500)
   ) +
   theme_light() +
   theme(
-    plot.title = element_text(size = 15, hjust = 0.5),
-    legend.title = element_text(size = 13),
-    legend.text = element_text(size = 12),
-    axis.title.y = element_text(size = 13),
-    axis.title.x = element_text(size = 13),
-    axis.text.y = element_text(size = 13),
-    axis.text.x = element_text(size = 13)
+    plot.title = element_text(size = 17, hjust = 0.5),
+    legend.title = element_text(size = 15),
+    legend.text = element_text(size = 15),
+    axis.title.y = element_text(size = 15),
+    axis.title.x = element_text(size = 15),
+    axis.text.y = element_text(size = 15),
+    axis.text.x = element_text(size = 15, vjust = 0.2),
+    plot.margin = margin(r = 1, unit = "cm")
   ) +
   guides(
     color = guide_legend(title = "Body mass index category:"),
@@ -78,19 +79,20 @@ BA_GRF_res_back <- cv_res_GRF_models$lower_back %>%
     breaks = seq(-600, 600, 200)
   ) +
   scale_x_continuous(
-    limits = c(400, 3100),
+    limits = c(500, 3000),
     expand = c(0, 0),
     breaks = seq(500, 3000, 500)
   ) +
   theme_light() +
   theme(
-    plot.title = element_text(size = 15, hjust = 0.5),
-    legend.title = element_text(size = 13),
-    legend.text = element_text(size = 12),
-    axis.title.y = element_text(size = 13),
-    axis.title.x = element_text(size = 13),
-    axis.text.y = element_text(size = 13),
-    axis.text.x = element_text(size = 13)
+    plot.title = element_text(size = 17, hjust = 0.5),
+    legend.title = element_text(size = 15),
+    legend.text = element_text(size = 15),
+    axis.title.y = element_text(size = 15),
+    axis.title.x = element_text(size = 15),
+    axis.text.y = element_text(size = 15),
+    axis.text.x = element_text(size = 15, vjust = 0.2),
+    plot.margin = margin(r = 1, unit = "cm")
   ) +
[…]

### 2022-03-09 figures/figS4.R

@@ -44,19 +44,20 @@ BA_LR_res_ankle <- cv_res_LR_models$ankle %>%
     breaks = seq(-30000, 30000, 10000)
   ) +
   scale_x_continuous(
-    limits = c(0, 55000),
+    limits = c(0, 50000),
     expand = c(0, 0),
-    breaks = seq(10000, 50000, 10000)
+    breaks = seq(0, 50000, 10000)
   ) +
   theme_light() +
   theme(
-    plot.title = element_text(size = 15, hjust = 0.5),
-    legend.title = element_text(size = 13),
-    legend.text = element_text(size = 12),
-    axis.title.y = element_text(size = 13),
-    axis.title.x = element_text(size = 13),
-    axis.text.y = element_text(size = 13),
-    axis.text.x = element_text(size = 13)
+    plot.title = element_text(size = 17, hjust = 0.5),
+    legend.title = element_text(size = 15),
+    legend.text = element_text(size = 15),
+    axis.title.y = element_text(size = 15),
+    axis.title.x = element_text(size = 15),
+    axis.text.y = element_text(size = 15),
+    axis.text.x = element_text(size = 15, vjust = 0.2),
+    plot.margin = margin(r = 1, unit = "cm")
   ) +
   guides(
     color = guide_legend(title = "Body mass index category:"),
@@ -80,19 +81,20 @@ BA_LR_res_back <- cv_res_LR_models$lower_back %>%
     breaks = seq(-30000, 30000, 10000)
   ) +
   scale_x_continuous(
-    limits = c(0, 55000),
+    limits = c(0, 50000),
     expand = c(0, 0),
-    breaks = seq(10000, 50000, 10000)
+    breaks = seq(0, 50000, 10000)
   ) +
   theme_light() +
   theme(
-    plot.title = element_text(size = 15, hjust = 0.5),
-    legend.title = element_text(size = 13),
-    legend.text = element_text(size = 12),
-    axis.title.y = element_text(size = 13),
-    axis.title.x = element_text(size = 13),
-    axis.text.y = element_text(size = 13),
-    axis.text.x = element_text(size = 13)
+    plot.title = element_text(size = 17, hjust = 0.5),
+    legend.title = element_text(size = 15),
+    legend.text = element_text(size = 15),
+    axis.title.y = element_text(size = 15),
+    axis.title.x = element_text(size = 15),
+    axis.text.y = element_text(size = 15),
[…]

### 2022-03-09 figures/fig1.R
Save figures in .png to include in manuscript.pdf

@@ -69,3 +69,14 @@ agg_tiff(
 )
 plot(fig1)
 dev.off()
+
+agg_png(
+  here("figures", "fig1.png"),
+  width = 90,
+  height = 50,
+  units = "cm",
+  res = 100,
+  scaling = 2
+)
+plot(fig1)
+dev.off()

### 2022-03-09 figures/fig2.R

@@ -228,3 +228,14 @@ agg_tiff(
 )
 plot(fig2)
 dev.off()
+
+agg_png(
+  here("figures", "fig2.png"),
+  width = 90,
+  height = 50,
+  units = "cm",
+  res = 100,
+  scaling = 2
+)
+plot(fig2)
+dev.off()

### 2022-03-09 figures/figS1.R

@@ -308,3 +308,14 @@ agg_tiff(
 )
 plot(figS1)
 dev.off()
+
+agg_png(
+  here("figures", "figS1.png"),
+  width = 120,
+  height = 50,
+  units = "cm",
+  res = 100,
+  scaling = 2
+)
+plot(figS1)
+dev.off()

### 2022-03-09 figures/figS2.R

@@ -308,3 +308,14 @@ agg_tiff(
 )
 plot(figS2)
 dev.off()
+
+agg_png(
+  here("figures", "figS2.png"),
+  width = 120,
+  height = 50,
+  units = "cm",
+  res = 100,
+  scaling = 2
+)
+plot(figS2)
+dev.off()

### 2022-03-09 figures/figS3.R

@@ -274,3 +274,14 @@ agg_tiff(
 )
 plot(figS3)
 dev.off()
+
+agg_png(
+  here("figures", "figS3.png"),
+  width = 120,
+  height = 50,
+  units = "cm",
+  res = 100,
+  scaling = 2
+)
+plot(figS3)
+dev.off()

### 2022-03-09 figures/figS4.R

@@ -280,3 +280,14 @@ agg_tiff(
 )
 plot(figS4)
 dev.off()
+
+agg_png(
+  here("figures", "figS4.png"),
+  width = 120,
+  height = 50,
+  units = "cm",
+  res = 100,
+  scaling = 2
+)
+plot(figS4)
+dev.off()

### 2022-03-09 manuscript/manuscript.Rmd
Update figures

@@
 Afterwards, Bland-Altman plots were built to visualize the agreement between the actual and predicted pGRF and pLR, for both the resultant vector and its vertical component from all accelerometer placements.
-Bland-Altman plots for the pGRF and pLR resultant vector from hip-worn accelerometers can be seen in Figure 3, and for all accelerometer placements and both resultant and vertical vectors in the Supplemental Figures S3 and S4.
+Bland-Altman plots for the pGRF and pLR resultant vector from hip-worn accelerometers can be seen in Figure 2 (Panels C and D, respectively), and for all accelerometer placements and both resultant and vertical vectors in the Supplemental Figures S3 and S4.
 Values for all pGRF and pLR prediction equations developed tended to assemble around zero and between the limits of agreement, as shown by the fact that, for all models, at least `r ba_agreement`\% of the data points lie within the limits of agreement.
@@
 Although all equations showed a significant proportional bias (*p* < 0.05), the linear regressions presented a low R^2^, with the highest value being `r myround(max(ba_regression_ankle_R2), 2)`.
-
-\begin{figure}[ht]
-	\captionsetup{singlelinecheck = false, format = hang, justification = raggedright, font = footnotesize, labelsep = space}
-	\centering
-	\begin{measuredfigure}
-		\includegraphics[width=\textwidth]{fig3.png}
-		\caption{Bland-Altman plots showing agreement between actual and predicted peak ground reaction force (Panel A) and loading rate (Panel B) for the resultant vector from hip-worn accelerometers.
-Continuous thick lines show bias (average of the differences between actual and predicted values) while dashed lines show the limits of agreement (bias \pm 1.96 standard deviation).
-Abbreviations: pGRF, peak ground reaction force; pLR, peak loading rate.}
-	\end{measuredfigure}
-\end{figure}
-
-\medskip
 

### 2022-03-10 manuscript/manuscript.Rmd
Update figures

@@
 During locomotion, pACC increases linearly as pGRF increases.
-Such relation can be observed in Figure 2, Panel A, for hip-worn accelerometers and the resultant vector, and in Supplemental Figure S1 for all placements and the resultant vector and its vertical component.
+Such relation can be observed in Figure 2, Panel A, for hip-worn accelerometers and the resultant vector, and in Supplemental Figure S2 for all placements and the resultant vector and its vertical component.
 Moreover, for the same pACC values, the registered pGRF tended to be higher in subjects with higher BMI.
 This demonstrates that accelerometer data can adequately discriminate between different BMI categories.
-However, data from the pAR x pLR relationship show a greater dispersion, especially at higher values, and the different BMI categories cannot be discriminated so well, as can be observed in Figure 2, Panel B, and Supplemental Figure S2.
+However, data from the pAR x pLR relationship show a greater dispersion, especially at higher values, and the different BMI categories cannot be discriminated so well, as can be observed in Figure 2, Panel B, and Supplemental Figure S3.
 
@@
 Afterwards, Bland-Altman plots were built to visualize the agreement between the actual and predicted pGRF and pLR, for both the resultant vector and its vertical component from all accelerometer placements.
-Bland-Altman plots for the pGRF and pLR resultant vector from hip-worn accelerometers can be seen in Figure 2 (Panels C and D, respectively), and for all accelerometer placements and both resultant and vertical vectors in the Supplemental Figures S3 and S4.
+Bland-Altman plots for the pGRF and pLR resultant vector from hip-worn accelerometers can be seen in Figure 2 (Panels C and D, respectively), and for all accelerometer placements and both resultant and vertical vectors in the Supplemental Figures S4 and S5.
 Values for all pGRF and pLR prediction equations developed tended to assemble around zero and between the limits of agreement, as shown by the fact that, for all models, at least `r ba_agreement`\% of the data points lie within the limits of agreement.
-For the pGRF equations, a tendency to higher dispersion can be observed for higher magnitudes, which correspond to the running trials, and especially in values derived from ankle-worn accelerometers (Supplemental Figure S3, Panels A and D).
-Prediction equations for pLR showed a more accentuated dispersion, and again mainly for higher magnitudes (Supplemental Figure S4).
+For the pGRF equations, a tendency to higher dispersion can be observed for higher magnitudes, which correspond to the running trials, and especially in values derived from ankle-worn accelerometers (Supplemental Figure S4, Panels A and D).
+Prediction equations for pLR showed a more accentuated dispersion, and again mainly for higher magnitudes (Supplemental Figure S5).
 None of the models presented a bias different from zero (*p* > 0.05), demonstrating that they do not consistently over or underestimate the pGRF and pLR values.
@@
 
-## Supplemental Figure S1 {.unnumbered}
 \begin{figure}[ht]
@@
 \end{figure}
-\footnotesize Relationship between peak ground reaction force and acceleration magnitudes.
-Data from accelerometers worn at ankle (Panels A and D), lower back (Panels B and E) and hip (Panels C and F), for both the resultant vector (Panels A to C) and its vertical component (Panels D to F).
-Linear trends for each body mass index category are also depicted.
-Abbreviations: pACC, peak acceleration; pGRF, peak ground reaction force.
+\footnotesize \textbf{Supplemental Figure S1.} Example of the ground reaction force and acceleration signals during the stance phase.
+Signals were mean centered and scaled to units of standard deviation to fit in the same plot window.
+In this image, the peak of each signal (star mark) represents the maximum impact between the foot and the ground.
+The circles mark the beginning of the curve, i.e., the start of the ascending phase of the signal that culminates in its peak.
+The region of the acceleration signal between the circle and the star marks was used to compute the acceleration rate, and the same region on the ground reaction force signal was used to compute the loading rate.
 
-\medskip
-
-## Supplemental Figure S2 {.unnumbered}
 \begin{figure}[ht]
@@
 \end{figure}
-\footnotesize Relationship between peak loading and acceleration rates.
+\footnotesize \textbf{Supplemental Figure S2.} Relationship between peak ground reaction force and acceleration magnitudes.
 Data from accelerometers worn at ankle (Panels A and D), lower back (Panels B and E) and hip (Panels C and F), for both the resultant vector (Panels A to C) and its vertical component (Panels D to F).
 Linear trends for each body mass index category are also depicted.
-Abbreviations: pAR, peak acceleration rate; pLR, peak loading rate.
+Abbreviations: pACC, peak acceleration; pGRF, peak ground reaction force.
 
-\newpage
-
-## Supplemental Figure S3 {.unnumbered}
 \begin{figure}[ht]
@@
 \end{figure}
-\footnotesize Bland-Altman plots showing agreement between actual and predicted peak ground reaction force for accelerometers worn at ankle (Panels A and D), lower back (Panels B and E) and hip (Panels C and F), for both the resultant vector (Panels A to C) and its vertical component (Panels D to F).
-Continuous thick lines show bias (average of the differences between actual and predicted values) while dashed lines show the limits of agreement (bias \pm 1.96 standard deviation).
-Abbreviations: pGRF, peak ground reaction force.
+\footnotesize \textbf{Supplemental Figure S3.} Relationship between peak loading and acceleration rates.
+Data from accelerometers worn at ankle (Panels A and D), lower back (Panels B and E) and hip (Panels C and F), for both the resultant vector (Panels A to C) and its vertical component (Panels D to F).
+Linear trends for each body mass index category are also depicted.
+Abbreviations: pAR, peak acceleration rate; pLR, peak loading rate.
 
-## Supplemental Figure S4 {.unnumbered}
 \begin{figure}[ht]
@@
 \end{figure}
-\footnotesize Bland-Altman plots showing agreement between actual and predicted peak loading rate for accelerometers worn at ankle (Panels A and D), lower back (Panels B and E) and hip (Panels C and F), for both the resultant vector (Panels A to C) and its vertical component (Panels D to F).
+\footnotesize \textbf{Supplemental Figure S4.} Bland-Altman plots showing agreement between actual and predicted peak ground reaction force for accelerometers worn at ankle (Panels A and D), lower back (Panels B and E) and hip (Panels C and F), for both the resultant vector (Panels A to C) and its vertical component (Panels D to F).
+Continuous thick lines show bias (average of the differences between actual and predicted values) while dashed lines show the limits of agreement (bias \pm 1.96 standard deviation).
+Abbreviations: pGRF, peak ground reaction force.
+
+\begin{figure}[ht]
+	\captionsetup{singlelinecheck = false, format = hang, justification = raggedright, font = footnotesize, labelsep = space}
+	\centering
+	\begin{measuredfigure}
+		\includegraphics[width=\textwidth]{figS5.png}
+	\end{measuredfigure}
+\end{figure}
+\footnotesize \textbf{Supplemental Figure S5.} Bland-Altman plots showing agreement between actual and predicted peak loading rate for accelerometers worn at ankle (Panels A and D), lower back (Panels B and E) and hip (Panels C and F), for both the resultant vector (Panels A to C) and its vertical component (Panels D to F).
 Continuous thick lines show bias (average of the differences between actual and predicted values) while dashed lines show the limits of agreement (bias ± 1.96 standard deviation).

### 2022-03-10 figures/figS2.R
Update figure names

@@ -17,30 +17,29 @@ mechanical_load_data <- mechanical_load_data %>%
     )
   )
 
-# LR x AR plots -----------------------------------------------------------
+# GRF x ACC plots ---------------------------------------------------------
 
 # Resultant: Ankle
-scatterplot_LR_res_ankle <- mechanical_load_data %>%
+scatterplot_GRF_res_ankle <- mechanical_load_data %>%
   filter(vector == "resultant" & acc_placement == "ankle") %>%
   ggplot() +
   geom_point(
-    aes(x = pAR_gs, y = pLR_Ns, color = BMI_cat, shape = activity), alpha = 0.5
+    aes(x = pACC_g, y = pGRF_N, color = BMI_cat, shape = activity), alpha = 0.5
   ) +
   geom_smooth(
-    aes(x = pAR_gs, y = pLR_Ns, color = BMI_cat),
+    aes(x = pACC_g, y = pGRF_N, color = BMI_cat),
     method = "lm", se = FALSE, show.legend = FALSE
   ) +
   scale_color_nejm() +
   scale_y_continuous(
-    labels = scales::label_number(),
-    limits = c(0, 60000),
+    limits = c(0, 3500),
     expand = c(0, 0),
-    breaks = seq(0, 60000, 10000)
+    breaks = seq(0, 3500, 500)
   ) +
   scale_x_continuous(
-    limits = c(0, 400),
+    limits = c(0, 14),
     expand = c(0, 0),
-    breaks = seq(0, 400, 50)
+    breaks = seq(0, 14, 2)
   ) +
   theme_light() +
   theme(
@@ -55,36 +54,36 @@ scatterplot_LR_res_ankle <- mechanical_load_data %>%
   ) +
   guides(
     color = guide_legend(title = "Body mass index category:"),
-    shape = guide_legend(title = "Locomotion type:")
+    shape = guide_legend(title = "Locomotion type:"),
+    alpha = FALSE
   ) +
   labs(
     title = "Resultant vector - Ankle placement",
-    x = quote("pAR" ~ (italic(g) %.% s^-1)),
-    y = quote("pLR" ~ (N %.% s^-1))
+    x = quote("pACC" ~ (italic(g))),
+    y = "pGRF (N)"
   )
 
 # Resultant: Lower back
-scatterplot_LR_res_back <- mechanical_load_data %>%
[…]

### 2022-03-10 figures/figS3.R

@@ -2,50 +2,45 @@
 
 library(here)
 library(tidyverse)
-library(lvmisc)
 library(ggsci)
 library(patchwork)
 library(ragg)
 
 # Load data ---------------------------------------------------------------
 
-load(here("output", "loocv_data.rda"))
-cv_res_GRF_models <- map(
-  cv_res_GRF_models,
-  ~ mutate(
-    .x,
+load(here("data", "mechanical_load_data.rda"))
+mechanical_load_data <- mechanical_load_data %>%
+  mutate(
     activity = fct_relevel(
       as.factor(ifelse(speed %in% 1:6, "Walking", "Running")),
       "Walking"
     )
   )
-)
-cv_ver_GRF_models <- map(
-  cv_ver_GRF_models,
-  ~ mutate(
-    .x,
-    activity = fct_relevel(
-      as.factor(ifelse(speed %in% 1:6, "Walking", "Running")),
-      "Walking"
-    )
-  )
-)
 
-# GRF plots ---------------------------------------------------------------
+# LR x AR plots -----------------------------------------------------------
 
 # Resultant: Ankle
-BA_GRF_res_ankle <- cv_res_GRF_models$ankle %>%
-  plot_bland_altman(color = BMI_cat, shape = activity, alpha = 0.5) +
+scatterplot_LR_res_ankle <- mechanical_load_data %>%
+  filter(vector == "resultant" & acc_placement == "ankle") %>%
+  ggplot() +
+  geom_point(
+    aes(x = pAR_gs, y = pLR_Ns, color = BMI_cat, shape = activity), alpha = 0.5
+  ) +
+  geom_smooth(
+    aes(x = pAR_gs, y = pLR_Ns, color = BMI_cat),
+    method = "lm", se = FALSE, show.legend = FALSE
+  ) +
   scale_color_nejm() +
   scale_y_continuous(
-    limits = c(-600, 600),
+    labels = scales::label_number(),
+    limits = c(0, 60000),
     expand = c(0, 0),
-    breaks = seq(-600, 600, 300)
+    breaks = seq(0, 60000, 10000)
   ) +
   scale_x_continuous(
-    limits = c(500, 3000),
+    limits = c(0, 400),
     expand = c(0, 0),
-    breaks = seq(500, 3000, 500)
+    breaks = seq(0, 400, 50)
   ) +
[…]

### 2022-03-10 figures/figS4.R

@@ -10,8 +10,8 @@ library(ragg)
 # Load data ---------------------------------------------------------------
 
 load(here("output", "loocv_data.rda"))
-cv_res_LR_models <- map(
-  cv_res_LR_models,
+cv_res_GRF_models <- map(
+  cv_res_GRF_models,
   ~ mutate(
     .x,
     activity = fct_relevel(
@@ -20,8 +20,8 @@ cv_res_LR_models <- map(
     )
   )
 )
-cv_ver_LR_models <- map(
-  cv_ver_LR_models,
+cv_ver_GRF_models <- map(
+  cv_ver_GRF_models,
   ~ mutate(
     .x,
     activity = fct_relevel(
@@ -31,22 +31,21 @@ cv_ver_LR_models <- map(
   )
 )
 
-# LR plots ----------------------------------------------------------------
+# GRF plots ---------------------------------------------------------------
 
 # Resultant: Ankle
-BA_LR_res_ankle <- cv_res_LR_models$ankle %>%
+BA_GRF_res_ankle <- cv_res_GRF_models$ankle %>%
   plot_bland_altman(color = BMI_cat, shape = activity, alpha = 0.5) +
   scale_color_nejm() +
   scale_y_continuous(
-    labels = scales::label_number(),
-    limits = c(-30000, 30000),
+    limits = c(-600, 600),
     expand = c(0, 0),
-    breaks = seq(-30000, 30000, 10000)
+    breaks = seq(-600, 600, 300)
   ) +
   scale_x_continuous(
-    limits = c(0, 50000),
+    limits = c(500, 3000),
     expand = c(0, 0),
-    breaks = seq(0, 50000, 10000)
+    breaks = seq(500, 3000, 500)
   ) +
   theme_light() +
   theme(
@@ -66,24 +65,23 @@ BA_LR_res_ankle <- cv_res_LR_models$ankle %>%
   ) +
   labs(
     title = "Resultant vector - Ankle placement",
-    x = quote("Mean of Actual and Predicted pLR" ~ (N %.% s^-1)),
-    y = quote("Actual - Predicted pLR" ~ (N %.% s^-1))
+    x = "Mean of Actual and Predicted pGRF (N)",
+    y = "Actual - Predicted pGRF (N)"
   )
 
 # Resultant: Lower back
-BA_LR_res_back <- cv_res_LR_models$lower_back %>%
[…]

### 2022-03-10 figures/figS5.R

@@ -2,44 +2,51 @@
 
 library(here)
 library(tidyverse)
+library(lvmisc)
 library(ggsci)
 library(patchwork)
 library(ragg)
 
 # Load data ---------------------------------------------------------------
 
-load(here("data", "mechanical_load_data.rda"))
-mechanical_load_data <- mechanical_load_data %>%
-  mutate(
+load(here("output", "loocv_data.rda"))
+cv_res_LR_models <- map(
+  cv_res_LR_models,
+  ~ mutate(
+    .x,
     activity = fct_relevel(
       as.factor(ifelse(speed %in% 1:6, "Walking", "Running")),
       "Walking"
     )
   )
+)
+cv_ver_LR_models <- map(
+  cv_ver_LR_models,
+  ~ mutate(
+    .x,
+    activity = fct_relevel(
+      as.factor(ifelse(speed %in% 1:6, "Walking", "Running")),
+      "Walking"
+    )
+  )
+)
 
-# GRF x ACC plots ---------------------------------------------------------
+# LR plots ----------------------------------------------------------------
 
 # Resultant: Ankle
-scatterplot_GRF_res_ankle <- mechanical_load_data %>%
-  filter(vector == "resultant" & acc_placement == "ankle") %>%
-  ggplot() +
-  geom_point(
-    aes(x = pACC_g, y = pGRF_N, color = BMI_cat, shape = activity), alpha = 0.5
-  ) +
-  geom_smooth(
-    aes(x = pACC_g, y = pGRF_N, color = BMI_cat),
-    method = "lm", se = FALSE, show.legend = FALSE
-  ) +
+BA_LR_res_ankle <- cv_res_LR_models$ankle %>%
+  plot_bland_altman(color = BMI_cat, shape = activity, alpha = 0.5) +
   scale_color_nejm() +
   scale_y_continuous(
-    limits = c(0, 3500),
+    labels = scales::label_number(),
+    limits = c(-30000, 30000),
     expand = c(0, 0),
-    breaks = seq(0, 3500, 500)
+    breaks = seq(-30000, 30000, 10000)
   ) +
   scale_x_continuous(
-    limits = c(0, 14),
+    limits = c(0, 50000),
     expand = c(0, 0),
-    breaks = seq(0, 14, 2)
[…]

### 2022-03-10 figures/figS2.R
Update figure names

@@ -285,7 +285,7 @@ scatterplot_GRF_ver_hip <- mechanical_load_data %>%
 
 # Combine and save plots --------------------------------------------------
 
-figS1 <- scatterplot_GRF_res_ankle +
+figS2 <- scatterplot_GRF_res_ankle +
   scatterplot_GRF_res_back +
   scatterplot_GRF_res_hip +
   scatterplot_GRF_ver_ankle +
@@ -299,23 +299,23 @@ figS1 <- scatterplot_GRF_res_ankle +
   )
 
 agg_tiff(
-  here("figures", "figS1.tiff"),
+  here("figures", "figS2.tiff"),
   width = 120,
   height = 50,
   units = "cm",
   res = 100,
   scaling = 2
 )
-plot(figS1)
+plot(figS2)
 dev.off()
 
 agg_png(
-  here("figures", "figS1.png"),
+  here("figures", "figS2.png"),
   width = 120,
   height = 50,
   units = "cm",
   res = 100,
   scaling = 2
 )
-plot(figS1)
+plot(figS2)
 dev.off()

### 2022-03-10 figures/figS3.R

@@ -285,7 +285,7 @@ scatterplot_LR_ver_hip <- mechanical_load_data %>%
 
 # Combine and save plots --------------------------------------------------
 
-figS2 <- scatterplot_LR_res_ankle +
+figS3 <- scatterplot_LR_res_ankle +
   scatterplot_LR_res_back +
   scatterplot_LR_res_hip +
   scatterplot_LR_ver_ankle +
@@ -299,23 +299,23 @@ figS2 <- scatterplot_LR_res_ankle +
   )
 
 agg_tiff(
-  here("figures", "figS2.tiff"),
+  here("figures", "figS3.tiff"),
   width = 120,
   height = 50,
   units = "cm",
   res = 100,
   scaling = 2
 )
-plot(figS2)
+plot(figS3)
 dev.off()
 
 agg_png(
-  here("figures", "figS2.png"),
+  here("figures", "figS3.png"),
   width = 120,
   height = 50,
   units = "cm",
   res = 100,
   scaling = 2
 )
-plot(figS2)
+plot(figS3)
 dev.off()

### 2022-03-10 figures/figS4.R

@@ -251,7 +251,7 @@ BA_GRF_ver_hip <- cv_ver_GRF_models$hip %>%
 
 # Combine and save plots --------------------------------------------------
 
-figS3 <- BA_GRF_res_ankle +
+figS4 <- BA_GRF_res_ankle +
   BA_GRF_res_back +
   BA_GRF_res_hip +
   BA_GRF_ver_ankle +
@@ -265,23 +265,23 @@ figS3 <- BA_GRF_res_ankle +
   )
 
 agg_tiff(
-  here("figures", "figS3.tiff"),
+  here("figures", "figS4.tiff"),
   width = 120,
   height = 50,
   units = "cm",
   res = 100,
   scaling = 2
 )
-plot(figS3)
+plot(figS4)
 dev.off()
 
 agg_png(
-  here("figures", "figS3.png"),
+  here("figures", "figS4.png"),
   width = 120,
   height = 50,
   units = "cm",
   res = 100,
   scaling = 2
 )
-plot(figS3)
+plot(figS4)
 dev.off()

### 2022-03-10 figures/figS5.R

@@ -257,7 +257,7 @@ BA_LR_ver_hip <- cv_ver_LR_models$hip %>%
 
 # Combine and save plots --------------------------------------------------
 
-figS4 <- BA_LR_res_ankle +
+figS5 <- BA_LR_res_ankle +
   BA_LR_res_back +
   BA_LR_res_hip +
   BA_LR_ver_ankle +
@@ -271,23 +271,23 @@ figS4 <- BA_LR_res_ankle +
   )
 
 agg_tiff(
-  here("figures", "figS4.tiff"),
+  here("figures", "figS5.tiff"),
   width = 120,
   height = 50,
   units = "cm",
   res = 100,
   scaling = 2
 )
-plot(figS4)
+plot(figS5)
 dev.off()
 
 agg_png(
-  here("figures", "figS4.png"),
+  here("figures", "figS5.png"),
   width = 120,
   height = 50,
   units = "cm",
   res = 100,
   scaling = 2
 )
-plot(figS4)
+plot(figS5)
 dev.off()

### 2022-03-12 manuscript/manuscript.Rmd
Improve suppl figs formating
Add reference to Suppl Fig 1 and update its caption

@@
 The peak loading rate (pLR) and peak acceleration rate (pAR) were defined as the maximum value in the array (Equation 2).
+An example of GRF and ACC curves during the stance phase of both walking and running can be seen in Supplemental Figure 1.
 
@@
 \end{figure}
-\footnotesize \textbf{Supplemental Figure S1.} Example of the ground reaction force and acceleration signals during the stance phase.
+\footnotesize \textbf{Supplemental Figure S1.} Example of ground reaction force and acceleration signals during the stance phase of walking (left panel) and running (right panel).
 Signals were mean centered and scaled to units of standard deviation to fit in the same plot window.
-In this image, the peak of each signal (star mark) represents the maximum impact between the foot and the ground.
-The circles mark the beginning of the curve, i.e., the start of the ascending phase of the signal that culminates in its peak.
-The region of the acceleration signal between the circle and the star marks was used to compute the acceleration rate, and the same region on the ground reaction force signal was used to compute the loading rate.
+In this image, the peak of each signal, represented by the start marks, is its maximum magnitude during the stance.
+The peak acceleration rate and peak loading rate were identified as the steepest part of each curve.
+
+\newpage
 
@@
 
+\newpage
+
 \begin{figure}[ht]

### 2022-05-28 manuscript/manuscript.Rmd
Revise the manuscript

@@
 
+To the best of our knowledge, only two studies have already attempted to develop pLR prediction models based on accelerometry data, and both studies only included walking activities.
+In accordance with our own findings, these studies also showed that the prediction error for these models is considerably larger than that for the pGRF prediction.
+One of these studies investigated the use of four different accelerometer placements (ankle, tibia, thigh and waist) to predict pLR and achieved a MAPE ranging from 23.4\% to 46.1\% [@Fortune2014], while the other obtained values from 19.5\% to 22.4\% using accelerometers at the lower back and hip [@Veras2020].
+Results from the models developed in these previous studies have both lower and higher prediction errors compared to our own results, which were of approximately 35\%.
+Although pLR is an important variable, known for a long time to determine to a great extent the bone formation response to mechanical loading [@Bassey1998; @Turner1995], based on our findings and in previous studies [@Fortune2014; @Veras2020], its prediction through accelerometer data, at the current state of technical accelerometer development, does not seem to yield trustworthy results that can be used in models to predict the bone anabolic response to mechanical loading.
+
+\medskip
+
 Our findings revealed that ankle, lower back or hip were all valid accelerometer placements to accurately predict pGRF.
@@
 
-To the best of our knowledge, only two studies have already attempted to develop pLR prediction models based on accelerometry data, and both studies only included walking activities.
-In accordance with our own findings, these studies also showed that the prediction error for these models is considerably larger than that for the pGRF prediction.
-One of these studies investigated the use of four different accelerometer placements (ankle, tibia, thigh and waist) to predict pLR and achieved a MAPE ranging from 23.4\% to 46.1\% [@Fortune2014], while the other obtained values from 19.5\% to 22.4\% using accelerometers at the lower back and hip [@Veras2020].
-Results from the models developed in these previous studies have both lower and higher prediction errors compared to our own results, which were of approximately 35\%.
-Although pLR is an important variable, known for a long time to determine to a great extent the bone formation response to mechanical loading [@Bassey1998; @Turner1995], based on our findings and in previous studies [@Fortune2014; @Veras2020], its prediction through accelerometer data, at the current state of technical accelerometer development, does not seem to yield trustworthy results that can be used in models to predict the bone anabolic response to mechanical loading.
+It is also important to highlight that, during peak detection, it was observed a mismatch in the stance phase in which the pACC and pGRF are detected.
+This mismatch occurred particularly at running speeds, in which the pACC is roughly at 20% of the stance and the pGRF at around 50% of the stance.
+Nevertheless, although we have detected this mismatch, and, effectively, the pACC and pGRF that were detected do not represent the same event in the stance phase, by using the prediction method developed in this paper we could accurately predict the active pGRF at both walking and running.
 
